### 2017-05-23 chronology.tex
add a chronology of things done

\documentclass[a4paper,11pt,oneside,onecolumn,openany,final,svgnames]{memoir}
\input{packages}
\input{abbreviations}
\title{PhD plans}
\title{Géraud Le Falher}
\begin{document}
\maketitle
\begin{description}
	\item[Oct. 14] State of the art for \pcc{} (see
		\href{run:support/cc.pdf}{A few words on Correlation Clustering})
	\item[Nov. 14 --- Jan. 15] Trying different ways of closing the
		triangles of a general graph without introducing bad cycle in
		order to transfer the \ccPivot{} algorithm (see
		\href{run:support/december.pdf}{Combinatorial
		Correlation Clustering on general unweighted graphs} and the Dating presentation
		\href{run:support/presentationeng.pdf}{Signed graphs: clustering and link prediction})
	\item[Jan. 15 --- Apr. 15]{Predicting link signs by building a low
		stretch spanning tree dubbed \gtx{}, which works by extracting
		stars, connecting them in a graph, and repeating on this reduced graph
		(see \href{run:support/nips2015.pdf}{some experimental results}).}
	\item[Mar. 15] Visit to Claudio. The setting was a graph where nodes have profiles and be
		similar along various ways, and the problems were either answering online
		similarity queries or recovering hidden profiles (see
		\href{run:support/contextual_clustering.pdf}{Contextual Clustering}).
	\item[May. 15] Implementing \shz{}~\autocite{Vitale2012} (actually it took a week, but I wasn't so
		much in the lab in May)
	\item[Jun. 15 --- Aug. 15] 
		I supervised the internship of Paul Dennetiere on the implementation of a parallel
		algorithm for \pcc{} described by \textcite{Pan2014} (see the
		\href{run:support/stage_cc.pdf}{internship description}, a
		\href{run:support/draft_paul.pdf}{draft of the technical part of Paul's report} and
		some ideas on \href{run:support/parallel_postprocessing.pdf}{parallel post
		processing} we didn't have time to explore).

		I worked on building trees that minimized the stretch of the distances between a
		set of revealed nodes and the other ones (see a
		\href{run:support/sujet2_lowstretch.pdf}{a MVA project topic} and
		\href{run:support/constrained_low_stretch.pdf}{some initial experiments}).

		Two others summaries of what I did until then can be found in my
		\href{run:support/postermlss.pdf}{MLSS poster} and my
		\href{run:support/progress.pdf}{first year progress report}.
	\item[Sep. 15]
		When Fabio came back from Finland, he told me about a problem on graph
		reconstruction. I can't fully remember the setting, but I think the general idea
		was to make some queries about the existence of certain edges. Trivially, by
		making $\Theta(|V|^2)$ queries, we know the full topology so the question of
		was can we do better. I didn't write much about except
		\href{run:support/permutation.pdf}{one thing about counting permutation on a
		line graph}, but this is very incomplete.
	\item[Oct. 15 --- Nov. 15] Following Michael's visit during the summer, we
		worked on a binary node classification problem where the bias is that a node
		is labeled $+1$ iff at least two of its neighbors are labeled $+1$. This
		lead to a propagation algorithm based on triangles chain (see
		\href{run:support/NodeClassification_triangle_mail_fabio_nov15.pdf}{some Fabio writeup on the topic}).
	\item[Nov. 15 --- May. 16] Edge classification in the troll trust model (see
		the \href{https://arxiv.org/pdf/1606.00182.pdf}{arxiv version of the AISTATS paper})
	\item[Jun.16 --- Dec. 16] Edge clustering
	\item[Dec. 16] Students of the MVA did a project on \pcc{} under perturbation,
		here is the \href{run:support/sujet1_ccpert.pdf}{topic description} and
		\href{run:support/Dubreuil_Ramzi_report.pdf}{their report}.
	\item[Jan. 17 --- Feb. 17] We worked with Fabio on a more robust version of
		the \shz{} algorithm that decreases its confidence to labeled nodes when
		they are implicated in further prediction mistake but
		\href{run:support/trees.pdf}{experimental results} are not very satisfying.
	\item[Mar. 17 --- May. 17] Edge clustering
\end{description}

During my thesis, I also published my Master's thesis in a
conference~\autocite{Thesis15} and kept working on Urban Informatics, which
resulted in poster~\autocite{WWWDemo16} and a journal
paper~\autocite{GeotopicTBD16}.

Furthermore I (loosely) maintained a list of references on
\href{run:support/dist.pdf}{distributed graph processing frameworks} and
\href{run:support/embedding.pdf}{neural network based graph embedding methods}.

\begingroup
% \setstretch{0.9}
\setlength\bibitemsep{2pt}
\printbibliography
\endgroup
\end{document}