### 2017-06-14 gtx_experiments.tex
add a previous document on galaxy tree experiments, plus small layout change

In this \namecref{sub:gtx_empirical_evaluation}, we provide empirical evidences of the properties of
\gtx{} over several classes of graph, and compare it with a \bfs{} baseline.\marginpars{If time
allows, it would be interesting to implement some methods of \vref{sub:gtx_state_of_the_art} and add
them to the comparison}\todo*{implement more low stretch methods} Namely, we consider three kinds of
graph topology (with both synthetic and real world instances that carry signs on their edges) and
evaluate $(i)$ what average stretch is reached by various trees and $(ii)$ how accurate is the sign
prediction.

\subsubsection{Graph topology}

The three kinds of topology we consider are:
\begin{description}
	\item[\grid{}] which are 2D lattices, where each node has four neighbors except on the boundary.
		The synthetic ones are square, while the \enquote{real world} ones represents the four neighbors
		pixel connectivity of the pictures showed in \autoref{fig:gtx_xp_bwpics}.
	\item[\lpa{}] which are built synthetically according to the model of \textcite{Barabasi1999}.
		While this does not follow the more rigorous specification of \textcite{PAmodel04}, informally,
		we start with a line graph of $m$ nodes and add node one by one until the graph consists of $n$
		nodes. Each time a new node is added, it is connected to $m$ of the existing nodes with a
		probability proportional to their degree. Here we choose $m=3.13$, that is when adding a new
		node, we pick $3$ or $4$ existing neighbors such the initial expected number of neighbors for
		each new nodes is $3.13$. Such graphs are quite sparse and have short diameter, thus providing a
		crude but reasonable approximation of online social networks. Therefore, the real world
		instances of the \lpa{} model are \wik{}, \sla{} and \epi{}\marginpars{as used in the first
		chapter}\todo*{add a ref to first chapter} along with \gplus{}. The last one is constructed from
		ego networks of \gplus{}\footnote{Available at
		\url{http://snap.stanford.edu/data/egonets-Gplus.html}} by keeping the largest connected
		component of users whose gender is known. Basic statistics of those real \lpa{} graphs are
		presented in (\autoref{tab:gtx_xp_dataset}). 
	\item[\triangle{}] which consists of a Delaunay triangulation of random 2D points\footnote{As
		implemented by the \textsf{graph-tool} library (\url{https://graph-tool.skewed.de})}.
\end{description}

\begin{table}[htpb]
	\centering
	\caption{Dataset description }\label{tab:gtx_xp_dataset}
	\begin{tabular}{lrrcc}
		\toprule
             & $|V|$  & $|E|$    & fraction of $+$ edges & $\frac{2|E|}{|V|\cdot(|V|-1)}$ \\
		\midrule
		\wik{}   & \np{7065}   & \np{99936}    & 78.5\%                & $4.00\cdot 10^{3}$             \\
		\gplus{} & \np{74917}  & \np{10130461} & 67.6\%                & $3.61\cdot 10^{3}$             \\
		\sla{}   & \np{82052}  & \np{498527}   & 76.4\%                & $1.48\cdot 10^{4}$             \\
		\epi{}   & \np{119070} & \np{701569}   & 83.2\%                & $9.90\cdot 10^{5}$             \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[t]
	\centering
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zmonastery}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zworld}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_poster}
	\end{subfigure}

	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zmonastery_bin}
		\caption{monastery}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/zworld_bin}
		\caption{world}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_poster_bin}
		\caption{poster}
	\end{subfigure}

	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_logo}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/space}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/waterfall}
	\end{subfigure}

	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/nips_logo_bin}
		\caption{logo}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/space_bin}
		\caption{space}
	\end{subfigure}~
	\begin{subfigure}[b]{0.32\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/waterfall_bin}
		\caption{waterfall}
	\end{subfigure}
	\caption{Real world pictures and their binarized version}\label{fig:gtx_xp_bwpics}
\end{figure}

\subsubsection{Stretch}

The first property of Galaxy trees we wish to evaluate is their stretch, which depends only of graph
topology. Namely, let $G$ be a graph over vertex set $V$ with $|V|=n$ and edge set
$E$.\Todo[MOVE]{Stretch definition is likely to happen somewhere earlier} Furthermore, let $T$ be a
spanning tree of $G$ and $\etest{}$ the edges of $G$ not in $T$. Then we define the \emph{average
test edge stretch} as $\frac{1}{|\etest{}|} \sum_{(u,v) \in \etest{}} |\mathrm{path}^T_{u,v}|$,
where $|\mathrm{path}^T_{u,v}|$ is the unique path between $u$ and $v$ in $T$.

As we consider unweighted graphs, we compare \gtx{} with a natural baseline, namely a spanning tree
rooted at the highest degree node and obtained through a breadth first visit of the graph. This
involves randomness in order in which nodes are visited. Likewise in \gtx{}, the choice of the edge
linking two stars is not always unique, meaning that we have to break ties at random.  Therefore,
for each graph, we repeat the tree construction 12 times and present the average result, noting that
the variance (showed as error bar in \autoref{fig:gtx_xp_st}) is small.

On \lpa{} and \triangle{}, we see that both trees exhibits logarithmic stretch, although with a
larger constant for \gtx{}. Note that this is also the case for others low stretch tree methods
\autocite[\S 5.3.1]{papplow}. On \grid{} however, \gtx{} preserves this logarithmic stretch growth
while this is visually no longer the case for \bfs{}.

\begin{figure}[tbh]
	\centering
	\begin{subfigure}[b]{0.9\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/gridst}
		\caption{\grid{} }\label{fig:gtx_xp_gridst}
	\end{subfigure}

	\begin{subfigure}[b]{0.9\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/past}
		\caption{\lpa{} }\label{fig:gtx_xp_past}
	\end{subfigure}

	\begin{subfigure}[b]{0.9\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/trst}
		\caption{\triangle{} }\label{fig:gtx_xp_trst}
	\end{subfigure}
	\caption{Stretch over graphs of increasing size}\label{fig:gtx_xp_st}
\end{figure}

\subsubsection{Sign prediction}

The second design goal of Galaxy trees is to accurately predict the sign of edges in $\etest{}$.
Except for the three real datasets that already include signs\footnote{We nonetheless perform some
preprocessing in order to make them undirected to remove the small proportion of conflicting edges
(e.g. positive from $u$ to $v$ but negative from $v$ to $u$).}, all the other are constructed,
meaning we have to set sign on their edges in the first place. This is done by partitioning the
nodes into two clusters. For \gplus{} we use node gender, for pictures we use node color (black or
white), and for all others, we propagate labels $0$ and $1$ from randomly selected high degree nodes.
Once each node belongs to one of the two clusters, we set the sign of an edge between two nodes to
be $+$ if they are in the same cluster and $-$ otherwise.  Predicting using path parity will thus
gives perfect result. To test performance in real or adversarial situation, we then add noise, that
is we select a fraction of edges uniformly at random and flip their sign.

We evaluate the performance of our prediction using the Matthews Correlation\Todo{merge this
definition of MCC with the one in first chapter} Coefficient (MCC)~\autocite{MCC00} \[
	\mathrm{MCC} = \frac{ TP \times TN - FP \times FN } {\sqrt{ (TP + FP) ( TP + FN ) (
			TN + FP ) ( TN + FN ) } } = \pm \sqrt{\frac{\chi^2}{n}}
\]
Since we do not have confidence score, we cannot use AUC. Yet we have to account for the large sign
unbalance and thus cannot rely on accuracy or $F_1$ measure.  Therefore we choose MCC, which
combines all the four numbers of the confusion matrix in a single metric. It ranges from $+1$
(perfect prediction) to $-1$ (inverse prediction) through $0$ (random prediction). As a
demonstration of MCC usefulness, predicting all edges but one to be positive on Slashdot gives
$.764$ accuracy, $.886$ $F_1$ score\marginpars{Actually the $F_1$ score is $.866$ for positive
edges, $0$ for negative ones and $.661$ if we can take an average weighted by class size.} but
$-0.0007$ MCC.

As showed in \autoref{fig:gtx_xp_mcc}, when the noise level is low, \gtx{} performs better than
\bfs{}. As the noise level gets higher, they have similar performance. Note also than in
\autoref{fig:gtx_xp_pasynthmcc}, \gtx{} is less sensible to the size of the graph.

\begin{figure}[tbh]
	\centering
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grsynthmcc}
		\caption{Synthetic \grid{} }\label{fig:gtx_xp_grsynthmcc}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grrwmcc}
		\caption{Pictures \grid{} }\label{fig:gtx_xp_grrwmcc}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/pasynthmcc}
		\caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthmcc}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/trmcc}
		\caption{\triangle{} }\label{fig:gtx_xp_trmcc}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/parwmcc}
		\caption{Real world network }\label{fig:gtx_xp_parwmcc}
	\end{subfigure}
	\caption{MCC over various graphs}\label{fig:gtx_xp_mcc}
\end{figure}

To further assess the quality of our trees, we plug them in them into a successful heuristic method
to predict edge sign: \asym{}~\autocite{Kunegis2009}. \Todo{It might also be interesting to see if
that would be a good training set for our troll method, although it has to be checked it makes sense
from a running time point of view.} It computes the exponential of the adjacency matrix after it has
been reduce to $z$ dimension. This allows to count the sign of all paths between two pairs of nodes
with decreasing weight depending of their length. To simulate an active learning setting, we reveal
only a subset of edge in $A$. This subset can be: $i)$ the edges forming a \bfs{}, $ii)$ the edges
forming a \gtx{} $iii)$ $|V|-1$ edges chosen uniformly at random.

We set the parameter $z$ equal to $15$ because $i)$ it is one of the best in \cite[Fig.
11]{Kunegis2009}, $ii)$ it performs well on real dataset in \cite[Fig.3]{Cesa-Bianchi2012a}, and
$iii)$ it was good in our initial testing (\texttt{20150401\_wed\_spectral.ipynb}).

As the \asym{} has a $O(n^3)$ complexity and uses quite some memory at prediction time, the larger
graphs used previously are not all included. The conclusion of \autoref{fig:gtx_xp_asym} is that
except on social network, it is better to use spanning trees than random edges. Specifically, \gtx{}
on \grid{} and \bfs{} elsewhere.

\begin{figure}[tbh]
	\centering
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grsynthasym}
		\caption{Synthetic \grid{} \label{fig:gtx_xp_grsynthasym}}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/grrwasym}
		\caption{\enquote{Real} \grid{} }\label{fig:gtx_xp_grrwasym}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/pasynthasym}
		\caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthasym}
	\end{subfigure}~
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/trasym}
		\caption{\triangle{} }\label{fig:gtx_xp_trasym}
	\end{subfigure}
	\begin{subfigure}[b]{0.47\textwidth}
		\includegraphics[width=\textwidth]{gtx_exp/parwasym}
		\caption{Real world network }\label{fig:gtx_xp_parwasym}
	\end{subfigure}
	\caption{\asym{} over various graphs}\label{fig:gtx_xp_asym}
\end{figure}

Finally\marginpars{Actually I never did it because \shz{} wasn't implemented at the time, so now is
a good occasion}\todo*{Run shazoo on galaxy tree} we also compare \gtx{} with \bfs{} and \rst{} on
the task of nodes prediction using \shz{} algorithm~\autocite{Vitale2012}.