### 2017-06-14 lowstretch_and_cc.tex

\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, final,
svgnames, marginals=raggedouter, english,
%symmetric, justified,
]{article}
\usepackage{charter}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\input{packages}
\input{abbreviations}

\title{Edge sign prediction in general graphs and connection with Correlation Clustering}

\begin{document}
\newgeometry{vmargin=1.5cm,left=1.5cm,textwidth=130mm,marginparsep=5mm,marginparwidth=52mm}
\maketitle
% \tableofcontents

\section{Limitations of the troll method}
\label{sec:limitations_of_the_troll_method}

\subsection{Reliance on the sign generative model}
\label{sub:reliance_on_the_sign_generative_model}

The method presented in the previous chapter hinges crucially upon our sign generative model. Yet
one can imagine contexts where this model is not applicable, especially when nodes do not represent
human beings. One way to alleviate this issue was introduced with our online algorithm, where this
time, signs are generated by an arbitrary adversary. However, in that case, we are still facing two
limitations\marginpars{limitations is not the right word, maybe \emph{strong constraints}…}
\begin{enumerate}
	\item Our bias remains that the labeling is regular (recall this means informally than the all
		the outgoing signs from a given node tend to be the same, and likewise for the incoming signs),
		since irregularities are the cost payed by the adversary to make our algorithm mispredict. While
		this bias is well suited to social networks, other applications may require other
		bias\marginpars{namely the \pcc{} bias we'll introduce next}.
	\item In the online setting, we evaluate our performance by the regret\marginpars{ref to regret equation},
		whereas in general we are interested in more classical measures of predictive accuracy.
\end{enumerate}

\subsubsection{Experience on directed biological network}
\label{ssub:experience_on_directed_biological_network}

Take for instance a biological network.

According to \href{https://web.stanford.edu/class/cs224w/slides/handout-bionets.pdf}%
{this description of various biological networks}, \emph{gene regulatory network} are directed and
contains activation and inhibition links, as \href{https://en.wikipedia.org/wiki/Gene_regulatory_network#Overview}%
{showed on Wikipedia}. There are some online databases such as
\href{http://regulondb.ccg.unam.mx/menu/download/datasets/index.jsp}%
{RegulonDB} or \href{http://www.pathwaycommons.org/pcviz/}{Pathway Commons} (this last one provide
network in the BioPAX format, which can be visualized by
\href{http://www.cytoscape.org/}{CytoScape}).
Another source is the \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2708159/table/T1}{Table 1}
of \cite{BioSigned09}.
%DasGupta also describe (in
%http://www.sciencedirect.com/science/article/pii/S0303264706001419#sec14) how
%they built a directed signed network from this SBML file
%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1681468/bin/msb4100014-sd1.xml

other domain? coref, images (is it directed?), entity resolution…\marginpars{Actually it's difficult
to find other kind of graph because all constructed ones (like image, coreference,
deduplication) are inherently symmetric}

\subsection{Need for a directed graph}
\label{sub:need_for_a_directed_graph}

A more serious limitation is that our sign model only applies to directed graphs. Consider
another model, where each node $i$ is endowed with an integer $c_i \in \{1, \ldots, k\}$ that
specifies its cluster and let the sign of the undirected edge $i,j$ be $+$ if $c_i = c_j$ and $-$
otherwise.  This corresponds naturally to the situation modeled by the \pcc{} problem.

\section{\pcc{}}
\label{sec:correlation_clustering}

\subsection{Problem setting}
\label{sub:problem_setting}

\subsection{Relation with edge sign prediction}
\label{sub:relation_with_edge_sign_prediction}

\iffalse
For the 2 clusters case ($k=2$), characterization proven already in 1936 by \textcite{Konig36}, as noted in
\cite{Zaslavsky2012} (commenting on \cite{harary1953}: \enquote{Although Thm. 3 was anticipated by
\textcite[Theorem X.11]{Konig36}  without the terminology of signs, here is the
first recognition of the crucial fact that labelling edges by elements of a
group—specifically, the sign group—can lead to a general theory.})
and \cite{Huffner2010} (\enquote{\textcite{Konig36} proved the following characterization of
	balanced graphs. For a graph $G = (V , E)$, the following are equivalent:\marginpars{There is a
	proof in \autocite[p. 111]{BookKleinberg2010}, maybe I can rewrite it as well}
	\begin{enumerate}
		\item $V$ can be partitioned into two sets $V_1$ and $V_2$ called sides such that there is no
			negative edge $\{v, w\} \in E$ with both $v, w \in V_1$ or both $v, w \in V_2$ and no negative
			edge $\{v, w\}$ with $v \in V_1$ and $w \in V_2$ .
		\item $V$ can be colored with two colors such that for all $\{v, w\} \in E^-$, the vertices $v$
			and $w$ have different colors, and for all $\{v, w\} \in E^+$, the vertices $v$ and $w$ have
			the same color. The color classes correspond to the sides.
		\item $G$ does not contain cycles with an odd number of negative edges.
	\end{enumerate}
	Using the characterization by a coloring, it is easy to see that balance of a signed
graph can be checked in linear time by depth-first search.})
\fi

\subsection{State of the art}
\label{sub:state_of_the_art}

\subsection{Variants and extensions}
\label{sub:variants_and_extensions}

\subsubsection{\pcc{} under stability assumption}
\label{ssub:cc_under_stability_assumption}

\iffalse
Haris Angelidakis, Konstantin Makarychev, and Yury Makarychev. 2017.
Algorithms for Stable and Perturbation-Resilient Problems. STOC’17
\href{http://ttic.uchicago.edu/~yury/papers/two-stable.pdf}{10.1145/3055399.3055487}
improves over the one cited in the internship description
\fi

\subsubsection{Parallel \pcc{}}
\label{ssub:parallel_cc}

\subsection{Empirical evaluation?}
\label{sub:cc_empiracal_evaluation}

\section{Low stretch trees and spanners}
\label{sec:low_stretch_trees_and_spanners}

\subsection{\gtx{}: a simple low stretch tree construction}
\label{sub:gtx_a_simple_low_stretch_tree_construction}

\subsection{State of the art}
\label{sub:gtx_state_of_the_art}

\iffalse
\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
n}{n}\right)\right)$ edges in $O(|E|)$ time.

While they have many applications [see first paragraph of \url{https://arxiv.org/pdf/1401.2454.pdf},
which was later merged in a STOC'14 paper] (a major one being solving linear system), in some
practical situations their advantages are less clear [from
\url{https://link.springer.com/chapter/10.1007/978-3-319-20086-6_16}\enquote{for reasonable inputs
the constant factors make the solver much slower than methods with higher asymptotic complexity.
One other aspect predicted by theory is confirmed by our findings: Spanning trees with lower
stretch indeed reduce the solver's running time. Yet, simple spanning tree algorithms perform
better in practice than those with a guaranteed low stretch.} this is improved by
\url{https://link.springer.com/chapter/10.1007%2F978-3-319-20086-6_17} although they seem to work
	mostly with the Laplacian of the tree ]
\fi

\subsection{Empirical evaluation}
\label{sub:gtx_empirical_evaluation}

\input{gtx_experiments}

\section{Conclusion}

\begingroup
\small
\setstretch{0.9}
\todos

\setlength\bibitemsep{2pt}
\printbibliography
\endgroup

% \bibliography{/home/orphee/data/projects/biblio/library.bib,more.bib}
% \bibliographystyle{plainnat}
\end{document}