### 2017-06-21 assets/tikz/gtx_run_level1_tikz.tex

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=8mm}]
  \node[vertex,fill=DodgerBlue,text=black] (s1) at (0.000, -0.000) {$S_1^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s2) at (0.845, -1.724) {$S_2^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s3) at (1.867, 0.448) {$S_3^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s4) at (2.122, -4.332) {$S_4^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s5) at (2.967, -6.056) {$S_5^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s6) at (3.989, -3.884) {$S_6^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s7) at (4.691, 1.126) {$S_7^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s8) at (5.535, -0.598) {$S_8^1$};
  \node[vertex,fill=DodgerBlue,text=black] (s9) at (6.558, 1.574) {$S_9^1$};

  \foreach \source/\dest in {1/2, 1/3, 2/8,3/6,4/5,4/6,7/8,7/9}
  \draw[edge] (s\source) -- (s\dest);
  \draw[edge] (s5) edge [bend left =-17]  (s9);
\end{tikzpicture}
\end{document}