### 2017-06-25 notations.tex

\begin{table*}[htpb]
  \centering
  \caption{List of notations used in this thesis}\label{tab:notations}%
  \vspace{-.5\baselineskip}
  \begin{tabulary}{187mm}{LL}
    \toprule
    Symbol & Meaning \\
    \midrule
    $G$ & A graph, either directed or undirected \\
    $V$, $V_t$ & The set of all the nodes of a graph. Unless noted otherwise, nodes are indexed from $1$ to $|V|$ \\
    $E$, $E_t$ & The set of all the edges of a graph \\
    $\degr(u)$ & The total degree of node $u$ (that is the number of edges incident to $u$
    regardless of their direction) \\
    $(u,v)$, $(i,j)$ & An undirected edges between nodes $u$ and $v$ (or $i$ and $j$) \\
    $\mathcal{N}(u)$ & The set of all neighbors of $u$, regardless of edge direction. We thus have
    $|\mathcal{N}(u)| = \degr(u)$ \\
    \bottomrule
  \end{tabulary}
\end{table*}