### 2017-06-27 assets/tikz/gtx_eccentricity_tikz.tex
change the example graph to add color on its edges

\documentclass[svgnames,11pt,tikz]{standalone}
\input{preambule}
\begin{document}
\tikzstyle{star}=[DodgerBlue,line width=2pt]
\tikzstyle{star-label}=[font=\Large]
\tikzstyle{removed}=[dashed,line width=1.7pt]
\begin{tikzpicture}[auto,vertex/.append style={minimum width=8mm},
  edge/.append style={line width=2.2pt,color=DodgerBlue!90!Black}]
  \draw[star,Orange] (2.260, -1.063) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$\mathbf{S_1^2}$};
  \draw[star,Orange] (6.504, -9.728) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$\mathbf{S_2^2}$};
  \draw[star,Orange] (11.641, 1.189) circle [radius=4.4]  node[star-label,font=\LARGE,xshift=-10, yshift=110] {$\mathbf{S_3^2}$};

  \draw[star] (0.000, -0.000) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$\mathbf{S_1^1}$};
  \draw[star] (1.689, -3.448) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$\mathbf{S_2^1}$};
  \draw[star] (3.734, 0.896) circle [radius=1.8]  node[star-label,xshift=-30, yshift=5] {$\mathbf{S_3^1}$};
  \draw[star] (4.245, -8.664) circle [radius=1.8]  node[star-label,xshift=-22, yshift=-25] {$\mathbf{S_4^1}$};
  \draw[star] (5.934, -12.113) circle [radius=1.8]  node[star-label,xshift=-15, yshift=35] {$\mathbf{S_5^1}$};
  \draw[star] (7.978, -7.768) circle [radius=1.8]  node[star-label,xshift=30, yshift=25] {$\mathbf{S_6^1}$};
  \draw[star] (9.381, 2.252) circle [radius=1.8]  node[star-label,xshift=-15, yshift=30] {$\mathbf{S_7^1}$};
  \draw[star] (11.071, -1.196) circle [radius=1.8]  node[star-label,xshift=-30, yshift=25] {$\mathbf{S_8^1}$};
  \draw[star] (13.115, 3.149) circle [radius=1.8]  node[star-label,xshift=-30, yshift=20	] {$\mathbf{S_9^1}$};

  \node[vertex,fill=black!0,] (n1) at (0.000, -0.000) {1};
  \node[vertex,fill=black!25,] (n2) at (0.223, -1.179) {2};
  \node[vertex,fill=black!25,] (n3) at (-1.179, -0.223) {3};
  \node[vertex,fill=black!25,] (n4) at (-0.223, 1.179) {4};
  \node[vertex,fill=black!25,] (n5) at (1.179, 0.223) {5};
  \node[vertex,fill=black!25,] (n6) at (1.689, -3.448) {6};
  \node[vertex,fill=black!50,] (n7) at (2.408, -4.409) {7};
  \node[vertex,fill=black!50,] (n8) at (0.728, -4.167) {8};
  \node[vertex,fill=black!50,] (n9) at (0.971, -2.487) {9};
  \node[vertex,fill=black!50,] (n10) at (2.650, -2.730) {10};
  \node[vertex,fill=black!25,] (n11) at (3.734, 0.896) {11};
  \node[vertex,fill=black!50,] (n12) at (4.391, -0.108) {12};
  \node[vertex,fill=black!50,] (n13) at (2.730, 0.240) {13};
  \node[vertex,fill=black!50,] (n14) at (3.077, 1.901) {14};
  \node[vertex,fill=black!50,] (n15) at (4.738, 1.553) {15};
  \node[vertex,fill=black!25,] (n16) at (4.245, -8.664) {16};
  \node[vertex,fill=black!50,] (n17) at (4.450, -9.846) {17};
  \node[vertex,fill=black!50,] (n18) at (3.062, -8.870) {18};
  \node[vertex,fill=black!50,] (n19) at (4.039, -7.482) {19};
  \node[vertex,fill=black!50,] (n20) at (5.427, -8.459) {20};
  \node[vertex,fill=black!50,] (n21) at (5.934, -12.113) {21};
  \node[vertex,fill=black!75,text=white] (n22) at (5.614, -13.269) {22};
  \node[vertex,fill=black!75,text=white] (n23) at (4.777, -11.793) {23};
  \node[vertex,fill=black!75,text=white] (n24) at (6.254, -10.956) {24};
  \node[vertex,fill=black!75,text=white] (n25) at (7.090, -12.433) {25};
  \node[vertex,fill=black!50,] (n26) at (7.978, -7.768) {26};
  \node[vertex,fill=black!75,text=white] (n27) at (8.018, -8.967) {27};
  \node[vertex,fill=black!75,text=white] (n28) at (6.779, -7.808) {28};
  \node[vertex,fill=black!75,text=white] (n29) at (7.939, -6.568) {29};
  \node[vertex,fill=black!75,text=white] (n30) at (9.178, -7.728) {30};
  \node[vertex,fill=black!25,] (n31) at (9.381, 2.252) {31};
  \node[vertex,fill=black!50,] (n32) at (8.964, 1.127) {32};
  \node[vertex,fill=black!50,] (n33) at (8.256, 2.669) {33};
  \node[vertex,fill=black!50,] (n34) at (9.799, 3.377) {34};
  \node[vertex,fill=black!50,] (n35) at (10.507, 1.835) {35};
  \node[vertex,fill=black!50,] (n36) at (11.071, -1.196) {36};
  \node[vertex,fill=black!75,text=white] (n37) at (11.078, -2.396) {37};
  \node[vertex,fill=black!75,text=white] (n38) at (9.871, -1.203) {38};
  \node[vertex,fill=black!75,text=white] (n39) at (11.064, 0.004) {39};
  \node[vertex,fill=black!75,text=white] (n40) at (12.271, -1.189) {40};
  \node[vertex,fill=black!50,] (n41) at (13.115, 3.149) {41};
  \node[vertex,fill=black!75,text=white] (n42) at (13.437, 1.993) {42};
  \node[vertex,fill=black!75,text=white] (n43) at (11.959, 2.827) {43};
  \node[vertex,fill=black!75,text=white] (n44) at (12.794, 4.305) {44};
  \node[vertex,fill=black!75,text=white] (n45) at (14.271, 3.470) {45};

  \draw[edge] (n1) edge [bend left =0] node[] {} (n2);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n3);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n4);
  \draw[edge] (n1) edge [bend left =0] node[] {} (n5);
  \draw[edge,Orange!90!Black] (n2) edge [bend left =25] node[] {} (n6);
  \draw[edge,removed,Orange!90!Black] (n2) edge [bend left =0] node[] {} (n9);
  \draw[edge,removed,Orange!90!Black] (n3) edge [bend left =-20] node[] {} (n12);
  \draw[edge,Orange!90!Black] (n4) edge [bend left =20] node[] {} (n14);
  % \draw[edge] (n3) edge [bend left =-20] node[] {} (n18);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n7);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n8);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n9);
  \draw[edge] (n6) edge [bend left =0] node[] {} (n10);
  % \draw[edge] (n7) edge [bend left =-15] node[] {} (n43);
  % \draw[edge] (n9) edge [bend left =-10] node[] {} (n32);
  \draw[edge,Green!90!Black] (n8) .. controls ++(-40:3) and ++(240: 2) .. (n37);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n12);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n13);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n14);
  \draw[edge] (n11) edge [bend left =0] node[] {} (n15);
  \draw[edge,removed,Green!90!Black] (n13) edge [bend left =0] node[] {} (n29);
  % \draw[edge] (n12) edge [bend left =0] node[] {} (n18);
  \draw[edge,Green!90!Black] (n15) .. controls ++(-40:3) and ++(64: 3.5) .. (n26);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n17);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n18);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n19);
  \draw[edge] (n16) edge [bend left =0] node[] {} (n20);
  \draw[edge,Orange!90!Black] (n17) edge [bend left =0] node[] {} (n23);
  %\draw[edge] (n19) edge [bend left =0] node[] {} (n27);
  \draw[edge,Orange!90!Black] (n19) edge [bend left =0] node[] {} (n29);
  \draw[edge,removed,Orange!90!Black] (n20) edge [bend left =0] node[] {} (n24);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n22);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n23);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n24);
  \draw[edge] (n21) edge [bend left =0] node[] {} (n25);
  \draw[edge,removed,Green!90!Black] (n25) edge [bend left =-20] node[] {} (n42);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n27);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n28);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n29);
  \draw[edge] (n26) edge [bend left =0] node[] {} (n30);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n32);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n33);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n34);
  \draw[edge] (n31) edge [bend left =0] node[] {} (n35);
  \draw[edge,Orange!90!Black] (n32) edge [bend left =0] node[] {} (n39);
  \draw[edge,Orange!90!Black] (n35) edge [bend left =0] node[] {} (n43);
  \draw[edge,removed,Orange!90!Black] (n33) edge [bend left =-25] node[] {} (n38);
  \draw[edge,removed,Orange!90!Black] (n34) edge [bend left =20] node[] {} (n43);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n37);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n38);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n39);
  \draw[edge] (n36) edge [bend left =0] node[] {} (n40);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n42);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n43);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n44);
  \draw[edge] (n41) edge [bend left =0] node[] {} (n45);

  \draw[star, Green!90!Black] (11.8, 6.5) 
  .. controls ++(190:1) and ++(55: 2) .. (-2, 2.5)
  .. controls  ++(55: -4) and ++(170:4) .. (6, -15) 
  .. controls  ++(-10:9) and ++(10:9) .. node[pos=.4,font=\Huge,xshift=-15pt] {$\mathbf{S_1^3}$} cycle ;
\end{tikzpicture}
\end{document}