### 2017-06-27 assets/tikz/gtx_singleton_tikz.tex

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
  \tikzstyle{star}=[DodgerBlue,line width=1pt]
  \tikzstyle{star-label}=[font=\Large]
  \begin{tikzpicture}[auto,vertex/.append style={minimum width=6.5mm}]
    \node[vertex,] (c1) at (-2.427,-0.000) {$c_1$};
    \node[vertex,] (c2) at (2.427,0.000) {$c_2$};
    \node[vertex,] (c3) at (0.000,0.000) {$c_3$};

    \draw[star] (0.000, -0.000) circle [radius=.4]  node[star-label,xshift=0, yshift=25] {$S_3^1$};
    \draw[star] (c1) circle [radius=2]  node[star-label,xshift=-30, yshift=30] {$S_1^1$};
    \draw[star] ( c2) circle [radius=2]  node[star-label,xshift=-30, yshift=30] {$S_2^1$};

    \node[vertex,medium] (p1) at (-1.283,-1.144) {$p_1$};
    \draw[edge,] (c1) edge [bend left =0] node[] {} (p1);
    \node[vertex,medium] (p2) at (-2.008,1.563) {$p_2$};
    \draw[edge,] (c1) edge [bend left =0] node[] {} (p2);
    \node[vertex,medium] (p3) at (-3.990,-0.419) {$p_3$};
    \draw[edge,] (c1) edge [bend left =0] node[] {} (p3);
    \node[vertex,medium] (p4) at (1.283,-1.144) {$p_4$};
    \draw[edge,] (c2) edge [bend left =0] node[] {} (p4);
    \node[vertex,medium] (p5) at (3.990,-0.419) {$p_5$};
    \draw[edge,] (c2) edge [bend left =0] node[] {} (p5);
    \node[vertex,medium] (p6) at (2.008,1.563) {$p_6$};
    \draw[edge,] (c2) edge [bend left =0] node[] {} (p6);
    \draw[edge,] (c3) edge [bend left =0] node[] {} (p4);
    \draw[edge,] (c3) edge [bend left =0] node[] {} (p1);

    \draw[->] (0,-2) -- (0,-2.9);

    \node[vertex,fill=DodgerBlue,text=black] (s1) at (-2,-3.7) {$S_1^1$};
    \node[vertex,fill=DodgerBlue,text=black] (s2) at (2,-3.7) {$S_2^1$};
    \node[vertex,fill=DodgerBlue,text=black] (s3) at (0,-3.7) {$S_3^1$};
    \foreach \source/\dest in {3/2, 1/3}
    \draw[edge] (s\source) -- (s\dest);
  \end{tikzpicture}
\end{document}