### 2017-07-10 gtx_related.tex
add more related works on low stretch tree

Looking for a subgraph $H$ of $G$ that best preserve the distance in $G$ while being sparse is an old
problem, especially driven by network design in fields such as transportation~\autocite{RoadNetworks60}
and electrical circuits~\autocite{electricalNetworks60}. For instance, \textcite{Johnson1978} define
the following problem\footnote{We adapt their notations to match ours}:  
\begin{problem}[Network Design Problem]
  \label{prob:gtx_ndp}
  Given an undirected integer-weighted graph $G=(V, E, w)$, a budget $B\in\Nbb$ and a criterion
  threshold $C\in \Nbb$, does there exist a spanning subgraph $G'=(V, E')$ of $G$ with weight
  $w(E') \leq B$ and criterion value $F(G') \leq C$, where the criterion function $F(G')$ denotes
  the sum of the weights of the shortest paths in $G'$ between all vertex pairs?
\end{problem}
Note that their definition of stretch includes all possible edges in $G$ and not only those in $E$.
They prove that this problem is \NPc{} by exhibiting a reduction from the \textsc{Knapsack}
problem. Finally they prove that the simpler problem of finding a spanning tree on an unweighted
graph, that is
\vspace{-.5\baselineskip}
\begin{problem}[Simple Network Design Problem]
  \autoref{prob:gtx_ndp} with $w$ being the equal to $1$ for all edges in $E$ and $B=|V|-1$.
\end{problem}%
\vspace{-.5\baselineskip}
\noindent is also \NPc{} by reduction from \textsc{Exact 3-Cover}.
In this setting of a spanning tree minimizing the stretch over all pairs of nodes, \enquote{
\textcite{AllPairStrech08} obtained a $(c log n)^{O(\log \Delta)}$ factor, where $c$ is the optimal
maximum stretch and $\Delta$ is the diameter.  They also showed $O(1)$ approximation for the case
where the graph is unweighted. The constant was later reduced to $6$ by \textcite{AllPairStrech10}}

\begin{problem}[Optimal Communication Spanning Tree~\autocite{Requirements74}]
Given a set of nodes $V=\{v_1, \ldots, v_n\}$, a set of distances $d_{ij}$ and a set of requirements
$r_{ij}$ between $v_i$ and $v_j$, find a spanning tree connecting these $n$ nodes such that the
total cost of communication of the spanning tree is a minimum among all spanning trees. The cost of
communication for a pair of nodes is $r_{i,j}$ multiplied by the sum of the distances of arcs which
form the unique path connecting $v_i$ and $v_j$ in the spanning tree. The cost of a spanning tree is
the sum of costs over all pairs of nodes.
\end{problem}

Given a weighted graph $G=(V,E,w)$, if we let $d_{ij}$ be $w((i,j))$ and $r_{ij}$ be $1$ if
$(i,j)\in E$ and $0$ otherwise, it seems to me this correspond to the weighted low stretch problem.
% However, \textcite[Section 3, page 453]{lognMetricBoundConf03} claim a $O(\log n)$ approximation so maybe I'm
% wrong. Turns out, they refer to a distribution over trees and this $O(\log n)$ is the expected
% stretch of a tree sampled from this distribution

This defines two kind of structures, spanning trees and spanners (which are still sparse subgraphs
yet containing more than $|V|-1$ edges).

\paragraph{Trees}
\label{par:trees}

Earlier but slightly less related, \textcite{OptimalNetwork69} studies the following problem:
\begin{problem}[Optimal Network Problem]
  \label{prob:gtx_scott}
  Given a set $V$ of $n$ vertices, find a set of spanning edges $E\subset V^2$ that minimizes
  the sum of the length of the shortest paths  between all vertex pairs while the
  total length of the resulting network does not exceed some upper bound $B\in\Nbb$.
\end{problem}
This can be seen as a special case of \autoref{prob:gtx_ndp} with $G$ being the unweighted
\marginpars{according to the experiments, it's not clear whether there are weights or
not…} $n$-complete graph. He proposes a backtracking solution and two local search approximate
algorithms. Some early branch and bound heuristic solutions to \autoref{prob:gtx_scott} are surveyed
in~\autocite[Section 2.3.2]{networkDesignSurvey89} although they do not come with asymptotic
guarantee on the stretch. Furthermore, \textcite{optimApproxNP80} proves that for any $\epsilon \in
(0,1)$, finding a $|V|^{1-\epsilon}$ approximation is \NPc{}.

Next is the seminal paper of \textcite{LowerBound95}. It touches on many topics, and frame the
problem in a game theoretic way but here we only focus on two of their results: a lower bound of
$\Omega(\log n)$ for the average stretch of any tree and their construction of a tree with $\exp
O(\sqrt{\log n\log\log n})$ average stretch in time $O(m^2)$. The lower bound follows from an
existing result in extremal graph theory~\autocite[pages 107--109]{ExtremalGraph04}: there is a
positive constant $a$ such that for all $n\in \Nbb$, one can construct a graph $G$ with $n$ vertices
and $2n$ edges such that every cycle $G$ has a length of at least $a\log n$. Now consider any
spanning tree $T$ of $G$.  While all the $n-1$ edges of $T$ have a stretch of $1$, the $n+1$
remaining ones form a cycle in $T$ hence in $G$ as well and thus incur a stretch of at least $a\log
n$. This shows that the average stretch is at least $\frac{1}{2}a\log n$.

They construct a low stretch spanning tree in a bottom up manner like the \gtx{} algorithm. First,
they extend the definition of stretch to multigraph~\autocite[Section 4]{LowerBound95} and then
describe a procedure to transform in linear time any multigraph $G$ with $n$ nodes to a multigraph
$G'$ on the same nodeset with at most $n(n+1)$ edges such the average stretch of $G'$ is at most
twice that of $G$~\autocite[Lemma 5.2]{LowerBound95}. The next ingredient is an algorithm to build a
low diameter decomposition of a multigraph $G$, parametrized by a number $x(n)$ depending of $n$. It
works by repeatedly selecting an arbitrary node and growing a ball around it until the number of
edges leaving the ball is at most a fraction $\nicefrac{1}{x(n)}$ of the number of edges with both
endpoints in the ball. The key property of this decomposition is that it yields a partition of $G$
in clusters such that the radius of each cluster is small (namely at most $O(x(n)\log n)$) and there
is most a fraction $\nicefrac{1}{x(n)}$ of edges between clusters. Finally, the iterative procedure
is a follows: once a partition has been built, we compute a shortest path spanning tree in each
cluster that are then collapsed into super nodes to form the next graph $G'$ and the process repeats.
Another difference from \gtx{}, besides the partition procedure, is that $G'$ is a multigraph,
taking into account the number of edges joining cluster, while \collapseStar{} picks only the most
direct one.

Another interesting idea from this paper is to consider a distribution over trees instead of a
single instance, especially when one is concern about the maximum stretch instead of the average
one. For instance, on a cycle with $n$ nodes, a tree is obtained by removing one edge, and that edge
incurs a stretch of $n-1$. The uniform distribution over such trees has a maximum stretch of
$2\left(1 - \frac{1}{n}\right)$~\autocite{circle2k89}.

\begin{table}[htbp]
  \centering
  \caption{Reproduction of Table 1 from~\autocite{Abraham2012} showing the evolution of the best
  asymptotic average stretch over time.}\label{tab:gtx_related}
  \begin{tabular}{lll}
    \toprule
    work                      & average stretch                          & time                    \\
    \midrule
    \autocite{LowerBound95}   & $\exp(O(\sqrt{\log n\log\log n}))$       & $O(m^2)$                \\
    \autocite{LowerStretch05} & $O((\log n)^2 \log \log n)$              & $O(m \log^2 n)$         \\
    \autocite{nearlyTight08}  & $O(\log n(\log \log n)^3)$              & $O(m \log^2 n)$         \\
    % \autocite{nearlyTight08}  & $O(\log n \log\log n(\log\log\log n)^3)$ & $O(m^2)$                \\
    \autocite{TighterSDD11}   & $O(\log n(\log \log n)^3 )$              & $O(m \log n\log\log n)$ \\
    \autocite{Abraham2012}    & $O(\log n \log \log n)$                  & $O(m \log n\log\log n)$ \\
    \bottomrule
  \end{tabular}
\end{table}

\autocite{LowerStretch05} is the star decomposition described in section 3 of \autocite{papplow},
which was modified by \autocite{nearlyTight08} to improve the stretch. Then \textcite{TighterSDD11}
improved the runtime by rounding the edge weights to the closest power of $2$ and using a modified
implementation of the Dijkstra's algorithm in the case of at most $k$ distinct edge
weights~\autocite{FastPathFewWeights10}. Finally, \textcite{Abraham2012} describe an even more
complex but tighter petal decomposition.
\begin{marginfigure}
  \centering
  \includegraphics[width=.95\textwidth]{assets/raw/star_decomp.pdf}
  \caption{Star decomposition (reproduced from Figure 1 of~\autocite{LowerStretch05})}
  \label{fig:gtx_star_decomp}
\end{marginfigure}

If the tree is non spanning (\ie{} it can include edges not from the original graph as long as the
distances in tree remain larger the distances in the graph), the average stretch can be reduced to
$\Theta(\log n)$~\autocite{lognMetricBoundConf03}. Furthermore, when the objective if to minimize
the average stretch over all possible pairs of node, then \textcite{constantDistortion07} achieve a
universal constant bound.

Special case of graph
series parallel
\enquote{In a subsequent paper, \textcite{seriesParallel06} proved that every series-parallel
unweighted graph admits a spanning tree of average stretch $O(log n)$. This bound is tight as it
matches the lower bound established in~\autocite{cutsTrees99}.}

\paragraph{Spanners}
\label{par:spanners}

% check some surveys of the 80's
% http://pubsonline.informs.org/doi/abs/10.1287/trsc.18.1.1
% http://onlinelibrary.wiley.com/doi/10.1002/net.3230190305/full
% early solutions
% http://onlinelibrary.wiley.com/doi/10.1002/net.3230090104/full
% http://onlinelibrary.wiley.com/doi/10.1002/net.3230130309/full
% later solution?
% http://ieeexplore.ieee.org/document/81738