### 2017-08-23 cc/cc_main.tex

% *TODO*
% say more clearly there are 3 methods for approximation so far: LP & region growing; LP & distance
% pivot; combinatorial pivot
% say that in practice, general graphs are assumed to be complete with missing edges being negative
% - “The problem is agnostic in the sense that the clustering of the data is not taken into account
%   or even assumed to exist.” another way to look at it is that we're trying to get a binary
%   classifier of edges (+ or -) using an hypothesis class (nodes clustering) with limited
%   expression power (as it has to respect transitivity constraints, whereas signs can be inconsistent)
% - when the graph is weighted, there is no much point separating complete from
%   general since weights can be 0

In this section, we introduce the \pcc{} problem (\autoref{sub:problem_setting}) and motivate it
both by general applications (\autoref{ssub:cc_applications}) and by its close connection with the
\esp{} problem (\autoref{sub:relation_with_edge_sign_prediction}). After discussing its theoretical
complexity and various approximation methods (\autoref{ssub:cc_harness_approx}) along other kinds of
methods (\autoref{ssub:cc_methods}), we will review related problems stemming from the original
formulation (\autoref{sub:variants_and_extensions}). In doing so, we present a thorough and updated
counterpart to existing surveys the same
topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017}.

% surveyCC16 is about mind on complete graph (ie cluster editing). They give a Russian article from
% 1971 that says it cluster editing on triangle-free graph can be solved in polynomial time, which
% include bipartite graph (so for CC, if G^+ is (complete? I think so) bipartite)
% Fridman, G.Š.: A graph approximation problem. Upravlyaemye Sistemy 8, 73–75 (1971). (in Russian)
% In [14] it was observed that the problem is polynomially solvable for instances of at most 2
% clusters??! Actually it means consensus clustering can be solved in polynomial time when there are
% only 2 clustering as input

\subsection{Problem setting and applications}
\label{sub:problem_setting}
\input{cc_problem_applications}

\subsection{Relation with edge sign prediction}
\label{sub:relation_with_edge_sign_prediction}
\input{cc_sign_relation}

\subsection{State of the art}
\label{sub:state_of_the_art}

\subsubsection{Hardness and approximation}
\label{ssub:cc_harness_approx}
\input{cc_approx}
\subsubsection{Proposed approaches}
\label{ssub:cc_methods}
\input{cc_proposed}

\subsection{Variants and extensions}
\label{sub:variants_and_extensions}
\input{cc_variants}

\subsubsection{\pcc{} under stability assumption}
\label{ssub:cc_under_stability_assumption}
\input{cc_multimincut_stable}