### 2017-08-23 cc/conclusions.tex

\begin{itemize}
	\item improve stable \pcc{}
	\item give guarantees for \gtx{}
	\item handle weights in \gtx{}
\end{itemize}