### 2017-08-23 cc/main.tex

In the previous chapter, we presented a learning bias for the \esp{} problem, namely our sign
generative model. By leveraging its rigorous theoretical guarantees, we have demonstrated
empirically its good performance on real \dssn{}. However, in this chapter, we consider other kinds
of signed networks and show they require another bias and different, more combinatorial algorithms.
More precisely, we start by pointing in \autoref{sec:new_bias} that our previous bias may not apply
to directed graphs from domains other than social science, or to undirected graphs. Motivated by
balance theories, we then introduce a new bias, assuming that ideally, nodes are grouped in $K$
clusters such that all edges within clusters are positive and all edges between cluster are
negative. Recovering such clusters from a given signed graph has been studied extensively in the
last decades under the \pcc{} name. In \autoref{sec:correlation_clustering}, we show how this
difficult combinatorial clustering problem is connected to our learning bias. We survey a wide range
of methods to solve it exactly, approximately or heuristically, paying special attention to settings
where such recovery is easier, such as noisy or stable instances.  Finally, focusing on the
important and convenient special case $K=2$, we develop in \autoref{sec:gtx} an algorithm that,
given an undirected general graph, produce a spanning tree designed to support fast and accurate
edge sign prediction.

\input{../notations}

\section{A bias for general graphs}
\label{sec:new_bias}

\input{bias_main}

\section{\pcc{}}
\label{sec:correlation_clustering}

\input{cc_main}

\section{Low stretch trees and spanners}
\label{sec:gtx}

\input{gtx_main}

\section{Conclusions}
\label{sec:cc_conclusions}

\input{conclusions}