### 2017-08-23 lowstretch_and_cc.tex

\documentclass[nobib, a4paper, 10pt, oneside, onecolumn, openany, notitlepage, draft,
svgnames, marginals=raggedouter, english,
%symmetric, justified,
]{article}
\usepackage{charter}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\input{packages}
\input{abbreviations}

\title{Edge sign prediction in general graphs and Correlation Clustering}
% \usetikzlibrary{external}
% \tikzexternalize[mode=list and make, failed ref warnings for={\ref,\cite,\pageref},]
\begin{document}
\newgeometry{vmargin=1.3cm,left=1.5cm,textwidth=150mm,marginparsep=5mm,marginparwidth=32mm}
\maketitle
% \tableofcontents

\subimport*{cc/}{main}

\begingroup
\newgeometry{hmargin=2cm,vmargin=1.3cm}
\setstretch{0.9}
\todos

\setlength\bibitemsep{2pt}
\printbibliography
\restoregeometry
\endgroup

\end{document}