### 2017-08-24 cc/cc_main.tex
change intro of CC

% *TODO*
% say more clearly there are 3 methods for approximation so far: LP & region growing; LP & distance
% pivot; combinatorial pivot
% say that in practice, general graphs are assumed to be complete with missing edges being negative
% - “The problem is agnostic in the sense that the clustering of the data is not taken into account
%   or even assumed to exist.” another way to look at it is that we're trying to get a binary
%   classifier of edges (+ or -) using an hypothesis class (nodes clustering) with limited
%   expression power (as it has to respect transitivity constraints, whereas signs can be inconsistent)
% - when the graph is weighted, there is no much point separating complete from
%   general since weights can be 0

As mentioned earlier, most real signed graphs are not fully balanced and thus do not have a
perfectly consistent clustering. The problem of quantifying how much a given graph departs from this
ideal situation is called \pcc{}. After defining it formally in \autoref{sub:problem_setting}, we
show that \pcc{} is a learning problem on its own, and discuss how it relates to the binary
classification problem of predicting edge signs, among many others applications. We then present
numerous methods to solve \pcc{} in \autoref{sub:state_of_the_art}, ranging from exact methods (that
implicitly leverage our bias) to heuristics (whose greedy and energy minimization frameworks are
classic in unsupervised graph cut problems). We also devote a large section to approximation
methods, in order to get a sense of why the problem is \NPc{}. This allows us to finish in
\autoref{sub:cc_non_worst_case} by describing situations related to our bias where is possible to
solve the problem efficiently and optimally. By doing so, we update existing surveys on the same
topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017}, highlighting the importance of
\pcc{} to solve our binary learning problem.

% surveyCC16 is about mind on complete graph (ie cluster editing). They give a Russian article from
% 1971 that says it cluster editing on triangle-free graph can be solved in polynomial time, which
% include bipartite graph (so for CC, if G^+ is (complete? I think so) bipartite)
% Fridman, G.Š.: A graph approximation problem. Upravlyaemye Sistemy 8, 73–75 (1971). (in Russian)
% In [14] it was observed that the problem is polynomially solvable for instances of at most 2
% clusters??! Actually it means consensus clustering can be solved in polynomial time when there are
% only 2 clustering as input

\subsection{Problem setting and connection to \esp{}}
\label{sub:problem_setting}
\input{cc_problem_applications}

\subsection{State of the art}
\label{sub:state_of_the_art}

\subsubsection{Hardness and approximation}
\label{ssub:cc_harness_approx}
\input{cc_approx}
\subsubsection{Proposed approaches}
\label{ssub:cc_methods}
\input{cc_proposed}

\subsection{Variants and extensions}
\label{sub:variants_and_extensions}
\input{cc_variants}

\subsubsection{\pcc{} under stability assumption}
\label{ssub:cc_under_stability_assumption}
\input{cc_multimincut_stable}