### 2017-08-24 cc/cc_multimincut_stable.tex
move text around for a state of the art following more the general direction of chapter

Whereas clustering objective functions are \NPh{} to optimize, we expect that meaningful instances
in which we are interested have additional structure which allows for guaranteed polynomial time
algorithms, see for instance \autocite{clusteringFeasibility15} for a critical overview of some
proposed such notions of structure. Informally, a general idea is that the clustering should not
change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
is \emph{$\alpha$-stable} (with $\alpha>1$) for some partition objective if its optimal partition
remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between $1$ and
$\alpha$. Finally, we say that an algorithm is \emph{robust} if, given an instance $\mathcal{I}$ and
in polynomial time, it: returns the optimal solution of $\mathcal{I}$ if $\mathcal{I}$ is
$\alpha$-stable; and if $\mathcal{I}$ is not $\alpha$-stable, either returns the optimal solution of
$\mathcal{I}$ or reports that $\mathcal{I}$ is not $\alpha$-stable. This is a handy property, as in
general, we cannot check the stability of an instance in practice as we do not know the optimal
solution hence we cannot tell whether it changes or not under perturbations.

\iffalse
\Textcite{StableCC17} provide a robust algorithm for $2-\nicefrac{2}{k}$-stable instance of
$k$-\textsc{Minimum Multiway Cut}, which therefore is not really useful in our case because I confused
multiway cut and multicut. To be fair, it seems I'm not the only one, \href{http://pages.cs.wisc.edu/~shuchi/papers/multicut-hardness-full.pdf}{on the hardness of
   approximating multicut}
   \enquote{multicut is known to be APX-hard
      (\href{https://pdfs.semanticscholar.org/1cf6/4c2bdd4f1c384a55910606a64c8d831a96ba.pdf}%
   {Dahlhaus et al. 1994}).} by that 1994 paper define the multiway cut problem, see
   \autoref{fig:cc_multiwhat}. \Textcite[Theorem 24]{Bansal2004} show that we can go from
   \textsc{Minimum Multiway Cut} to \pcc{} but I doubt about the other direction, mostly because
   \textsc{Minimum Multiway Cut} can actually be approximated with factor smaller than $1.3$.
   see \url{http://www.nowozin.net/sebastian/papers/nowozin2009lpstability.pdf} about LP multicut
   stability.
\begin{figure}[htpb]
   \centering
   \includegraphics[width=0.9\linewidth]{assets/raw/multicut_vs_multiway.png}
   \caption{Confusing statement} \label{fig:cc_multiwhat}
\end{figure}
\fi

% \enquote{We prove that if the Unique Games Conjecture of Khot (2002) is true, then for every constant L > 0
% it is NP-hard to approximate Multicut within factor L. If a quantitatively stronger version of the
% conjecture is true, then Multicut is NP-hard to approximate within a factor of $\Omega(\sqrt{\log
% \log n})$.}

% It's probably limited because LP with a lot of constraints do not scale that well (someone claimed
% $O(n^{4.5})$)

% SDP Gaps and UGC Hardness for Multiway Cut, 0-Extension and Metric Labelling,
% On the Unique Games Conjecture: Subhash Khot
% under the UGC, one cannot get better than $O(\log n)$ approximation using LP or SDP relaxation.