### 2017-08-25 cc/cc_main.tex
reorganize CC state of the art

As mentioned earlier, most real signed graphs are not fully balanced and thus do not have a
perfectly consistent clustering. The problem of quantifying how much a given graph departs from this
ideal situation is called \pcc{}. After defining it formally in \autoref{sub:problem_setting}, we
show that \pcc{} is a learning problem on its own, and discuss how it relates to the binary
classification problem of predicting edge signs, among many others applications. We then present
numerous methods to solve \pcc{} in \autoref{sub:state_of_the_art}, ranging from exact methods (that
implicitly leverage our bias) to heuristics (whose greedy and energy minimization frameworks are
classic in unsupervised graph cut problems). We also devote a large section to approximation
methods, in order to get a sense of why the problem is \NPc{} and even hard to approximate. This
allows us to finish in
\autoref{sub:cc_non_worst_case} by describing situations related to our bias where is possible to
solve the problem efficiently and optimally. In this section, we thus update existing surveys on the same
topic~\autocites{bonchi2014correlation}{surveyCC16}{CCWirth2017} and highlight the importance of
\pcc{} to solve our binary learning problem.

% surveyCC16 is about mind on complete graph (ie cluster editing). They give a Russian article from
% 1971 that says it cluster editing on triangle-free graph can be solved in polynomial time, which
% include bipartite graph (so for CC, if G^+ is (complete? I think so) bipartite)
% Fridman, G.Š.: A graph approximation problem. Upravlyaemye Sistemy 8, 73–75 (1971). (in Russian)
% In [14] it was observed that the problem is polynomially solvable for instances of at most 2
% clusters??! Actually it means consensus clustering can be solved in polynomial time when there are
% only 2 clustering as input

\subsection{Problem setting and connection to \esp{}}
\label{sub:problem_setting}
\input{cc_problem_and_esp}

\subsection{State of the art}
\label{sub:state_of_the_art}

\input{cc_state_of_art}

\subsection{Variants and extensions}
\label{sub:variants_and_extensions}
\input{cc_variants}

\subsubsection{\pcc{} under stability assumption}
\label{ssub:cc_under_stability_assumption}
\input{cc_multimincut_stable}