### 2017-09-02 troll/algs_online.tex

\section{Algorithms in the Online Setting}\label{s:algonline}

For the online scenario, we have the following result.
%
\begin{theorem}\label{t:online}
There exists a randomized online prediction algorithm $A$ whose expected number of mistakes satisfies
$
\E M_A(Y) = \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V| \right)
$
on any edge-labeled graph $G(Y) = (V,E(Y))$.
\end{theorem}
%
The algorithm used in Theorem~\ref{t:online} is a combination of randomized Weighted Majority instances. Details are reported in the supplementary material. We complement the above result by providing a mistake lower bound. Like Theorem~\ref{t:online}, the following result holds for all graphs, and for all label irregularity levels $\Psi_G(Y)$.
%
\begin{theorem}
\label{t:mistake_bound}
Given any edge-labeled graph $G(Y) = (V,E(Y))$ and any integer $K \le \big\lfloor \tfrac{|E|}{2}\big\rfloor$, a randomized labeling $Y\in\spin^{|E|}$ exists such that $\Psi_G(Y) \leq K$, and the expected number of mistakes that any online algorithm $A$ can be forced to make satisfies
\(
\E M_A(Y) \ge \frac{K}{2}\,.
\)
Moreover, as $\frac{K}{|E|} \rightarrow 0$ then
\(
\E M_A(Y) = K
\).
\end{theorem}
%

%\footnote
%{
%In fact, we can refine the above statement by proving that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$ for any $G(Y)$, hence asymptotically matching the upper bound of Theorem~\ref{t:online} (\textit{proof in the supplementary material}).
%}