### 2017-09-02 troll/introduction.tex
small pass over introduction

\section{Introduction}
% say that we are facing edge sign prediction from many viewpoints.
% say about trolls and trustworthy people in a social network
% say that we are using reductions to node classification (novelty)
% generative model
% novel and cool

Connections in social networks are mostly driven by the \emph{homophily} assumption: linked
individuals tend to be similar, sharing personality traits, attitudes, or interests. However,
homophily alone is clearly not sufficient to explain the variety of social links\Todo{link with
chapter 3}. In fact, sociologists have long studied networks, hereafter called \emph{signed} social
networks\Todo{DSSN?}, where also {\em negative} relationships ---like dissimilarity, disapproval or
distrust--- are explicitly displayed. The presence of negative relationships is also a feature of
many technology-mediated social networks. Known examples are \textsc{Ebay}, where users trust or
distrust agents in the network based on their personal interactions, \textsc{Slashdot}, where each
user can tag another user as friend or foe, and \textsc{Epinion}, where users can rate positively
or negatively not only products, but also other users. Even in social networks where connections
solely represent friendships, negative links can still emerge from the analysis of online debates
among users.

%Consider, in particular, a user joining an online community within a social network. His/her behavior will often fit one of these two stereotypes: the new member could play well with other users, establishing positive relationships with those who have been helpful. Or, the new user could try and disrupt the community by engaging into anti-social behavior, and creating conflictual relationships with other members. This behavioral dichotomy is supported by decades of research in psychology, starting with the seminal work~\autocite{Dissonance57} about cognitive dissonance (when someone is acting in contradiction with his/her personal beliefs, ideas, or values). This kind of attitude expressed publicly on social media leads to the definition of \enquote{trolls}~\autocite{Hardaker10}: users whose real intentions are to cause disruption and/or to trigger or exacerbate conflict for the purposes of their own amusement. The paper~\autocite{Shachaf10} elaborates on their motives, adding that boredom, attention seeking, and revenge motivate trolls; they find pleasure from causing damage to other people or to the community as a whole.

When the social network is signed, specific challenges arise in both network analysis and learning.
On the one hand, novel methods are required to tackle standard tasks (e.g., user clustering, link
prediction, targeted advertising/recommendation, prediction of user interests, and analysis of the
spreading of diseases in epidemiological models). On the other hand, new problems such as edge sign
prediction, which we consider here, naturally emerge. Edge sign prediction is the problem of
classifying the positive or negative nature of the links based on the network topology. Prior
knowledge of the network topology is often a realistic assumption, for in several situations the
discovery of the link sign can be more costly than acquiring the topological information of the
network. For instance, when two users of an online social network communicate on a public web page,
we immediately detect a link. Yet, the classification of the link sign as positive or negative may
require complex techniques.

From the modeling and algorithmic viewpoints, because of the huge amount of available networked
data, a major concern in developing learning methods for edge sign prediction is algorithmic
scalability. Many successful, yet simple heuristics for edge sign prediction are based on the
troll-trust features, i.e., on the fraction of outgoing negative links (trollness) and incoming
positive links (trustworthiness) at each node. We study such heuristics by defining a probabilistic
generative model for the signs on the directed links of a given network, and show that these
heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our
generative model. We also gather empirical evidence supporting our probabilistic model by observing
that a logistic model trained on trollness and trustworthiness features alone is able to learn
weights that, on all datasets considered in our experiments, consistently satisfy the properties
predicted by our model.

We then introduce suitable graph transformations defining reductions from edge sign prediction to
node sign prediction problems. This opens up the possibility of using the arsenal of known
algorithmic techniques developed for node classification. In particular, we show that a Label
Propagation algorithm, combined with our reduction, approximates the maximum likelihood estimator
of our probabilistic generative model. Experiments on real-world data show the competitiveness of
our approach in terms of both prediction performance (especially in the regime when training data
are scarce) and scalability.

Finally, we point out that the notions of trollness and trustworthiness naturally define a measure
of complexity, or learning bias, for the signed network that can also be used to design
\emph{online} (i.e., sequential) learning algorithms for the edge sign prediction problem. The
learning bias encourages settings where the nodes in the network have polarized features (e.g.,
trollness/trustworthiness are either very high or very low). Our online analysis holds under
adversarial conditions, namely, without any stochastic assumption on the assignment of signs to the
network links.

%Hence, it is necessary to depart from well-established yet computationally expensive approaches by relying on novel algorithmic techniques.

%In this paper, we consider the problem of learning link classifiers, and study this problem from different viewpoints. Our contributions are summarized as follows.

\subsection{Related work}

Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} with its weak
version~\autocite{davis1967clustering}.  The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. Among the several approaches
related to our work, some extend the spectral properties of a graph to the signed case in order to
find good embeddings for classification~\autocites{Kunegis2009}{SignedEmbedding15}. However, the
use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
makes those methods hardly scalable to large graphs. Another approach is based on mining ego
networks with SVM. Although this method seems to deliver good results~\autocite{Papaoikonomou2014},
the running time makes it often impractical for large real-world datasets. An alternative approach,
based on local features only and proposed in~\autocite{Leskovec2010}, relies on the so-called
status theory for directed graphs~\autocite{guha2004propagation}. Some works in active learning,
using a more sophisticated bias based on the correlation clustering (CC)
index~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}, provide strong theoretical guarantees.
However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering
of the nodes with a small CC index.\Todo{don't put it that way since we use that bias later}

Whereas our focus will be on \emph{binary} prediction, researchers have also considered a weighted
version of the problem, where edges measure the amount of trust or distrust between two
users~\autocites{guha2004propagation}{Tang2013}{Bachi2012}{Qian2014sn}. Other works have also
considered versions of the problem where side information related to the network is available to
the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
Epinion in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
transfer learning from one network to another. While many of these approaches have interesting
performances, they often require extra information which is not always available (or reliable) and,
in addition, may face severe scaling issues.

The survey~\autocite{Tang2015a} contains pointers to many papers on edge sign prediction for
signed networks, especially in the Data Mining area. Additional references, more closely related to
our work, will be mentioned at the end of \autoref{ss:passive}.