### 2017-09-05 troll/supp_body.tex
first pass on troll appendix

\section{Proofs from \autoref{s:algbatch}}

\subsection{Proof of \autoref{t:active}}

The following ancillary results will be useful.

\begin{lemma}[Hoeffding's inequality for sampling without replacement]
\label{l:hoeff-wr}
Let $\scX = \{x_1,\dots,x_N\}$ be a finite subset of $[0,1]$ and let
\[ \mu = \frac{1}{N}\sum_{i=1}^N x_i~.  \]
If $X_1,\dots,X_n$ is a random sample drawn at random from $\scX$ without replacement, then, for
every $\ve > 0$,
\[
  \Pr\left( \left|\frac{1}{n}\sum_{t=1}^n X_t - \mu \right| \ge \ve \right) \le 2e^{-2n\ve^2}~.
\]
\end{lemma}

\begin{lemma}\label{l:bins}
% Given a function $f\,:\, \Nat \rightarrow \R$ such that $f(n) \rightarrow \infty$ as $n \rightarrow \infty$, 
Let $\NN_1, \ldots, \NN_n$ be subsets of a finite set $E$.
% with $|E| \geq n\,f(n)$
% Let $\theta = \frac{|E|}{n\,f(n)}$, 
% and $k = k(\theta)$ be the number of subsets $\NN_i$ such that $|\NN_i| \geq \theta$. Let 
Let $E_0 \subseteq E$ be sampled uniformly at random {\em without} replacement from $E$, with $|E_0| = m$. Then, for $\delta \in (0,1)$, $Q > 0$, and $\theta \geq 2\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$, we have
\[
\Pr\Bigl( \exists i \,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\Bigl) \leq \delta
\]
provided $|E| \geq m \geq \frac{2|E|}{\theta}\times\max\left\{Q,4\ln \frac{n}{\delta}\right\}$.
\end{lemma}

\begin{proof}[Proof of Lemma \ref{l:bins}]
Set for brevity $p_i = |\NN_i|/|E|$. Then, due to the sampling without replacement, each random
variable $|\NN_i \cap E_0|$ is the sum of $m$ dependent Bernoulli random variables $X_{i,1}, \ldots,
X_{i,m}$ such that $\Pr(X_{i,t} = 1) = p_i$, for $t = 1, \ldots, m$. Let $i$ be such that $|\NN_i|
\geq \theta$. Then the condition $m \geq \frac{2|E|Q}{\theta}$ implies
\[
Q \leq \frac{m\theta}{2|E|} \leq \frac{m\,p_i}{2} = \frac{\E\big[|\NN_i \cap E_0|\big]}{2}~.
\]
Since the variables $X_{u,v}$ are negatively associated, we may apply a (multiplicative) Chernoff
bound~\autocite[Section~3.1]{dpbook}.
This gives
\[
\Pr\big(|\NN_i \cap E_0| < Q\big) \leq e^{-\frac{m\,p_i}{8}} \leq e^{-\frac{m\,\theta}{8|E|}}
\]
so that
$
\Pr\bigl( \exists i\,:\, |\NN_i| \geq \theta, |\NN_i \cap E_0| < Q\bigr) \leq n\,e^{-\frac{m\,\theta}{8|E|}}
$,
which is in turn upper bounded by $\delta$ whenever 
$m \geq \frac{8|E|}{\theta}\ln \frac{n}{\delta}$. % = 8\,n\,f(n)\,\ln \frac{n}{\delta}$.
\end{proof}

Let now $E_{\theta} = \theset{(u,v) \in E}{\din(v) \ge \theta,\, \dout(u) \ge \theta} \setminus
E_0$, where $E_0 \ss E$ is the set of edges sampled by the learning algorithm of
\autoref{s:algbatch}. Then \autoref{t:active} in the main paper is an immediate consequence of the
following lemma.

\begin{lemma}\label{l:active}
Let $G(Y) = (V,E(Y))$ be a directed graph with labels on the edges generated according to the model
in \autoref{s:gen}.
For all $0 < \alpha,\delta < 1$ and $0 < \ve < \tfrac{1}{16}$, if the learning algorithm of
\autoref{s:algbatch} is run with parameters $Q = \tfrac{1}{2\ve^2}\ln\frac{4|V|}{\delta}$ and
$\alpha$, then with probability at least $1-11\delta$ the predictions $\yhat(u,v)$ satisfy
$\yhat(u,v) = y^*(u,v)$ for all $(u,v) \in E_{\theta}$ such that $\big|\eta(u,v) - \tfrac{1}{2}\big|
> 8\ve$.
\end{lemma}

\begin{proof}[Proof of Lemma \ref{l:active}]
We apply \autoref{l:bins} with $\theta = \tfrac{2Q}{\alpha} \ge 2\times\max\big\{Q,4\ln
\frac{2|V|+1}{\delta}\big\}$ to the $2|V|+1$ subsets of $E$ consisting of $E_L$ and
$\Nin(u),\Nout(u)$, for $i \in V$. We have that, with probability at least $1-\delta$, at least $Q$
edges of $E_L$ are sampled, at least $Q$ edges of $\Nin(u)$ are sampled for each $i$ such that
$\NNin(u) \ge \theta$, and at least $Q$ edges of $\Nout(v)$ are sampled for each $j$ such that
$\NNout(v) \ge \theta$.
For all $(u,v) \in E_{\theta}$ let
\[
  \pbar_j = \frac{1}{\din(v)}\sum_{i\in\NNin(v)} p_i \quad\text{and}\quad \qbar_i = \frac{1}{\dout(u)}\sum_{j\in\NNout(u)} q_j
\]
and set for brevity $\hdeltain(v) = 1-\hun(v)$ and $\hdeltaout(u) = 1-\htr(u)$.
We now prove that $\hdeltain(v)$ and $\hdeltaout(u)$ are concentrated around their expectations for
all $(u,v) \in E_{\theta}$. Consider $\hdeltaout(u)$ (the same argument works for $\hdeltain(v)$).
Let $J_1,\dots,J_Q$ be the first $Q$ draws in $E_0 \cap \NNout(u)$ and define
\[
  \muhat_p(u) = \frac{1}{Q} \sum_{t=1}^Q \frac{p_i+q_{J_t}}{2}~.
\]
Applying \autoref{l:hoeff-wr} to the set $\theset{\tfrac{p_i+q_j}{2}}{j \in \NNout(u)}$, and using
our choice of $Q$, we get that $\big|\muhat_p(u) - \mu_p(u)\big| \le \ve$ holds with probability at
least $1-\delta/(2|V|)$, where
\[
  \mu_p(u) = \frac{1}{\dout(u)}\sum_{j \in \NNout(u)} \frac{p_i + q_j}{2} = \frac{p_i + \qbar_i}{2}~.
\]
Now consider the random variables $Z_t = \Ind{y_{u,v_t}=1}$, for $t=1,\dots,Q$. Conditioned on
$J_1,\dots,J_Q$, these are independent Bernoulli random variables with $\E[Z_t \mid J_t] =
\tfrac{p_i+q_{J_t}}{2}$. Hence, applying a standard Hoeffding bound for independent variables and
using our choice of $Q$, we get that
\[
  \left|\frac{1}{Q}\sum_{t=1}^Q Z_t - \muhat_p(u) \right| \le \ve 
\]
with probability at least $1-\delta/(2|V|)$ for every realization of $J_1,\dots,J_Q$.

Since $\hdeltaout(u) = (Z_1+\cdots+Z_Q)/Q$, we get that $\big|\hdeltaout(u) - \mu_p(u)\big| \le
2\ve$ with probability at least $1-2\delta/(2|V|)$. Applying the same argument to $\hdeltain(v)$,
and the union bound\footnote{The sample spaces for the ingoing and outgoing edges of the vertices
occurring as endpoints in $E_{\theta}$ overlap. Hence, in order to prove a uniform concentration
result, we need to apply the union bound over the random variables defined over these sample spaces,
which motivates the presence of the factor $\ln(2|V|)$ in the definition of $Q$.}
on the set $\theset{\hdeltain(v),\hdeltaout(u)}{(u,v) \in E_{\theta}}$, we get that
\begin{equation}
  \label{eq:first}
  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\pbar_j + \qbar_i}{2} \right| \le 4\ve
\end{equation}
simultaneously holds for all $(u,v) \in E_{\theta}$ with probability at least $1 - 4\delta$. Now
notice that $\pbar_j$ is a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables drawn from the
prior marginal $\int_0^1 \mu\big(\cdot,q\bigr) dq$ with expectation $\mu_p$. Similarly, $\qbar_i$ is
a sample mean of $Q$ i.i.d.\ $[0,1]$-valued random variables independently drawn from the prior
marginal $\int_0^1 \mu\big(p,\cdot\big) dp$ with expectation $\mu_q$. By applying Hoeffding bound
for independent variables, together with the union bound to the set of pairs of random variables
whose sample means are $\pbar_j$ and $\qbar_i$ for each $(u,v) \in E_{\theta}$ (there are at most
$2|V|$ of them) we obtain that
\begin{align*}
  \big|\pbar_j - \mu_p\big| \le \ve
  \qquad\text{and}\qquad
  \big|\qbar_i - \mu_q\big| \le \ve
\end{align*}
hold simultaneously for all $(u,v) \in E_{\theta}$ with probability at least $1-2\delta$. Combining
with~\eqref{eq:first} we obtain that
\begin{equation}
  \label{eq:this}
  \left| \hdeltaout(u) + \hdeltain(v) - \frac{p_i + q_j}{2} - \frac{\mu_p + \mu_q}{2} \right| \le 5\ve
\end{equation}
simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 6\delta$. Next,
let $E_L'$ be the set of the first $Q$ edges drawn in $E_L \cap E_0$. Then
\[
  \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \Pr\big(y_{u,v} = 1\big) = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~,
\]
where the expectation is w.r.t.\ the independent draws of the labels $y_{u,v}$ for $(u,v) \in E_L'$.
Hence, by applying again Hoeffding bound (this time without the union bound) to the $Q =
\tfrac{1}{2\ve^2}\ln\tfrac{2}{\delta}$ independent Bernoulli random variables $\Ind{y_{u,v} = 1}$,
$(u,v) \in E_L'$, the event
$
\big| \tauhat - \E\big[\tauhat\big] \big| \le \ve
$
holds with probability at least $1-\delta$. Now, introduce the function
\[
  F(\bp,\bq) = \E\big[\tauhat\big] = \frac{1}{Q}\sum_{(u,v) \in E_L'} \frac{p_i+q_j}{2}~.
\]
For any realization $\bq_0$ of $\bq$, the function $F_1(\bp) = F(\bp,\bq_0)$ is a sample mean of $Q
= \tfrac{1}{2\ve^2}\ln\tfrac{4|V|}{\delta}$ i.i.d.\ $[0,1]$-valued random variables
$\theset{p_i}{(u,v) \in E_L'}$ (recall that if $i \in V$ is the origin of an edge $(u,v) \in E_L'$,
then it is not the origin of any other edge $(u,v') \in E_L'$). Using again the standard Hoeffding
bound, we obtain that
\[
  \left| F(\bp,\bq) - E_{\bp}\big[F(\bp,\bq)\big] \right| \le \ve
\]
holds with probability at least $1 -\delta$ for each $\bq\in [0,1]^{|V|}$. With a similar argument,
we obtain that
\[
  \left| E_{\bp}\big[F(\bp,\bq)\big] - E_{\bp,\bq}\big[F(\bp,\bq)\big] \right| \le \ve
\]
also holds with probability at least $1 -\delta$. Since
\[
  E_{\bp,\bq}\big[F(\bp,\bq)\big] = \frac{\mu_p+\mu_q}{2}
\]
we obtain that
\begin{equation}
  \label{eq:that}
  \Big| \tauhat - \frac{\mu_p+\mu_q}{2} \Big| \le 3\ve
\end{equation}
with probability at least $1-3\delta$. Combining~(\ref{eq:this}) with~(\ref{eq:that}) we obtain
\[
  \left| \hdeltaout(u) + \hdeltain(v) - \tauhat - \frac{p(u) + q(v)}{2} \right| \le 8\ve
\]
simultaneously holds for each $(u,v) \in E_{\theta}$ with probability at least $1 - 10\delta$.
Putting together concludes the proof.
\end{proof}

\subsection{Derivation of the maximum likelihood equations}

Recall that the training set $\trainset = \theset{\big(u_t,v_t),y_{u_t,v_t}\big)}{t=1,\dots,m}$
is drawn uniformly at random from $E$ without replacement. 
We can write

\begin{align*}
  \Pr&\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) = 
  \frac{1}{\binom{|E|}{m}\,m!}\prod_{k=1}^m \left(\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=+1}}\,
  \times\,\prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{v_k}}{2} \right)^{\Ind{y_{u_k,v_k}=-1}}\\
  &=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=+1}}\,\times\,
    \prod_{k=1}^m\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)^{\Ind{u_k = \ell,\,y_{\ell,v_k}=-1}}\Biggl)
\end{align*}

so that $\log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)$ is proportional to

\begin{equation*}
  \sum_{\ell=1}^{|V|}\left(\sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1}\,
  \log\left(\frac{p_{\ell}+q_{v_k}}{2} \right)\, +\,
  \sum_{\ell=1}^{|V|}\sum_{k=1}^m \Ind{u_k = \ell,y_{\ell,v_k}=+1}\, 
\log\left(1-\frac{p_{\ell}+q_{v_k}}{2} \right)\right)
\end{equation*}

and

\begin{equation*}
  \frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial p_{\ell}}
  = \sum_{k=1}^m 
  \frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
   - \,
  \sum_{k=1}^m \frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}}\,.
\end{equation*}

By a similar argument,

\begin{align*}
  \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right) 
=
  \frac{1}{\binom{|E|}{m}\,m!}
  \prod_{\ell=1}^{|V|}\Biggl(\prod_{k=1}^m\,
    \left(\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=+1}}\,
\times\,
    \prod_{k=1}^m\left(1-\frac{p_{u_k}+q_{\ell}}{2} \right)^{\Ind{v_k = \ell,\,y_{u_k,\ell}=-1}}\Biggl)
\end{align*}

so that

\begin{align*}
\frac{\partial \log \Pr\left(E_0\,\Big|\, \{p_i, q_i\}_{i=1}^{|V|}\right)}{\partial q_{\ell}}
= \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=+1}}{p_{u_k}+q_{\ell}}\,
- \, \sum_{k=1}^m \frac{\Ind{v_k = \ell,y_{u_k,\ell}=-1}}{2-p_{u_k}-q_{\ell}}\,.
\end{align*}

We then derive the approximation presented in the main paper. Namely, equating to zero the
gradient of the log likelihood w.r.t $p_\ell$ gives
\begin{align*}
\sum_{k=1}^m 
\frac{\Ind{u_k = \ell,y_{\ell,v_k}=+1}}{p_{\ell}+q_{v_k}}\,
 - \,
\frac{\Ind{u_k = \ell,y_{\ell,v_k}=-1}}{2-p_{\ell}-q_{v_k}} &= 0
\end{align*}

To simplify the notation, let $a_k = \Ind{u_k = \ell,y_{\ell,v_k}=+1}$, $b_k=\Ind{u_k =
\ell,y_{\ell,v_k}=+1}$ and $c_k = p_{\ell} + q_{v_k}$, we can rewrite the previous equation as 
\begin{align*}
  \sum_{k=1}^m \frac{a_k}{c_k} - \frac{b_k}{2-c_k} = 0 =
  \sum_{k=1}^m  \frac{a_k(2-c_k) - b_k c_k }{ c_k(2-c_k) }
\end{align*}

The approximation consists in assuming that the denominator $c_k(2-c_k)$ is a
constant for all $k$, and can therefore be disregarded.
Moving $b_kc_k$ to the right hand side and returning to the original variables,
it yields the approximate equation presented in the main paper, namely
\begin{align*}
  \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=+1} \left(2-p_{\ell}-q_{v_k}\right)
  = \sum_{k=1}^m \Ind{i_k = \ell,y_{\ell,v_k}=-1}
  (p_{\ell}+q_{v_k})
\end{align*}

By applying a similar process to the gradient w.r.t. $q_\ell$, we obtain the
following approximate equation
\begin{align*}
\sum_{k=1}^m \Ind{v_k = \ell,y_{u_k,\ell}=+1} \left(2-p_{u_k}-q_{\ell}\right)
= \sum_{k=1}^m  \Ind{v_k = \ell,y_{u_k,\ell}=-1}
\left(p_{u_k}+q_{\ell}\right)
\end{align*}

\subsection{Label propagation on $G''$}
\label{ssec:troll_lprop_weights}

Here we provide more details on the choice of weight for the edges of $G''$, as well as an
explanation on why we temporarily use symmetrized variables lying in $[-1, 1]$ (which we will denote
with primes, so that for instance $p'_u = 2p_u-1$). Since only the ratio between the negative and
positive weights matters, we fix the negative weight of the edges in $E''\setminus E'$ to be $-1$
and we denote by $\eps$ the weight of edges in $E'$.  With these notations, Label Propagation on
$G''$ seeks the harmonic minimizer of the following expression

\begin{equation*}
  \frac{1}{16} \sum_{u,v \in E} \Bigl[ \eps\left(y_{u,v} - p_u'\right)^2 + \eps\left(y_{u,v} - q_v'\right)^2 + (p'_i+q_v')^2 \Bigr]
\end{equation*}

which can be successively rewritten as

\begin{align*}
  \frac{1}{16} \sum_{u,v \in E} & \Bigl[ \eps\left(y_{u,v}+1 - 2p_u\right)^2 + \eps\left(y_{u,v}+1 - 2q_v\right)^2
	+ (2p_u +2q_v -2)^2 \Bigr]
\\=
	\frac{1}{8} \sum_{u,v \in E} & \Biggl[ 2\eps\!\left(\frac{y_{u,v}+1}{2} - p_u\right)^2\!\! +  2\eps\!\left(\frac{y_{u,v}+1}{2} - q_v\right)^2
	+ 8\left(\frac{p_u +q_v -1}{2}\right)^2 \Biggr]
\\=
  \frac{1}{8}\sum_{u,v \in E} & \Biggl[ 2\eps\left(\left(\frac{y_{u,v}+1}{2}\right)^2 -
    p_u(1+y_{u,v}) + p_u^2\right) + %\\
                              2\eps\left(\left(\frac{y_{u,v}+1}{2}\right)^2 - q_v(1+y_{u,v}) + q_v^2\right) +  \\
                              & 8\left(\left(\frac{p_u +q_v}{2}\right)^2 - \frac{p_u+q_v}{2} + \frac{1}{4} \right) \Biggr]
\\=
  \frac{1}{8}\sum_{u,v \in E} & 4\Biggl(\eps\!\left(\frac{y_{u,v}+1}{2}\right)^2\!\! -
  2\eps\!\left(\frac{y_{u,v}+1}{2}\right)\left(\frac{p_u+q_v}{2}\right)
% \\&
  + 2\!\left(\frac{p_u +q_v}{2}\right)^2\Biggr)
\\&+
  \sum_{u,v \in E} \Bigl[ \left(2\eps p_u^2 - 4p_u + 1\right) + \left(2\eps q_v^2 - 4q_v + 1\right) \Bigr]
\end{align*}

By setting $\eps=2$, we can factor this expression into 
\begin{align*}
  \sum_{u,v \in E} &\left(\frac{y_{u,v}+1}{2} - \frac{p_u+q_v}{2}\right)^2
  + \frac{1}{2}\sum_{u,v \in E} \left(\left(p_u - \frac{1}{2}\right )^2 + \left(q_v -
  \frac{1}{2}\right)^2 \right)~.
\end{align*}

\section{Proofs from \autoref{s:algonline}}

\begin{proof}[Proof of \autoref{t:online}]

Let each node $u \in V$ host two instances of the randomized Weighted Majority (RWM)
algorithm~\cite{LittlestoneWa94} with an online tuning of their learning rate~\cite{cb+97,acg02}:
one instance for predicting the sign of outgoing edges $(u,v)$, and one instance for predicting the
sign of incoming edges $(v,u)$. Both instances simply compete against the two constant experts,
predicting always $+1$ or always $-1$. Denote by $M(u,v)$ the indicator function (zero-one loss) of
a mistake on edge $(u,v)$. Then the expected number of mistakes of each RWM instance
satisfy~\cite{cb+97,acg02}:
\[
    \sum_{v \in \NNout(u)} \E\,M(u,v) = \Psiout(u,Y) + O\left(\sqrt{\Psiout(u,Y)}+ 1\right)
\]
and
\[
    \sum_{u \in \NNin(v)} \E\,M(u,v) = \Psiin(v,Y) + O\left(\sqrt{\Psiin(v,Y)} + 1\right)~.
\]
We then define two meta-experts: an ingoing expert, which predicts $y_{u,v}$ using the prediction of
the ingoing RWM instance for node $v$, and the outgoing expert, which predicts $y_{u,v}$ using the
prediction of the outgoing RWM instance for node $u$.
The number of mistakes of these two experts satisfy
\begin{align*}
  \sum_{u \in V}\sum_{v \in \NNout(u)} \E\,M(u,v)
                                       &= \Psiout(Y) + O\left(\sqrt{|V|\Psiout(Y)} + |V|\right)\\
  \sum_{v \in V}\sum_{u \in \NNin(v)} \E\,M(u,v)
                                      &= \Psiin(Y)  + O\left(\sqrt{|V|\Psiin(Y)}  + |V|\right)~,
\end{align*}
where we used $\sum_{v \in V} \sqrt{\Psiin(v,Y)} \leq \sqrt{|V|\Psiin(Y)}$, and similarly for
$\Psiout(Y)$.
Finally, let the overall prediction of our algorithm be a RWM instance run on top of the ingoing and
the outgoing experts. Then the expected number of mistakes of this predictor satisfies~
\begin{align*}
  \sum_{(u,v) \in E} \E\,M(u,v) &= \Psi_G(Y) + O\Biggl(\sqrt{|V|\Psi_G(Y)} + |V|
    + \sqrt{\left(\Psi_G(Y) + |V| + \sqrt{|V|\Psi_G(Y)}\right)} \Biggl)\\
    &= \Psi_G(Y) + O\left(\sqrt{|V|\Psi_G(Y)} + |V|\right)
\end{align*}
as claimed.
\end{proof}

\begin{proof}[Proof sketch of \autoref{t:mistake_bound}]

Let $\mathcal{Y}_K$ be the set of all labelings $Y$ such that the total number of negative and
positive edges are $K$ and $|E|-K$, respectively (without loss of generality we will focus on
negative edges). Consider the randomized strategy that draws a labeling $Y \in \spin^{|E|}$ of the
edges of the input graph \uar{} from $\mathcal{Y}_K$. For each node $u \in V$, we have $\Psiin(u,Y)
\le \din^-(u)$, which implies $\Psiin(Y) \le K$. A very similar argument applies to the outgoing
edges, leading to $\Psiout(Y) \le K$. The constraint $\Psi_G(Y) \le K$ is therefore always
satisfied.

The adversary will force on average \shalf{} mistakes in each one of the first $K$ rounds of the
online protocol by repeating $K$ times the following: (i) A label value $\ell\in\{-1,+1\}$ is
selected uniformly at random. (ii) An edge $(u,v)$ is sampled uniformly at random from the set of
all edges that were not previously revealed and whose labels are equal to $\ell$.

The learner is required to predict $y_{u,v}$ and, in doing so, \shalf{} mistakes will be clearly
made on average because of the randomized labeling procedure. Observe that this holds even when $A$
knows the value of $K$ and $\Psi_G(Y)$. Hence, we can conclude that the expected number of mistakes
that $A$ can be forced to make is always at least $K/2$, as claimed.

\medskip

We now show that, as $\frac{K}{|E|} \rightarrow 0$, the lower bound gets arbitrarily close to $K$
for any $G(Y)$ and any constant $K$.
Let $\mathcal{E}$ be the following event: There is at least one unrevealed negative label.
The randomized iterative strategy used to achieve this result is identical to the one described
above, except for the stopping criterion. Instead of repeating step (i) and (ii) only for the first
$K$ rounds, these steps are repeated until $\mathcal{E}$ is true.
%
Let $m_{r,c}$ be defined as follows: For $c=1$ it is equal to the expected number of mistakes forced
in round $r$ when $K=1$. For $c > 1$ it is equal to the difference between the expected number of
mistakes forced in round $r$ when $K=c$ and $K=c-1$. One can see that $m_{r,c}$ is 
%either 
null when $r<c$. 
%and, for any given $|E|$ and $K$, when $r>c+|E|-K-1$.
%at least one of the following  conditions is met: , . 
%$c>K$, 
%
When $K=1$, the probability that $\mathcal{E}$ is true in round $r$ is clearly equal to
$\frac{1}{2^{r-1}}$.
Hence, the expected number of mistakes made by $A$ when $K=1$ in any round $r$ is equal to 
\(
\frac{1}{2}\,\frac{1}{2^{r-1}} = \frac{1}{2^{r}}.
\)
We can therefore conclude that $m_{r,1}=\frac{1}{2^{r}}$ for all $r$.

%When $r> 1$, 
A simple calculation shows that if $r=c$ then $m_{r,c}=\frac{1}{2^r}$. Furthermore, when $r>1$ and $c>1$, 
%for all $r$ and $c$ values such that $1 < c \le K$ and $c < r \le c+|E|-K-1$, 
we have the following recurrence:
%(holding even when $\frac{K}{|E|}$ does not converge to 0)
\[
m_{r,c}=\frac{m_{r-1,c}+m_{r-1,c-1}}{2}~.
\]  
In order to calculate $m_{r,c}$ for all $r$ and $c$, we will rest on the ancillary quantity
$s_j(i)$, defined recursively for any positive integer variables $i$ and $j$ by
$ s_0(i)=1 $
and
\[
s_j(i)=\sum_{k=1}^i s_{j-1}(k)~.
\]
(see \autoref{tab:sij}) for examples)

\begin{center}
  \begin{tabular}{|c |M{.45cm}|M{.45cm}|M{.45cm}|M{.45cm}|N}
    \hline
    \diaghead{j\ i}{\ \ \ 	{\footnotesize \textit{\textbf{j}}}\ }{\ \ {\footnotesize \textit{\textbf{i}}}\ \ \ } & \textbf{0}& \textbf{1} & \textbf{2} & \textbf{3}    &\\[13pt] \hline
    \textbf{1} & 1 & 1 & 1 & 1  &\\[13pt] \hline
    \textbf{2} & 1 & 2 & 3 & 4 &\\[13pt] \hline
    \textbf{3} & 1 & 3 & 6 & 10 &\\[13pt] \hline
    \textbf{4} & 1 & 4 & 10 & 20 &\\[13pt] \hline
    \textbf{5} & 1 & 5 & 15 & 35 &\\[13pt] \hline
    \textbf{6} & 1 & 6 & 21 & 56 &\\[13pt] \hline
    \textbf{7} & 1 & 7 & 28 & 84 &\\[13pt] \hline
  \end{tabular}
  \captionof{table}{Values of $s_j(i)$ for $i \le 7$  and $j \le 3$.} \label{tab:sij} 
\end{center}

It is not difficult to verify that 
\[
m_{r,c}=\frac{s_{c-1}(r-c+1)}{2^{r}}. 
\]
Since $s_j(i)=\frac{\langle i\rangle_j}{j!}$, where $\langle i\rangle_j$ is the rising factorial
$i(i+1)(i+2)\ldots(i+j-1)$, we have 
\[
m_{r,c}=\frac{\langle r-c+1\rangle_{c-1}}{(c-1)!2^{r}}.
\] 
(see \autoref{tab:mrc}) for examples)

\begin{center}
  \begin{tabular}{|c |M{.77cm}|M{.77cm}|M{.77cm}|M{.77cm}|N}
    \hline
    \diaghead{\theadfont i}%
    {{\footnotesize \textit{\textbf{c}}}}{{\footnotesize \textit{\textbf{r}}}}
               & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} &  \\ [13pt] \hline
    \textbf{1} & $1/2^1$    & 0          & 0          & 0          &  \\ [13pt] \hline
    \textbf{2} & $1/2^2$    & $1/2^2$    & 0          & 0          &  \\ [13pt] \hline
    \textbf{3} & $1/2^3$    & $2/2^3$    & $1/2^3$    & 0          &  \\ [13pt] \hline
    \textbf{4} & $1/2^4$    & $3/2^4$    & $3/2^4$    & $1/2^4$    &  \\ [13pt] \hline
    \textbf{5} & $1/2^5$    & $4/2^5$    & $6/2^5$    & $4/2^5$    &  \\ [13pt] \hline
    \textbf{6} & $1/2^6$    & $5/2^6$    & $10/2^6$   & $10/2^6$   &  \\ [13pt] \hline
    \textbf{7} & $1/2^7$    & $6/2^7$    & $15/2^7$   & $20/2^7$   &  \\ [13pt] \hline
  \end{tabular}
  \captionof{table}{Values of $m_{r,c}$ for $r \le 7$, $c \le 4$ and $|E| \rightarrow \infty$.}
  \label{tab:mrc} 
\end{center}

\bigskip
%Given any $K'>1$, the difference between the expected number of mistakes %forced when $K=K'$ and $K=K'-1$ is equal to 
%\[
%\sum_{r=K'}^{|E|-1} m_{r,K'}=
%\sum_{r=K'}^{|E|-1} \frac{\langle r-K'+1\rangle_{K'-1}}{(K'-1)!2^{r}}.
%\]

When $\frac{K}{|E|} \rightarrow 0$, given any integer $K'>1$, the difference between the expected
number of mistakes forced when $K=K'$ and $K=K'-1$ is equal to
%
\begin{align*}
  \sum_{r=K'}^{\infty} m_{r,K'}
  &=
  \frac{1}{(K'-1)!}\sum_{r=K'}^{\infty} \frac{\langle r-K'+1\rangle_{K'-1}}{2^{r}}\\
  &=
  \frac{1}{(K'-1)!2^{K'-1}}\sum_{r'=1}^{\infty} \frac{\langle r'\rangle_{K'-1}}{2^{r'}}~,
\end{align*}
% 
where we set $r'=r-K'+1$.
%Taking into account the following recurrence relation sequence 
Setting $i'=i-1$ and recalling that
\[
\langle i\rangle_j=j!{{i+j-1}\choose{i-1}}~,
\]
we have
\[
\frac{1}{j!}\sum_{i=1}^{\infty}\frac{\langle i\rangle_j}{2^i}=
\sum_{i=1}^{\infty}\frac{{{i+j-1}\choose{i-1}}}{2^i}=
\sum_{i'=0}^{\infty}\frac{{{i'+j}\choose{i'}}}{2^{i'+1}}~.
\]
Now, using the identity
\[
{i'+j+1 \choose i'} = {i'+j \choose i'} + {i'+j \choose i'-1}~,
\]
we can easily prove by induction on $j$ that
\[
\sum_{i'=0}^{\infty}\frac{{{i'+j}\choose{i'}}}{2^{i'+1}}=2^j~.
\]
Hence, we have
%\[
%\frac{1}{j!}\sum_{i=1}^{\infty}\frac{\langle i\rangle_j}%{2^i}=2^j~,
%\]
\[
\sum_{r=K'}^{\infty} m_{r,K'}=1.
\]
Moreover, as shown earlier, $m_{r,1}=\frac{1}{2^{r}}$ for all $r$. Hence we can conclude that when $\frac{K}{|E|} \rightarrow 0$ 
\[
\E M_A(Y) \ge
\sum_{r=1}^{\infty} \frac{1}{2^{r}} +
\sum_{K'=2}^{K} \sum_{r=K'}^{\infty} m_{r,K'}=K
\]
for any edge-labeled graph $G(Y)$ and any constant $K$, as claimed.
\end{proof}

\iffalse %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the special case when the bias is
\[
    \Psi(Y) = \min\big\{\Psiin(Y),\Psiout(Y)\big\}
\]
where
\[
    \Psiin(Y) = \sum_{j \in V} \min\big\{\din^-(v),\din^+(v)\big\}
\]
and
\[
    \Psiout(Y) = \sum_{i \in V} \min\big\{\dout^-(u),\dout^+(u)\big\}
\]
then using Randomized Weighted Majority on the two experts
\[
    \Yhat^{(1)}_{u,v} = \sgn\left(\frac{1}{2}-\frac{\hdout^-(u)}{\hdout(u)}\right)
\qquad\text{and}\qquad
    \Yhat^{(2)}_{u,v} = \sgn\left(\frac{1}{2}-\frac{\hdin^-(v)}{\hdin(v)}\right)
\]
we get an expected mistake bound of
\[
    2\Psi(Y) + \sqrt{(4\ln 2) \Psi(Y)}~.
\]
\fi %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Further Experimental Results}

This section contains more evidence related to the experiments in \autoref{s:exp}.
In particular, we experimentally demonstrate the alignment between \usrule{} and \uslogregp{}.

\iffalse
\subsection{Accuracy Results}
%
\autoref{tab:all_acc} contains the accuracy results on training sets of increasing size. Figures are
averaged over 12 random draws of the training set.

\setlength{\tabcolsep}{3pt}
{\scriptsize
\begin{longtable}{lrccc|cccc}
\caption{Like in the main paper, we tabulate Accuracy $\pm$ one standard deviation with increasing training set size, averaged over 12 random draws of the training set. The best accuracy performance is highlighted in \textbf{\textcolor{brown}{bold brown}} and the second one in \textit{\textcolor{red}{italic red}}. If the difference is statistically significant (as defined by a paired Student's $t$-test $p$-value less than $0.005$), the best score is underlined. The "time" rows contain the time taken to train on a $15\%$ training set. \label{tab:all_acc}}\\
\toprule
                                                  & $\frac{|\trainset{}|}{|E|}$ &       \uslogregp{} &          \usrule{} &     \uslpropGsec{} &                \compranknodes{} &                \compbayesian{} &     \complowrank{} &                \comptriads{} \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\aut{}}} & $3\%$  &  $57.14 \pm 0.78$  &  $57.15 \pm 1.31$  &  $60.76 \pm 2.76$  &               $68.99 \pm 0.89$  &    $\vsecond{72.02} \pm 0.26$  &  $70.18 \pm 0.33$  &   $\vfirst{72.44} \pm 0.13$  \\
                                                  %& $5\%$  &  $62.07 \pm 0.79$  &  $61.97 \pm 1.17$  &  $66.80 \pm 1.53$  &               $69.44 \pm 0.69$  &     $\vfirst{71.90} \pm 0.33$  &  $70.49 \pm 0.27$  &  $\vsecond{71.87} \pm 0.65$  \\
                                                  %& $7\%$  &  $65.01 \pm 0.50$  &  $65.98 \pm 0.72$  &  $68.76 \pm 1.63$  &  $\vsecondSig{71.21} \pm 0.69$  &  $\vfirstSig{72.89} \pm 0.54$  &  $71.03 \pm 0.27$  &            $70.69 \pm 0.54$  \\
                                                  & $9\%$  &  $67.07 \pm 0.56$  &  $68.29 \pm 0.79$  &  $70.65 \pm 1.04$  &  $\vsecondSig{71.75} \pm 0.65$  &  $\vfirstSig{73.37} \pm 0.78$  &  $71.41 \pm 0.42$  &            $70.48 \pm 0.77$  \\
                                                  & $15\%$ &  $71.07 \pm 0.36$  &  $72.79 \pm 0.38$  &  $74.10 \pm 1.07$  &  $\vsecondSig{74.47} \pm 0.56$  &  $\vfirstSig{75.29} \pm 0.61$  &  $72.72 \pm 0.34$  &            $70.46 \pm 0.59$  \\
                                                  & $20\%$ &  $73.10 \pm 0.21$  &  $74.63 \pm 0.64$  &  $75.17 \pm 1.09$  &  $\vsecondSig{75.71} \pm 0.49$  &  $\vfirstSig{76.44} \pm 0.43$  &  $73.48 \pm 0.23$  &            $71.30 \pm 0.49$  \\
                                                  & $25\%$ &  $74.79 \pm 0.26$  &  $76.32 \pm 0.28$  &  $76.74 \pm 0.78$  &     $\vsecond{77.06} \pm 0.59$  &     $\vfirst{77.46} \pm 0.40$  &  $74.49 \pm 0.35$  &            $71.66 \pm 0.35$  \\
                                                  & time (ms)   &                           1.9 &                           0.4 &                          16.1 &              128 &                            5398 &               1894 &              4.6 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\wik{}}} & $3\%$  &  $74.84 \pm 0.38$  &           $75.12 \pm 0.63$  &            $76.45 \pm 1.50$  &           $77.94 \pm 0.24$  &  $\vsecondSig{78.06} \pm 0.35$  &  $77.97 \pm 0.17$  &  $\vfirstSig{78.76} \pm 0.17$  \\
                                                  %& $5\%$  &  $78.32 \pm 0.34$  &           $78.86 \pm 0.37$  &            $79.33 \pm 0.79$  &  $\vfirst{79.67} \pm 0.48$  &     $\vsecond{79.42} \pm 0.60$  &  $78.78 \pm 0.25$  &              $78.83 \pm 0.17$  \\
                                                  %& $7\%$  &  $80.22 \pm 0.28$  &           $80.55 \pm 0.45$  &   $\vfirst{80.86} \pm 0.78$  &           $80.44 \pm 0.90$  &     $\vsecond{80.68} \pm 0.30$  &  $79.60 \pm 0.22$  &              $79.29 \pm 0.31$  \\
                                                  & $9\%$  &  $81.32 \pm 0.18$  &  $\vfirst{81.98} \pm 0.38$  &  $\vsecond{81.62} \pm 0.84$  &           $81.17 \pm 0.31$  &               $81.61 \pm 0.39$  &  $80.15 \pm 0.14$  &              $79.74 \pm 0.22$  \\
                                                  & $15\%$ &  $83.27 \pm 0.14$  &  $\vfirst{83.78} \pm 0.14$  &            $83.30 \pm 0.44$  &           $82.91 \pm 0.41$  &     $\vsecond{83.47} \pm 0.34$  &  $81.09 \pm 0.21$  &              $80.66 \pm 0.30$  \\
                                                  & $20\%$ &  $84.13 \pm 0.19$  &  $\vfirst{84.60} \pm 0.17$  &            $84.36 \pm 0.22$  &           $83.84 \pm 0.38$  &     $\vsecond{84.39} \pm 0.25$  &  $81.73 \pm 0.16$  &              $81.35 \pm 0.18$  \\
                                                  & $25\%$ &  $84.63 \pm 0.13$  &  $\vfirst{85.09} \pm 0.18$  &            $84.63 \pm 0.28$  &           $84.44 \pm 0.28$  &     $\vsecond{84.96} \pm 0.21$  &  $82.08 \pm 0.16$  &              $81.92 \pm 0.13$  \\
                                                  & time (ms)   &                           4.2 &              1.1 &                           35.3 &              210 &              14089 &              4858 &                                                   10.6\\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\sla{}}} & $3\%$  &  $69.06 \pm 0.19$  &  $69.39 \pm 0.37$  &               $74.01 \pm 0.84$  &  $\vfirstSig{80.74} \pm 0.21$  &               $78.33 \pm 0.14$  &  $\vsecondSig{79.90} \pm 0.17$  &  $74.59 \pm 0.58$  \\
                                                  %& $5\%$  &  $72.92 \pm 0.16$  &  $73.34 \pm 0.32$  &               $76.68 \pm 0.76$  &  $\vfirstSig{81.82} \pm 0.29$  &               $79.41 \pm 0.10$  &  $\vsecondSig{80.38} \pm 0.11$  &  $76.57 \pm 0.54$  \\
                                                  %& $7\%$  &  $75.11 \pm 0.10$  &  $75.58 \pm 0.19$  &               $78.95 \pm 0.36$  &  $\vfirstSig{82.34} \pm 0.30$  &               $79.99 \pm 0.10$  &  $\vsecondSig{80.59} \pm 0.07$  &  $78.03 \pm 0.55$  \\
                                                  & $9\%$  &  $76.58 \pm 0.04$  &  $77.06 \pm 0.21$  &               $79.91 \pm 0.40$  &  $\vfirstSig{82.65} \pm 0.20$  &               $80.64 \pm 0.11$  &  $\vsecondSig{80.82} \pm 0.04$  &  $79.58 \pm 0.60$  \\
                                                  & $15\%$ &  $79.16 \pm 0.08$  &  $79.68 \pm 0.16$  &               $81.95 \pm 0.23$  &  $\vfirstSig{83.12} \pm 0.20$  &  $\vsecondSig{82.02} \pm 0.09$  &               $81.16 \pm 0.09$  &  $81.71 \pm 0.40$  \\
                                                  & $20\%$ &  $80.29 \pm 0.12$  &  $80.99 \pm 0.11$  &  $\vsecondSig{82.89} \pm 0.14$  &  $\vfirstSig{83.26} \pm 0.09$  &               $82.82 \pm 0.09$  &               $81.64 \pm 0.14$  &  $82.73 \pm 0.35$  \\
                                                  & $25\%$ &  $81.08 \pm 0.08$  &  $81.86 \pm 0.05$  &  $\vsecondSig{83.45} \pm 0.10$  &  $\vfirstSig{83.66} \pm 0.14$  &               $83.39 \pm 0.08$  &               $81.94 \pm 0.15$  &  $83.42 \pm 0.16$  \\
& time (ms)   &              21.2 &              5.7 &                           655 &                            1918 &              77042 &              56252 &              78 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\epi{}}} & $3\%$  &  $80.23 \pm 0.19$  &  $80.44 \pm 0.17$  &               $86.56 \pm 0.30$  &  $\vfirstSig{89.44} \pm 0.10$  &               $86.73 \pm 0.13$  &  $\vsecondSig{86.98} \pm 0.08$  &  $86.37 \pm 0.65$  \\
                                                  %& $5\%$  &  $83.60 \pm 0.14$  &  $83.96 \pm 0.10$  &  $\vsecondSig{88.09} \pm 0.25$  &  $\vfirstSig{90.28} \pm 0.08$  &               $87.59 \pm 0.10$  &               $87.35 \pm 0.06$  &  $86.64 \pm 0.66$  \\
                                                  %& $7\%$  &  $85.24 \pm 0.10$  &  $85.69 \pm 0.12$  &  $\vsecondSig{88.96} \pm 0.16$  &  $\vfirstSig{90.70} \pm 0.08$  &               $88.45 \pm 0.07$  &               $87.63 \pm 0.04$  &  $86.55 \pm 0.37$  \\
                                                  & $9\%$  &  $86.43 \pm 0.07$  &  $86.83 \pm 0.06$  &  $\vsecondSig{89.47} \pm 0.10$  &  $\vfirstSig{91.07} \pm 0.09$  &               $89.10 \pm 0.07$  &               $87.92 \pm 0.10$  &  $87.20 \pm 0.65$  \\
                                                  & $15\%$ &  $88.33 \pm 0.06$  &  $88.76 \pm 0.06$  &               $90.49 \pm 0.06$  &  $\vfirstSig{91.56} \pm 0.05$  &  $\vsecondSig{90.53} \pm 0.10$  &               $88.58 \pm 0.15$  &  $88.10 \pm 0.56$  \\
                                                  & $20\%$ &  $89.19 \pm 0.06$  &  $89.65 \pm 0.06$  &               $90.93 \pm 0.05$  &  $\vfirstSig{91.86} \pm 0.04$  &  $\vsecondSig{91.38} \pm 0.05$  &               $88.96 \pm 0.15$  &  $88.80 \pm 0.58$  \\
                                                  & $25\%$ &  $89.76 \pm 0.06$  &  $90.21 \pm 0.05$  &               $91.75 \pm 0.05$  &  $\vfirstSig{92.07} \pm 0.08$  &  $\vsecondSig{91.95} \pm 0.07$  &               $89.42 \pm 0.20$  &  $89.37 \pm 0.63$  \\
                                                  & time (ms)   &              32 &              7.1 &                            1226 &                           2341 &             116838 &             121530 &              129 \\
\midrule
\multirow{7}{*}{\rotatebox[origin=c]{90}{\kiw{}}} & $3\%$  &  $73.91 \pm 0.37$  &  $73.93 \pm 0.58$  &  $80.44 \pm 1.05$  &  $86.95 \pm 0.19$  &  $\vsecondSig{87.72} \pm 0.32$  &               $87.02 \pm 0.05$  &   $\vfirstSig{87.88} \pm 0.01$  \\
                                                  %& $5\%$  &  $77.89 \pm 0.21$  &  $78.22 \pm 0.20$  &  $82.32 \pm 0.46$  &  $87.65 \pm 0.08$  &   $\vfirstSig{88.40} \pm 0.22$  &               $87.24 \pm 0.05$  &  $\vsecondSig{87.90} \pm 0.01$  \\
                                                  %& $7\%$  &  $79.65 \pm 0.20$  &  $80.34 \pm 0.15$  &  $83.43 \pm 0.23$  &  $86.72 \pm 0.73$  &   $\vfirstSig{88.71} \pm 0.13$  &               $87.37 \pm 0.05$  &  $\vsecondSig{87.53} \pm 0.35$  \\
                                                  & $9\%$  &  $80.87 \pm 0.09$  &  $81.61 \pm 0.09$  &  $83.97 \pm 0.27$  &  $86.53 \pm 0.20$  &   $\vfirstSig{89.02} \pm 0.07$  &  $\vsecondSig{87.49} \pm 0.07$  &               $87.23 \pm 0.38$  \\
                                                  & $15\%$ &  $82.57 \pm 0.07$  &  $83.38 \pm 0.06$  &  $85.03 \pm 0.11$  &  $86.97 \pm 0.29$  &   $\vfirstSig{89.67} \pm 0.05$  &  $\vsecondSig{87.76} \pm 0.04$  &               $87.12 \pm 0.33$  \\
                                                  & $20\%$ &  $83.34 \pm 0.06$  &  $84.11 \pm 0.09$  &  $85.48 \pm 0.13$  &  $87.02 \pm 0.24$  &   $\vfirstSig{90.00} \pm 0.07$  &  $\vsecondSig{87.92} \pm 0.07$  &               $87.12 \pm 0.21$  \\
                                                  & $25\%$ &  $83.82 \pm 0.08$  &  $84.56 \pm 0.06$  &  $85.81 \pm 0.11$  &  $87.02 \pm 0.16$  &   $\vfirstSig{90.29} \pm 0.04$  &  $\vsecondSig{88.15} \pm 0.07$  &               $87.24 \pm 0.20$  \\
                                                  & time (ms)   &                           28 &                           6.6 &                           824 &               2938 &                         103264 &             130036 &              104 \\
\bottomrule
\end{longtable}}
********************************************************
\fi

After training on the two features $1-\htr(u)$ and $1-\hun(v)$, \uslogregp{} has learned three
weights $w_0$, $w_1$ and $w_2$, which allow to predict $y_{u,v}$ according to 
\[ 
  \sgn\Big(\big(w_1(1-\htr(u)\big) + w_2 \big(1-\hun(v)\big) + w_0\Big)~.
\]
This can be rewritten as 
\[
\sgn\Big(\big(1-\htr(u)\big) + w_2'\big(1-\hun(v)\big) - \tfrac{1}{2} -\tau'\Big)~,
\]
with $w_2' = \frac{w_2}{w_1}$ and $\tau' = - \left(\frac{1}{2} + \frac{w_0}{w_1}\right)$~.

As shown in \autoref{tab:coeff}, and in accordance with the predictor built out of Equation
\eqref{eq:predictor}, $w_2'$ is almost $1$ on \emph{all datasets}, while $\tau'$ tends to be always
close the fraction of positive edges in the dataset.

\begin{table}[bt]
  \centering
  \caption{Normalized logistic regression coefficients averaged over 12 runs (with one standard
  deviation) \label{tab:coeff}}
  \small
  \begin{tabular}{lrccc}
    \toprule
    & $\frac{|\trainset{}|}{|E|}$ &              $w_2'$ & $\tau'$ \\
    % \multirow{5}{*}{\aut{}} & $3\%$  &  $0.909 \pm 0.07$  &  $0.592 \pm 0.04$  \\
    %                         & $9\%$  &  $0.985 \pm 0.04$  &  $0.707 \pm 0.03$  \\
    \multirow{5}{*}{\aut{}} & $5\%$  &  $0.965 \pm 0.04$  &  $0.662 \pm 0.03$  \\
                            & $10\%$ &  $0.983 \pm 0.03$  &  $0.705 \pm 0.02$  \\
                            & $15\%$ &  $1.001 \pm 0.03$  &  $0.729 \pm 0.03$  \\
                            & $20\%$ &  $1.013 \pm 0.02$  &  $0.747 \pm 0.02$  \\
                            & $25\%$ &  $1.011 \pm 0.02$  &  $0.746 \pm 0.01$  \\
    \midrule
    % \multirow{5}{*}{\wik{}} & $3\%$  &  $0.919 \pm 0.04$  &  $0.666 \pm 0.02$  \\
    %                         & $9\%$  &  $0.940 \pm 0.03$  &  $0.726 \pm 0.01$  \\
    \multirow{5}{*}{\wik{}} & $5\%$  &  $0.920 \pm 0.02$  &  $0.691 \pm 0.02$  \\
                            & $10\%$ &  $0.940 \pm 0.01$  &  $0.730 \pm 0.01$  \\
                            & $15\%$ &  $0.947 \pm 0.01$  &  $0.741 \pm 0.01$  \\
                            & $20\%$ &  $0.963 \pm 0.01$  &  $0.760 \pm 0.01$  \\
                            & $25\%$ &  $0.962 \pm 0.02$  &  $0.764 \pm 0.01$  \\
    \midrule
    % \multirow{5}{*}{\sla{}} & $3\%$  &  $1.012 \pm 0.02$  &  $0.669 \pm 0.01$  \\
    %                         & $9\%$  &  $1.014 \pm 0.01$  &  $0.703 \pm 0.01$  \\
    \multirow{5}{*}{\sla{}} & $5\%$  &  $1.024 \pm 0.02$  &  $0.693 \pm 0.01$  \\
                            & $10\%$ &  $1.017 \pm 0.01$  &  $0.705 \pm 0.01$  \\
                            & $15\%$ &  $1.007 \pm 0.01$  &  $0.707 \pm 0.01$  \\
                            & $20\%$ &  $1.002 \pm 0.00$  &  $0.710 \pm 0.00$  \\
                            & $25\%$ &  $0.995 \pm 0.01$  &  $0.712 \pm 0.00$  \\
    \midrule
    % \multirow{5}{*}{\epi{}} & $3\%$  &  $1.124 \pm 0.02$  &  $0.786 \pm 0.02$  \\
    %                         & $9\%$  &  $1.066 \pm 0.01$  &  $0.783 \pm 0.01$  \\
    \multirow{5}{*}{\epi{}} & $5\%$  &  $1.099 \pm 0.02$  &  $0.791 \pm 0.02$  \\
                            & $10\%$ &  $1.059 \pm 0.01$  &  $0.782 \pm 0.01$  \\
                            & $15\%$ &  $1.037 \pm 0.01$  &  $0.774 \pm 0.01$  \\
                            & $20\%$ &  $1.018 \pm 0.01$  &  $0.765 \pm 0.01$  \\
                            & $25\%$ &  $1.010 \pm 0.01$  &  $0.763 \pm 0.01$  \\
    \midrule
    % \multirow{5}{*}{\kiw{}} & $3\%$  &  $1.075 \pm 0.02$  &  $0.851 \pm 0.02$  \\
    %                         & $9\%$  &  $1.041 \pm 0.01$  &  $0.872 \pm 0.01$  \\
    \multirow{5}{*}{\kiw{}} & $5\%$  &  $1.047 \pm 0.02$  &  $0.853 \pm 0.01$  \\
                            & $10\%$ &  $1.038 \pm 0.01$  &  $0.872 \pm 0.01$  \\
                            & $15\%$ &  $1.025 \pm 0.01$  &  $0.876 \pm 0.01$  \\
                            & $20\%$ &  $1.012 \pm 0.01$  &  $0.874 \pm 0.01$  \\
                            & $25\%$ &  $1.007 \pm 0.01$  &  $0.874 \pm 0.01$  \\
    \bottomrule
  \end{tabular}
\end{table}