### 2017-09-05 assets/tikz/troll_notations_tikz.tex
move and add figures in the troll notations part

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
  \begin{tikzpicture}[auto,vertex/.append style={minimum width=10mm,font=\Large,fill=White},
        scale=1.0,]
      \tikzset{>=latex}
      \tikzstyle{edge}=[line width=2pt,color=Black,-{Latex[length=3mm,width=1.0mm,angle'=40]}]
      \tikzstyle{trainEdge}=[line width=10pt,color=Black!15,opacity=.7]
      \tikzstyle{elabel}=[color=Black]
      \tikzstyle{posc}=[scale=1.2,text=Green]
      \tikzstyle{negc}=[scale=1.2,text=red]
      \tikzstyle{testset}=[opacity=1.0]
      \tikzstyle{revealed}=[circle,draw=none,fill=Black!15,inner sep=1pt,]

      \node[vertex] (u) at (0,0) {\LARGE $u$};
      \node[vertex] (v) at (5.5,-1.26) {\LARGE $v$};

      \draw[trainEdge] (u) --                                                       (5, -2.5);
      \draw[edge] (u) -- node [revealed,posc,xshift=-2mm,] {$\mathbf{+}$}           (5, -2.5);

      \draw[testset,edge] (u) edge  node [posc,xshift=-5mm]
        {${\textcolor{Black}{y_{u,v}=}}{\textcolor{Green}{+}}$} (v);
      \draw[testset,edge] (u) edge  node [posc,xshift=2mm,pos=0.7] {$\mathbf{+}$}            (5, 0.66);
      \draw[testset,edge, dashed] (u)   edge   node [negc,xshift=2mm] {$\mathbf{-}$} (5, 2);

      \begin{scope}[on background layer]
        \draw[trainEdge] (-5, -2) --    (u.south west);
        \draw[edge, dashed] (-5, -2)       -- node[revealed,negc,yshift=-3mm,xshift=-5pt]
          {$\mathbf{-}$} (u.south west);
        \draw[trainEdge,]          (-5, 2) -- (u.north west);
        \draw[edge,]               (-5, 2) -- node[revealed,posc,above,pos=0.2,xshift=5pt] {$\mathbf{+}$} (u.north west);
      \end{scope}

      \draw[testset,edge,dashed] (-5, -1)-- node[negc,yshift=-1mm] {$\mathbf{-}$} (u);
      \draw[testset,edge,dashed] (-5, 0) -- node[negc] {$\mathbf{-}$} (u);
      \draw[testset,edge,]       (-5, 1) -- node[posc,xshift=-4mm] {$\mathbf{+}$} (u);

      \node[fill=white!90!DarkOrange,fill opacity=.9,text opacity=1,rounded corners,inner sep=2pt] at (1,-1.5) {$tr(u)=\frac{1}{4}$};
      \node[fill=white!90!Blue,fill opacity=.5,text opacity=1,rounded corners,inner sep=2pt] at (-1.0,1.7) {$un(u)=\frac{3}{5}$};
\end{tikzpicture}
\end{document}