### 2017-09-05 troll/related.tex

\section{Related work}

Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} with its weak
version~\autocite{davis1967clustering}.  The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. Among the several approaches
related to our work, some extend the spectral properties of a graph to the signed case in order to
find good embeddings for classification~\autocites{Kunegis2009}{SignedEmbedding15}. However, the
use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
makes those methods hardly scalable to large graphs. Another approach is based on mining ego
networks with SVM. Although this method seems to deliver good results~\autocite{Papaoikonomou2014},
the running time makes it often impractical for large real-world datasets. An alternative approach,
based on local features only and proposed in~\autocite{Leskovec2010}, relies on the so-called
status theory for directed graphs~\autocite{guha2004propagation}. Some works in active learning,
using a more sophisticated bias based on the correlation clustering (CC)
index~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}, provide strong theoretical guarantees.
However, the bias used there is rather strong, since it assumes the existence of a $2$-clustering
of the nodes with a small CC index.\Todo{don't put it that way since we use that bias later}

Whereas our focus will be on \emph{binary} prediction, researchers have also considered a weighted
version of the problem, where edges measure the amount of trust or distrust between two
users~\autocites{guha2004propagation}{Tang2013}{Bachi2012}{Qian2014sn}. Other works have also
considered versions of the problem where side information related to the network is available to
the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
Epinion in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
analysing the textual content of their post, and~\autocite{SNTransfer13} uses SVM to perform
transfer learning from one network to another. While many of these approaches have interesting
performances, they often require extra information which is not always available (or reliable) and,
in addition, may face severe scaling issues.

The survey~\autocite{Tang2015a} contains pointers to many papers on edge sign prediction for
signed networks, especially in the Data Mining area.

The vast majority of existing edge sign prediction algorithms for
directed graphs are based on the computation of local features of the graph. These features are
evaluated on the subgraph induced by the training edges, and the resulting values are used to train
a supervised classification algorithm (e.g., logistic regression). The most basic set of local
features used to classify a given edge $(u,v)$ are defined by
$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$. In turn, these can be used to define
more complicated features, such as
\(
	\frac{\din^+(v) + |E^+|\uin(v)}{\din(v) + \uin(v)}
\quad\text{and}\quad
	\frac{\dout^+(u) + p^+\uout(u)}{\dout(u) + \uout(u)}
\)
introduced in~\autocite{Bayesian15}, together with their negative counterparts, where $|E^+|$ is the
overall fraction of positive edges, and $\uin(v),\uout(u)$ are, respectively, the number of test
edges outgoing from $u$ and the number of test edges incoming to $v$.
Other types of features are derived from social status theory~(e.g., \autocite{Leskovec2010}), and
involve the so-called triads; namely, the triangles formed by $(u,v)$ together with $(u,w)$ and
$(w,v)$ for any $w \in \NNout(u) \cap \NNin(v)$. A third group of features is based on node ranking
scores. These scores are computed using a variery of methods, including
Prestige~\autocite{zolfaghar2010mining}, exponential ranking~\autocite{traag2010exponential},
PageTrust~\autocite{de2008pagetrust}, Bias and Deserve~\autocite{mishra2011finding},
TrollTrust~\autocite{wu2016troll}, and generalizations of PageRank and HITS to signed
networks~\autocite{shahriari2014ranking}. Examples of features using such scores are \textsl{reputation}
and \textsl{optimism}~\autocite{shahriari2014ranking}, defined for a node $u$ by
\(
	\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}
\quad\text{and}\quad
	\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}\,,
\)
where $\sigma(v)$ is the ranking score assigned to node $v$. Some of these algorithms will be used
as representative competitors in our experimental study of \autoref{s:exp}.

\iffalse
We compare these methods with four competitors, each exploiting the sign information and network
topology in a different way:

Looking at undirected graphs with a global perspective, \textcite{LowRankCompletion14} consider the
observed adjacency matrix $A$, made of the edges in \trainset{}, as a noisy sampling of the
adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly balance condition
(that is with no triangle containing only one negative edge).
%TODO the previous sentence is too long
This condition implies the existence of a small number $k$ of node clusters with positive links
within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
recovering a complete matrix $\tilde{A}$ which matches the non-zeros entries of $A$, it is possible
to predict the sign of $\eij \notin \trainset$ as $\yhat_{i,j} = \sgn(\tilde{A}_{i,j})$. Although
the exact version of this problem is NP-hard, authors assume that $k$ is an hyperparameter known
beforehand\footnote{The method is showed to be robust to the choice of its value, thus we pick
$k=7$.} and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ which minimise a sigmoidal
reconstruction loss with $A$, subject to a regularization term. We refer to this method as
\emph{\complowrank{}}.

The status theory heuristic posits that a positive link from $i$ to $j$ denotes that user $i$
considers user $j$ as having a higher status or skill than himself~\autocite{Leskovec2010}. It
implies that among the 16 possible triads of three nodes with two directed signed edges among them,
some must be more represented than others. The \emph{\comptriads{}} method exploits this fact by
counting for each edge in the training set how frequently it is involved in each of the 16 triad
types. Based on these edge features and 7 degree features of each edge endpoints, a Logistic
Regression model is trained and used to predict the sign of the test edges.

Inspired by the PageRank algorithm, \textcite{wu2016troll} propose a recursive definition of
trollness based on the opinion of one node's neighbors weighted by their own trollness, which allow
to assign a trustworthiness $\pi(i)$ to each node $i$ through a set of non linear equations solved
by an iterative method. This $\pi$ values are used to rank the nodes by computing their Reputation
and Optimism scores, thus providing four features for each edge, which are in turn used to train a
Logistic Regression model for the classification task. We refer to this method as
\emph{\compranknodes{}}.  Computing $\pi$ requires to choose two hyperparameters of the model,
$\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
perform cross validation but in practice we found little difference in performance in our
implementation while it consumes a lot of time.

The last competitors is also local. \textcite{Bayesian15} note that  a node can belong to one of the
16 types based of whether is number of its positive (resp.  negative) outgoing (resp. incoming)
edges is zero or not.  The number of unobserved incoming and outgoing edges of each node $i$ let us
define its 16-dimensional vector $V_i$ containing the probability of transitioning to any other
state.  Then for each edge \eij we compute the outer product of $V_i$ with $V_j$ and also include
additional features such as triads count and degree information before training a Logistic
Regression model. We refer to this method as \emph{\compbayesian{}}.
\fi