### 2017-09-14 packages.tex
add new dataset, competitor and results

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{csquotes}
\usepackage{babel}
\usepackage[heightrounded]{geometry}
\usepackage{setspace}
\usepackage{xspace}
\usepackage{microtype}
\usepackage[citestyle=alphabetic, bibstyle=ieee-alphabetic, isbn=false, maxnames=1, minnames=1,
sorting=nyvt, backref=true, backend=biber, defernumbers=true,maxbibnames=60]{biblatex}
\renewcommand*{\bibfont}{\normalfont\footnotesize}
\AtEveryBibitem{
   % \clearfield{arxivId}
   % \clearfield{booktitle}
   % \clearfield{doi}
   % \clearfield{eprint}
   % \clearfield{eventdate}
   \clearfield{isbn}
   \clearfield{issn}
   % \clearfield{journaltitle}
   \clearfield{month}
   % \clearfield{number}
   % \clearfield{pages}
   \clearfield{series}
   \clearfield{url}
   \clearfield{urldate}
   \clearfield{venue}
   % \clearfield{volume}
   % \clearlist{location} % alias to field 'address'
   \clearlist{publisher}
   \clearname{editor}
}
\addbibresource{/home/orphee/data/projects/biblio/library.bib}
\addbibresource{more.bib}
\addbibresource{gtx.bib}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{import}
\usepackage{mathtools}
\usepackage{sidenotes}
\let\bottomfootnote\footnote
\renewcommand{\footnote}[1]{\sidenote{\footnotesize #1}}
\usepackage[shortlabels]{enumitem}
\usepackage{tabulary}
\usepackage{bm}
\usepackage{multirow}
\usepackage[marginpar]{todo}
\usepackage[np,autolanguage]{numprint}
\usepackage{listings}
\lstset{frame=none, language=Python, tabsize=4,}
% \usepackage{inconsolata}
\usepackage{nicefrac}
\input{drawing}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\newcommand*\Let[2]{\State #1 $\gets$ #2}
\def\algorithmautorefname{Algorithm}

\usepackage{mdframed}
\newenvironment{aside}%
{\begin{mdframed}[hidealllines=true, leftline=true, innertopmargin=0.5ex, innerbottommargin=0.5ex,
      innermargin=2.3em, leftmargin=2.3em, outermargin=0em, innerleftmargin=0.7em, innerrightmargin=0pt,
      linewidth=1.5pt, skipabove=\baselineskip, skipbelow=\baselineskip,
   linecolor=DodgerBlue!30!White]\small}%
{\end{mdframed}}
\usepackage{stmaryrd}

\usepackage{varioref}
\usepackage{hyperref}
\hypersetup{%
    % draft,    % = no hyperlinking at all (useful in b/w printouts)
    colorlinks=true, linktocpage=true, pdfstartpage=3, pdfstartview=FitV,%
    % uncomment the following line if you want to have black links (e.g., for printing)
    %colorlinks=false, linktocpage=false, pdfborder={0 0 0}, pdfstartpage=3, pdfstartview=FitV,%
    breaklinks=true, pdfpagemode=UseNone, pageanchor=true, pdfpagemode=UseOutlines,%
    plainpages=false, bookmarksnumbered, bookmarksopen=true, bookmarksopenlevel=1,%
    hypertexnames=true, pdfhighlight=/O,%nesting=true,%frenchlinks,%
    urlcolor=Chocolate, linkcolor=RoyalBlue, citecolor=LimeGreen, %pagecolor=RoyalBlue,%
}
\newcommand{\marginpars}[1]{\marginpar{\small#1}}
\usepackage{caption}
\usepackage[margin=0pt,labelformat=parens,labelsep=space,%font+=small,
skip=6pt,list=false,hypcap=false]{subcaption}
\captionsetup{compatibility=false}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{makecell}
\renewcommand\theadfont{\small}
\graphicspath{{./assets/}}
% \usepackage[capitalize,noabbrev]{cleveref}
\usepackage[]{extdash}