### 2017-09-14 troll/diff_mcc_pp.tex

\begin{table}[htpb]
  \centering
  \caption{100 time the difference of MCC between using the twin heuristic and not using it for our
  two main methods. Those values are computed on the same training/testing split as of
\autoref{tab:all_mcc}.\label{tab:twin}}
  \begin{tabular}{lrrrrrrr}
    \toprule
                                & $5\%$                    & $10\%$                   & $15\%$                   & $20\%$                   & $25\%$                   & $50\%$                   & $90\%$                   \\
    \midrule
    \multicolumn{8}{c}{\aut{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.7$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Green}{$0.1$} & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.7$}  \\
    \multicolumn{8}{c}{\adv{}} \\
    \uslpropGsecTwin{}          & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.6$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.4$}  & \textcolor{Red}{$-10.8$} \\
    \usruleTwin{}               & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.5$}  & \textcolor{Red}{$-1.0$}  & \textcolor{Red}{$-1.6$}  & \textcolor{Red}{$-2.2$}  & \textcolor{Red}{$-5.7$}  & \textcolor{Red}{$-11.1$} \\
    \multicolumn{8}{c}{\wik{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.2$}  \\
    \multicolumn{8}{c}{\sla{}} \\
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$2.0$} & \textcolor{Green}{$3.3$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.9$} & \textcolor{Green}{$1.1$} & \textcolor{Green}{$1.3$} & \textcolor{Green}{$2.2$} & \textcolor{Green}{$3.5$} \\
    \multicolumn{8}{c}{\epi{}} \\
    \uslpropGsecTwin{}          & \textcolor{Green}{$0.3$} & \textcolor{Green}{$0.5$} & \textcolor{Green}{$0.7$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.9$} & \textcolor{Green}{$3.0$} \\
    \usruleTwin{}               & \textcolor{Green}{$0.4$} & \textcolor{Green}{$0.6$} & \textcolor{Green}{$0.8$} & \textcolor{Green}{$1.0$} & \textcolor{Green}{$1.2$} & \textcolor{Green}{$1.8$} & \textcolor{Green}{$2.8$} \\
    \multicolumn{8}{c}{\kiw{}} \\
    \uslpropGsecTwin{}          & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.1$}  & \textcolor{Red}{$-0.3$}  \\
    \usruleTwin{}               & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & $0.0$                    & \textcolor{Red}{$-0.2$}  & \textcolor{Red}{$-0.5$}  \\
    \bottomrule
  \end{tabular}
\end{table}