### 2017-09-29 intro/draft.tex
first finished version of the edspi report

Graphs are useful to represent relationships between entities. In particular, a lot can be learned
from graphs encompassing several types of relations in one integrated structure. However, many
existing works assume that such graphs are given, before studying interesting properties like
community structure and information spreading. In this thesis instead, we develop methods to predict
the type of edges given a graph topology, a few initial label and possibly attributes of the nodes.
\smallskip

\paragraph{Learning on graphs}

Graph theory has been around for almost three centuries~\autocite{biggs1976graph}, and it provides a
conceptually simple yet immensely rich framework to model situations where entities are connected
with each other~\autocite{ComplexNetworksApp11}. Coupled with the increasing availability of large
amount of relational data, it has recently spurred a lot of interest across various and tangible
applications:
\begin{itemize}[nosep,leftmargin=*]
  \item community detection~\autocite{FortunatoSurvey10}, where the goal is to cluster nodes in
    tightly connected group that are loosely connected with the rest of the graph, in order to
    better understand their relationships and the graph organization. Is has been used for instance
    to identify proteins functional groups~\autocite{clusterBio03} or see how different scientific
    fields relate based on publication data~\autocite{clusterScience08}.
  \item semi-supervised learning~\autocites{SSL06}{graphSSL14}, where during training, we are
    provided with unlabeled data at the same time as labeled data. Connecting similar instances
    allow propagating information along the graph, with applications in classifying text
    documents~\autocite{sslText09} or aligning categories and relations across multiple knowledge
    base~\autocite{sslKB13}.
  % \item node classification~\autocite{nodeClassif11} Idem
    % could add visualisation (Roberto Tamassia. 2013. Handbook of Graph Drawing and Visualization.
    % CRC Press.) but it's not exactly learning
  \item node embedding~\autocite{representationLearning17}, where we seek a low dimensional
    representation of nodes based on their structural patterns, usually in an unsupervised way,
    although it is also possible to include problem supervision when available. Such representation can
    then be used in downstream tasks, for instance visualisation~\autocite{LINE15} or the
    aforementioned node classification, even when new nodes join the graph after
    training~\autocite{inductiveRepresentation17}.
  \item link prediction~\autocite{linkPredSurvey16}, where the goal is, given a snapshot of the
    graph at time $t$, to return a set of non-existing links that will be created by time
    $t+\Delta$. Most methods are based on the assumption that link creation is driven by node
    similarity and it has been successfully applied to inferring potential interaction between
    proteins without expensive experiments~\autocite{linkPredBio06} and uncovering hidden
    association in criminal networks~\autocite{linkPredCrime08}.
  \item information and influence propagation~\autocite{infmax13}, which is the study of processes
    by which content is spread across networks, and how such processes can be influenced to speed
    them up or slow them down. Two main applications are selecting the best seed in a social
    network to promote a viral marketing campaign~\autocite{infmaxKempe15} and better contain the
    diffusion of actual biological viruses~\autocite{influenceBio13}.
  \item network evolution~\autocite{networkEvolution14}, which focuses on the mechanisms and
    consequences of the growth of most networks, as well as how to keep the results on some data
    mining algorithms \enquote{fresh} and relevant. Monitoring the changes in the interactions of
    proteins can indeed be used as an early indicator for some kind of
    diseases~\autocite{evolBio10}, while sudden changes in networks of computers are usually worth
    investigating for they might signal some external attack~\autocite{evolSecurity04}.
\end{itemize}

\vspace{-\baselineskip}
\paragraph{Complex networks}

The prevalent setting for most of the works discussed above is that the graph connections are: (i)
homogeneous, in the sense that all edges are of a single type, having the same semantics  between
any pairs of nodes; and (ii) based on the \emph{homophily} principle~\autocite{Homophily01}, roughly
meaning that nodes are linked because they are similar. However, many real world situations require
a more complex model, with several kind of edges and where nodes can be connected if they are
similar only on some dimension or even if they are dissimilar. For instance, a e-commerce website
could allow its users to express whether they find the review of other users trustworthy or not. As
more involved example, consider a social network such as Facebook. Whereas the website offers only a
single \emph{friendship} relation, we may want to distinguish between family, colleague and sport
buddies~\autocite{egoNips12}. Moreover, among one user's \enquote{friend}, there might be some
relations that are actually not positive~\autocite{Yang2012}, whether because the receiving end if
spam or behave in strange way. Note finally that friendship is only kind of relationship between two
users of Facebook, we could also link them whenever they attend the same event or are tagged in the
same post. Yet another possibility would be to study the same set of users across multiple social
networks~\autocite{mergingNetworks16}, in which case the edge type could denote some common
combinations of connections (\eg Twitter+LinkedIn or Facebook+Instagram). In this thesis, we thus
defend the following statement: \textbf{There exists efficient and accurate methods to predict edge
type in complex networks (reyling either only on the graph topology or also on node metadata).}

A first motivation is that by labeling all edges, we can split a complex networks in a superposition
of simple networks with a single edge type, on which we can apply existing methods. Of course,
different kind of relation are intertwined, thus a second stronger motivation is to identify the
different layers in order to study their interactions.

\vspace{-\baselineskip}
\paragraph{Predicting polarity in directed social networks}

The first kind of such complex networks we consider are \emph{signed graphs}~\autocite{Tang2015a},
where edges carry a positive or a negative sign. Among the many applications of signed graphs, a
natural, and historical~\autocite{harary1953}, one is directed social networks. In such systems,
users can explicitly tag others, stating whether they trust them, support them, like them, or not.
Being able to predict edge signs in such networks improve the user experience, for instance by
promoting reasonable online debates~\autocite{journalism15}, assembling efficient teams within
companies~\autocite{MLinHR16} and mitigating cyberbullying~\autocite{cyberbullying13}. Existing
approaches attack the problem along several angles: finding a low dimensional representation of the
nodes based on their connection patterns~\autocites{SIGNet17}{SNE17}, finding low rank approximation
of the adjacency matrix~\autocites{LowRankCompletion14}{OnlineCompletion17} or extracting
topological features and training classifiers~\autocites{Leskovec2010}{Bayesian15}{wu2016troll}.
Guided by the idea that users exhibit various degree of trollness, we propose a simple sign
generative model. It naturally leads to two batch algorithms, one that probably approximate the
Bayes optimal predictor when the training set is dense enough and one that approximate the maximum
likelihood estimates through label propagation. Extensive experiments on synthetic and real datasets
show that in addition to their sound theoretical foundations, our methods are competitive with the
state of the art in predictive performance, while being more scalable. To illustrate this point, a
short extract of our results is showed in \autoref{tab:troll}.

\begin{table}[ptbh]
  \centering
  \caption{Matthews Correlation Coefficient measuring (the closer to 100 the better) in a \aut{}
    network as the proportion of labeled training edges grows. The best result is in red and the
  second one in blue.\label{tab:troll}}
  \vspace{0.5\baselineskip}
  \begin{tabular}{lccccc|r}
    \toprule
    {}               & 3\%                & 9\%                & 15\%               & 20\%               & 25\%               & time (ms)            \\
    \midrule
    \uslogregp{}     & \vsecondSig{15.19} & \vsecondSig{26.46} & 32.98              & 36.57              & 39.90              & 2                    \\
    \rowcolor{gray!20!white}
    \usrule{}        & 15.09              & 26.40              & \vsecondSig{32.98} & \vsecondSig{36.72} & 40.16              & \textbf{\textless 1} \\
    \rowcolor{gray!20!white}
    \uslpropGsec{}   & \vfirstSig{19.00}  & \vfirstSig{30.25}  & \vfirstSig{35.73}  & \vfirstSig{38.53}  & \vfirstSig{41.32}  & \textbf{16}          \\
    \midrule
    \compranknodes{} & 12.28              & 24.44              & 31.03              & 34.57              & 38.26              & 128         \\
    \complowrank{}   & 8.85               & 17.08              & 22.57              & 25.57              & 29.24              & 1894                 \\
    \compbayesian{}  & 10.91              & 23.75              & 32.25              & 36.52              & \vsecondSig{40.32} & 5398                 \\
    \comptriads{}    & 8.62               & 16.42              & 22.01              & 24.77              & 27.13              & \textbf{5}           \\
    \bottomrule
  \end{tabular}
\end{table}

\vspace{-\baselineskip}
\paragraph{Active learning in undirected signed graphs}

There are many other domains where signed graphs have proved useful. For instance in computer
vision, it is used to segment images in 2D~\autocites{Kim2011}{Bagon2011}{CellSeg14} or
3D~\autocites{VolumeSegmentation12}{Beier2015}, to simplify 3D shapes~\autocite{Shape3D17} and to
track target across sequential video frames~\autocite{multiTracking15}. When the nodes are words,
signed graphs are employed to identify coreference~\autocites{graphicalCoreference04}{Elsner2009} or
cluster synonym words~\autocite{SignedWordRatings}. They plays a large role in biology, to clusters
genes~\autocite{Ben-Dor99}, identify mutation regions in chromosomes~\autocite{Das2015} or finding
stable subsystems~\autocite{monotoneBiology07}. Finally, given large databases with duplicated
records, they naturally help solving entity
resolution~\autocites{Crosslingual07}{DeDup09}{LargeScaleDeDup09}{WeightedER14}.

In all these cases, graphs are not directed and nodes are not human beings, meaning we can hardly
assign them trollness.  Therefore, we make a different assumption, positing that nodes are
partitioned in $K$ groups in such a way that links within groups are positive and links across
groups are negative. Recovering the optimal partition after observing all the signs of a graph is
called \pcc{}~\autocite{Bansal2002} in the Machine Learning field. Whereas it has been proved to be
\NPc{} to even approximate in the worst case~\autocites{Charikar2003}{Demaine2006}, our thorough
literature review paints a less gloomy landscape. Not only does there exist
exact~\autocite{Berg2015} and fixed parameters algorithms~\autocites{GoldenCE12}{Fomin2014},
efficient approximation~\autocites{CCPivotConf05}{ChawlaArxiv14} and large scale
heuristics~\autocites{Levorato2015}{Facchetti2011isingmodel}{Kappes2016}, but beyond the worst case,
we are interested in instances that exhibit stability under
perturbation~\autocites{clusteringFeasibility15}{StableCC09}{StableLP09} or are obtained through
random moves from an ideal case~\autocites{plantedAilon09}{Makarychev2014}. This leads us to develop
a spanning tree tailored to predicting edge signs in such a setting, which we call \gtx{}. It does
so by aiming to approximately preserve the distances in the
graph~\autocites{LowerBound95}{Abraham2012}. We present preliminary results showing that on certain
kind of graphs, this indeed results in better predictive performance than a competitive baseline.
Namely, we create synthetic graphs of growing size with the \lpa{} model~\autocite{Barabasi1999},
partition the nodes in two clusters, set the edge signs accordingly and finally flip a fraction of
them \uar{}. As showed on \autoref{fig:gtx_xp_pasynthmcc}, \gtx{} always predict signs better than a
\bfs{} tree rooted at the highest degree node, especially when the noise level is low, and is less
sensible to the graph size.

\begin{figure}[phtb]
  \centering
  \includegraphics[width=.6\textwidth]{gtx_exp/pasynthmcc}
  \caption{Synthetic \lpa{} }\label{fig:gtx_xp_pasynthmcc}
\end{figure}

\vspace{-\baselineskip}
\paragraph{Predicting more than two kinds of relations}

Finally, we consider the case where there are more than two possible type of relations, which is
often called \emph{multilayer graphs}~\autocites{Kivela2014}{multiSurvey14}. In addition, we assume
that the graph is attributed, that is every node is described by a vector of numeric features called
the \emph{profile} of the node. Our assumption is then that links cannot be fully described by a
global homophily. Thus we take a more nuanced view, combining partial homophily (that is, nodes are
connected when they are similar on a subset of the attributes) and heterophily (that is, nodes are
connected when they are dissimilar on a subset of the attributes). As examples of the latter, think
of dating websites ---where most users are linked with users of the opposite
gender~\autocites{homophilyMyspace09}{Tinder16}; spreading innovation ---where meeting people with
different background and point of view is crucial to favor diversity and
creativity~\autocite{rogers2003diffusion}; and online news consumption ---where connecting people
from different sides of the political spectrum helps avoiding echo chambers and fueling a democratic
debate~\autocite{balancedNews17}. Our goal is to predict to which category a given belong. This is
widely applicable, for multilayer graphs are present  in various fields. In social networks, where
players of the Pardus massively multilayer online game can have six types of relations (friendship,
enmity, private message, trading, aggression and bounty)~\autocite{Szell2010} or where photo sharers
of the Flickr website can interact in eleven ways (either directly or through comments, tags, groups
and so on)~\autocite{RecoFlickrMulti11}. In citation networks, where authors of the DBLP dataset are
connected if they have co-authored a paper in one the \np{1000} conferences that form the edge
type~\autocite{communityDBLPbyConf05}, or where \np{5000} SIAM papers can be connected for five
reasons (abstract similarity, title similarity, keywords similarity, author similarity or
citation)~\autocite{articlesMultiSim11} in economic networks, where 951 ports are connected by 3
kind of ships (bulk dry carriers, container ships and oil tankers)~\autocite{ports3kindofships10} or
in the International Trade Network, where 162 countries are connected by 96 kind of commodities they
can exchange~\autocite{worldTradeNetwork10}. And in biology, where genes are connected in a
co-expression network under 130 different experimental conditions~\autocite{bioLayerExp11}.

There exists previous work that leans towards analyzing and characterising edges in such multilayer
graphs, although they do not share exact our goal of predicting category. For instance, one can
maximize the probability of a generative model in ego networks with $K$
categories~\autocites{LeskovecEgo12}{LeskovecEgo14}, look for communities that are both tightly
connected and share common keywords~\autocite{AttributedCommunity16}, find a partition maximizing
the density of intra community edges ---weighted by the endpoints profile similarities and a learned
community weight vector~\autocite{ZhangModelFree16}, recover with low distortion $K$ metric spaces
whose small-world superposition yield the observed graph~\autocite{Abraham2012a}, discover edge role
by computing high order topological features~\autocite{ahmed2017roles}, group nodes with common
attributes and extract frequent links between such group as edge type~\autocite{conceptualLinks12},
or cluster the profiles into a set of affine subspace of $\mathbb{R}^d$~\autocite{SCSurvey11}.

Our approach is to seek a small number of vectors that, once assigned to every edge, best explain
the graph (\ie maximize a score function between the edge vector and the profiles of its endpoints).
From this initial formulation, we derive two optimization problems, and initially solve them on
synthetic data, generated with a few natural topological constraints. We compared these solutions
with the result of a baseline approach that simply clusters the pairwise profile vectors using
$k$-means. As showed in \autoref{fig:edge}, on a graph with \np{500} nodes, \np{1300} edges and 7
types of edges (where each node is involved in at most 3 types of relation), we recover the hidden
edge assignment more accurately than the $k$-means baseline.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.7\linewidth]{assets/raw/Gfixed_over0_step2_7dirs3_ambi.pdf}
  \caption{Distribution of the Adjusted Mutual Information (the closer to one the better) over
  \np{600} generation of 7 random directions, comparing our \textsf{optim} method with the $k$-means
baseline. \label{fig:edge}}
\end{figure}

% Surveys are also available about applications of unsigned networks
% such as data classification [Sen et al. 2008], recommendation [Tang et al. 2013], and
% information propagation [Chen et al. 2013a]

While our thesis is therefore primarily concerned with learning on graphs, instead of focusing at
nodes (to predict their class, their cluster or to find their representation), we look at edges (\ie
pairs of nodes), and instead of having one kind of relationship, we have two, or more than two when
the nodes are attributed.