### 2017-10-27 edge/signed.tex
2-CC version of the connection between edge and signed
The previous approaches was indeed not correct, for reasons explained
in the aside paragraph :(

\paragraph{Generalization of signed graphs}
\label{par:generalization_of_signed_graphs}

The learning bias of \autoref{chap:cc} implicitly defined an idealized \esp{} problem as:
\begin{problem}[idealized \esp{}]
  \label{p:edge_esp}
  Given $n$ nodes belonging latently to $K$ groups, predict whether two nodes $u$ and $v$ belong to
  the same group (that is $(u,v)$ is positive) or not (that is $(u,v)$ is negative).
\end{problem}
In \autoref{sec:gtx} we study in particular the important special case where $K=2$. Now consider an
instance of \autoref{p:edge_full} where $k=2$, and where the profiles and edges are based on group
membership. We show that the optimal solution of such an instance is exactly the solution of this
idealized \esp{} problem on the same graph.

Let us first define this instance formally. Given a graph $G=(V,E)$, we partition the nodes in two
groups $L$ and $R$. We then define the nodes profiles $X$ as follow. We first draw \uar{} a vector $p$
from $\{-1, 1\}^d$, then let the profiles in $L$ be $\frac{p}{\sqrt{d}}$ and the
profiles in $R$ be $-\frac{p}{\sqrt{d}}$. We call this instance $\mathcal{I} = (G, X, k=2)$. With
this choice, we have
\begin{equation*}
  s_{uv} =
  \begin{cases}
    \frac{\onev}{d} = s_{\mathrm{inner}} & \quad \text{if $u$ and $v$ are in the same group,} \\
    -\frac{\onev}{d} = s_{\mathrm{outer}} & \quad \text{if $u$ and $v$ are in different groups.}
  \end{cases}
\end{equation*}
Let us denote by $m_i$ the number of edges that are internal to $L$ and $R$, and by $m_o = |E| -
m_i$ the number of edges between $L$ and $R$.

\begin{prop}
  Let $\mathcal{D}_2 = \{w_1^\star, w_2^\star\}$ be an optimal solution of \autoref{p:edge_full} on the
  instance $\mathcal{I}$. Then $\mathcal{D}_2$ provides the following optimal solution to
  \autoref{p:edge_esp}: each every edge explained by $w_1^\star$ is predicted positive, and every
  edge explained by $w_2^\star$ is predicted negative.
\end{prop}
\begin{proof}
  Take any $w_1, w_2 \in \dsphere \times \dsphere$ such that $w_1 \neq w_2$. Assume without loss of
  generality that $w_1 = \argmax_{w\in \{w_1, w_2\}} {s_{\mathrm{inner}}}^T w$. This implies that:
  \begin{equation*}
    {s_{\mathrm{inner}}}^T w_1 = \frac{1}{d}\sum_{i=1}^d {w_1}_{;i} >
    \frac{1}{d}\sum_{i=1}^d {w_2}_{;i} = {s_{\mathrm{inner}}}^T w_2 \,.
  \end{equation*}
  In turn, we have that:
  \begin{equation*}
    {s_{\mathrm{outer}}}^T w_2 = -\frac{1}{d}\sum_{i=1}^d {w_2}_{;i} >
    -\frac{1}{d}\sum_{i=1}^d {w_1}_{;i} = {s_{\mathrm{outer}}}^T w_1 \,.
  \end{equation*}
  We can thus rewrite the maximization objective \eqref{eq:edge_full} of \autoref{p:edge_full} as:
  \begin{align}
    \label{eq:edge_special}
    \argmax_{w_1,w_2 \in \dsphere} & \quad \sum_{u,v \in E} \max_{w \in \{w_1, w_2\}} {s_{uv}}^T w = \\
    \argmax_{w_1,w_2 \in \dsphere} & \quad
    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{inner}}}} {s_{\mathrm{inner}}}^T w_1 +
    \sum_{\substack{u,v \in E\\ s_{uv} = s_{\mathrm{outer}}}} {s_{\mathrm{outer}}}^T w_2 = \notag \\
    \argmax_{w_1,w_2 \in \dsphere} & \quad m_i \frac{1}{d} \onev + m_o \frac{-1}{d} \onev \notag
  \end{align}
  One can check using Lagrange multipliers that for any $c \in \Rbb^n$, $\max_{w \in \dsphere} c^T w
  = \frac{c}{\|c\|}$. The solution of \eqref{eq:edge_special} is therefore:
  \begin{equation*}
    w_1^\star = \frac{\onev}{\sqrt{d}} \qquad \text{and} \qquad w_2^\star = \frac{-\onev}{\sqrt{d}}
  \end{equation*}
  As claimed, the edges associated with $w_1^\star$ (respectively $w_2^\star$) are the edges that
  are to be predicted positive (respectively negative) in \autoref{p:edge_esp}.
\end{proof}

\begin{aside}
  It is natural to ask how this result can be extended to more than $K=2$ groups. The most natural
  idea is to use one-hot encoding of the group membership as profiles, with additional negative
  values as in the example of \autoref{fig:edge_exe}. Namely, set the profiles of the \ith{} group
  as follow: all component are equal to $-b$ except for the \ith{} one, which is equal to $a$, where
  $a$ and $b$ are real constants such that $a^2 + (K-1)b^2=1$. However, in that case, we check
  numerically that $w_+ = \frac{\onev}{\sqrt{d}}$ and $w_- = \frac{-\onev}{\sqrt{d}}$ is the not the
  optimal solution, and that a solution achieving a larger objective function value does not cluster
  the edges between inner and outer.

  The next thing I tried was to use the approach described later in \autoref{sub:edge_generation} in
  order to optimize the profiles of a complete graphs with $K$ groups, given $w_+$, $w_-$ and the
  inner/outer edges assignment. As one could expect from the symmetry of the problem, all the nodes
  in the same group have the same profile. However, there are no clear pattern in those profile, and
  once again there exists better solutions than $\{w_+, w_-\}$, which partition the edges in two
  roughly balanced sets.
\end{aside}