### 2017-11-21 intro/main.tex
say complex networks have 2 features: k edge types and homo/hetero mix

Graphs are a natural way to represent the relationships between a set of entities. Because of the
simplicity and flexibility of this formalism, graphs are ubiquitous and have been used in countless
fields. To make the rest of our explanations more concrete, we now give three examples of graphs
that we consider later on.
\begin{enumerate}[nosep]%,leftmargin=*]
\item In a social network\footnote{In the rest of this thesis, we use the terms \emph{graph} and
\emph{network} interchangeably.}, the nodes of the graph are human users, and relationships
between nodes denote interactions between the corresponding users. One social network we study
in \autoref{chap:troll} summarizes the connections between the editors of \wik{}. These editors
can be promoted to administrators after a vote from their peers. An edge \euv{} in this network
means that user $u$ has voted on the possible promotion of $v$.
\item In computer vision, we can represent an image as a graph. Each pixel is a node, and those
nodes are connected to four neighbors, namely the adjacent pixel from the top, bottom, left and
right sides.
\item In e-commerce, we can consider a co-purchase network. Nodes are products being sold on a
website, and two nodes are connected if the corresponding products are frequently bought
together by customers, for instance a phone and a memory card.
\end{enumerate}

Not only are there graphs in many domains, but the progress of technology in the last few decades
has made it easier to collect many graphs in every single domain, sometimes with up to billion of
nodes and hundred of billions of edges. The availability of such large amount of structured data has
prompted the development of automated methods to extract insight from it. For instance, it is
possible to cluster nodes into coherent groups, predict the category of the nodes or study how to
best propagate information within a graph. We describe in more details these possibilities and
others in \autoref{sec:learning_in_graphs}. By doing so, our goal is to illustrate the wide
potential of \emph{learning in graphs}.

\medskip

At this point of the discussion though, we have only considered the most common and simple kind of
graph, one representing a single type of relation between nodes. However, in many situations, there
are two dimensions along which graphs exhibit more complexity. First, nodes have more than one type
of relation among each other. Second, two nodes are not only connected because of their global
similarity but also for more nuanced reasons. As a case in point, let us look again at our
three examples. In the \wik{} network, a vote can support the promotion or oppose it. This
additional information enrich our understanding of the relationships between the nodes in the graph.
It also points out that two editors can be connected because they share a common topic of interest,
but come from different socio-economic backgrounds and therefore cannot agree on this topic.
Likewise in an image, an edge between two pixels is positive if the two pixels belong to the same
object (say a car or a building) and negative otherwise. This information could be used to segment
the image. In the co-purchase network, there are even more than two types of relation. Assuming the
products are movies, two movies can be frequently bought together because they are part of a series
(like Star Wars), because they have the same director but different gender, because they receive the
same prestigious awards in different years, and so on. The second type of relation (\enquote{same
director} and \enquote{different gender}) is actually an example of a mixed relation, for it mixes
similarity and dissimilarity.

We call \emph{complex networks} graphs where edges have different \emph{semantics}, or types, and
where connections are explained by more than simple global similarity. In
\autoref{sec:edge_semantics}, we give additional examples of such complex networks. As
showed in our three previous examples, we distinguish between two cases. The first is when there are two types of
edge having opposite semantics. Such graphs are called \emph{signed graphs} and have been
extensively studied since the fifties, for we shall see they have many applications. The second case
is when there are more than two types of edge. Such graphs have also been studied for a long time
under different names and we refer to them as \emph{multilayer graphs}.

\medskip

It is natural to expect that we can extract finer insights from graphs with edge semantics. However,
in many cases, the information about edge types is not available. We therefore present in
\autoref{sec:predicting_edge_type} the problem of characterizing edges in complex networks.
Informally, given an input graph and possibly some extra information, we want to predict the type of
every edge. This problem can take several forms depending of what information is available as input.
First, the graph may be directed or undirected. Second, there might be two edge types (in the case of
signed graphs) or more (for general multilayer graphs). Third, the problem can be supervised or not.
In the supervised setting, we are provided with labels for some of the edges. In the unsupervised
setting, there is no label at all. We thus make the additional assumption that we observe some
\emph{attributes} of the nodes. Once again, we return to our three examples to illustrate what
those attributes can be. In the \wik{} graph, we could have for each user data about age, experience and
area of expertise. In an image, each pixel is associated with a color, along with higher order
visual features. Each product of a co-purchase network comes with information about, \eg{}, its price,
popularity, category and availability. In the absence of label, our intuition is that these
attributes can inform us about the type of edge between nodes.

\medskip

We list in \autoref{sec:outline} three concrete instantiations of this general problem that we
consider in this thesis. Our solutions to these three problems offer evidence in the defense of our
thesis statement:
\begin{quote}
\bf
There exists efficient and accurate methods to predict edge type in complex networks, relying
only on the graph topology or also on node attributes.
\end{quote}

\section{Learning in graphs}
\label{sec:learning_in_graphs}
\input{learning}

\section{Graph with several edge semantics}
\label{sec:edge_semantics}
\input{semantics}

\section{Predicting edge type}
\label{sec:predicting_edge_type}
\input{predicting}

\section{Outline}
\label{sec:outline}
\input{outline}