### 2017-11-23 intro/semantics.tex
examples of signed graphs in intro

The three graphs we described in the introduction are examples of what we call complex networks.
Such graphs have more than one type of edge, and edges are not explained by mere global node
similarity. The point of this section is to illustrate that this notion is not simply a mathematical
variation of a \enquote{simple} graph, where the adjacency matrix would take values in $\{-1, 0, 1\}$ or
$\{0, 1, \ldots, k\}$ instead of $\{0, 1\}$. Rather, we reviews many applications in various domains
where complex networks are the right model to represent rich networked data. First, we showcase many
uses of signed networks. In that case, we do not insist on the fact that similarity is not the only
driver of node connections, for it is implicit that negative edges actually denote dissimilarity.
Second, we present multilayer graphs, that is graphs with more than two types of edges. When
applicable, we highlight how the connections in such graphs indeed rely on a nuanced similarity or
dissimilarity across some of the nodes attributes.

To elaborate on this last point, our hypothesis is that connections in complex networks are the
results of both partial homophily (that is, nodes are connected when they are similar on a subset of
the attributes) and partial heterophily (that is, nodes are connected when they are dissimilar on a
subset of the attributes). As examples of the latter, think of dating websites ---where most users
are linked with users of the opposite gender~\autocites{homophilyMyspace09}{Tinder16}; diffusing
innovations ---where meeting people with different backgrounds and point of views is crucial to
favor diversity and creativity~\autocite{rogers2003diffusion}; and online news consumption ---where
connecting people from different sides of the political spectrum helps to avoid echo chambers and
fuel a democratic debate~\autocite{balancedNews17}.

\subsection{Signed graphs}
\label{sub:intro_signed_graphs}

In this section, we present a list of signed graph usages, sorted by domains. Many of those signed
graphs are the input of some clustering algorithms. In the context of signed graph, the clustering
task can be captured by the \pcc{} problem. We provide a thorough overview of this problem in
\autoref{chap:cc}. Here we simply give a broad, informal definition. The objective in \pcc{} is to
cluster the nodes of a signed graph in a way that minimizes the number of positive edges across
clusters and the number of negative edges within clusters. Such edges are called disagreements.

\input{use_of_sign_graphs.tex}