### 2017-11-25 intro/semantics.tex
example of multilayer graphs in intro

The three graphs we described in the introduction are examples of what we call complex networks.
Such graphs have more than one type of edge, and edges are not explained by mere global node
similarity. The point of this section is to illustrate that this notion is not simply a mathematical
variation of a \enquote{simple} graph, where the adjacency matrix would take values in $\{-1, 0, 1\}$ or
$\{0, 1, \ldots, k\}$ instead of $\{0, 1\}$. Rather, we reviews many applications in various domains
where complex networks are the right model to represent rich networked data. First, we showcase many
uses of signed networks. In that case, we do not insist on the fact that similarity is not the only
driver of node connections, for it is implicit that negative edges actually denote dissimilarity.
Second, we present multilayer graphs, that is graphs with more than two types of edges. When
applicable, we highlight how the connections in such graphs indeed rely on a nuanced similarity or
dissimilarity across some of the nodes attributes.

To elaborate on this last point, our hypothesis is that connections in complex networks are the
results of both partial homophily (that is, nodes are connected when they are similar on a subset of
the attributes) and partial heterophily (that is, nodes are connected when they are dissimilar on a
subset of the attributes). As examples of the latter, think of dating websites ---where most users
are linked with users of the opposite gender~\autocites{homophilyMyspace09}{Tinder16}; diffusing
innovations ---where meeting people with different backgrounds and point of views is crucial to
favor diversity and creativity~\autocite{rogers2003diffusion}; and online news consumption ---where
connecting people from different sides of the political spectrum helps to avoid echo chambers and
fuel a democratic debate~\autocite{balancedNews17}.

\subsection{Signed graphs}
\label{sub:intro_signed_graphs}

In this section, we present a list of signed graph usages, sorted by domains. Many of those signed
graphs are the input of some clustering algorithms. In the context of signed graph, the clustering
task can be captured by the \pcc{} problem. We provide a thorough overview of this problem in
\autoref{chap:cc}. Here we simply give a broad, informal definition. The objective in \pcc{} is to
cluster the nodes of a signed graph in a way that minimizes the number of positive edges across
clusters and the number of negative edges within clusters. Such edges are called disagreements.

\input{use_of_sign_graphs.tex}

\subsection{Multilayer graphs}
\label{sub:intro_multilayer_graphs}

Besides signed graphs, in this thesis we also consider at multilayer
graphs~\autocites{Kivela2014}{multiSurvey14}. Those are graphs with $k$ edge types, and the name
refer to the fact we can see them as the superposition of the $k$ subgraphs induced by each edge
type. Even when $k=2$, we make a distinction between signed graphs and multilayer graphs. In signed
graphs, the two types of edge have reverse semantic, whereas in a 2-layers graph, they simply
denotes two possible interaction, for instance advisor-advisee or regular coauthors relation in a
citation network~\autocite{Advisor10}. In general though, we focus on cases where $k$ is larger than
$2$, and not larger than a few dozen in order to preserve interpretability.
Like signed graphs, these multilayer graphs are versatile enough to be used in many fields.

\paragraph{Social networks} \Textcite{Szell2010} model the interactions of the players of
\emph{Pardus}, a massively multilayer online game. These players can be friend or enemy, send
private messages, trade resources, attack each other and set a bounty on the head of another
players. These six types of interactions are either positive or negative, but their nuances cannot
simply by explained by global similarity and dissimilarity. Another examples are the users of the
photo sharing website Flickr.  They can interact in eleven ways, either directly or through
comments, shared tags, groups membership and so on~\autocite{RecoFlickrMulti11}. Again, while being
part of a common group denote shared interests between two users, overall they must also differs in
some other attributes (for instance location) in order to bring diversity to this groups.

\paragraph{Citation networks} In these networks, nodes are research papers or authors, and edges
typically connect two nodes whenever one cites the other. For instance, using the DBLP dataset,
\textcite{communityDBLPbyConf05} connect two authors if they have co-authored a paper in one the
\np{1000} conferences appearing in the data. One can then consider that the resulting graph was
obtained as the superposition of these \np{1000} subgraphs. Besides direct citations,
\textcite{articlesMultiSim11} consider four others reasons to connect \np{5000} SIAM papers, based
on their similarity in terms of abstract words, title words, keywords and authors.

\paragraph{Economic networks} In our ever increasingly globalized economy, entities around the world
are getting more and more tightly connected. However, it would be overly simplistic to consider
there is only kind of link between entities. For something seemingly as simple as the connections
between the largest \np{951} ports, one must already noticed that these connections can be
implemented by any combinations of three kinds of ships: bulk dry carriers, container ships and oil
tankers~\autocite{ports3kindofships10}. Likewise, the 162 countries of the International Trade
Network are connected by 96 kinds of commodities they can exchange~\autocite{worldTradeNetwork10}.
Finally, \textcite{KantPeace15} studies international relations through the lens of Kant three
folded program for peace,  based on democracy, economic dependence and supra national governance.
They build the graph of all countries and connect them in three layers. All democracies form a
clique in the first layer, countries are connected with weights proportional to the amount of yearly
trade in the second layer and with weights proportional to the number of international organisations
they belong to together in the third layer. While the essence of commerce is to exploit differential
between the partners involved (\ie{} heterophily), trade (and in the last example participating in
common institution) intensity also involve geographical, historical and cultural ties (\ie{}
homophily).

\paragraph{Biological networks} The interactions in biology also take several forms and studying
them as a whole has proved fruitful. One example is a genes co-expression network, where each
connection was tested under 130 different experimental conditions, providing as many
layers~\autocite{bioLayerExp11}.
Multilayer graphs are also a relevant way to represent ecological
networks~\autocite{EcologyMultiReview17}. For instance, \textcite{EcoChile15} build the graph of
more than 100 species living on the Chilean
coast\footnote{\url{http://app.mappr.io/play/chile-marine-intertidal-network}} and divide their
interactions in three categories: trophic (that is one specie eats another one), positive
non-trophic (\eg{} refuge providing) and negative non-trophic (\eg{} competition for shared
resources).
Finally, in neuroscience, multilayer graphs have recently emerged as a useful tool to better
understand the human brain~\autocite{Neuroscience16}. The nodes of such graphs are neurons, and the
edges can be labeled in various ways: some correspond to actual physical links while others are
functional ---\ie{} neurons responding in the same way to external stimuli, some are present in
healthy subject and others in patient cohort, some are acquired through MRI and others by EEG.