### 2017-11-28 cc/cc_non_worst_case.tex
typos and stability in cc

We now look at two settings where our bias is likely to be present.  In the first one, we assume the
input graph is randomly perturbed version of a consistent clustering. In the second one, we do not
assume anything about such a consistent clustering but expect instead that the optimal solution is
\enquote{clear} enough that it does not change when weights are multiplicatively perturbed.

\paragraph{\pcc{} under noise}

As we have seen, assuming the Unique Games Conjecture, the \mmc{} problem, and therefore
the \pcc{} problem on a general graph, cannot be approximated to within a constant factor in the worst
case.  But maybe we can do better in the average case, which motivates the study of semi-random
model, where real graphs are seen as being obtained from the controlled perturbation of a perfectly
clusterable graph. In the simplest case, each edge sign is independently flipped with probability
$p \in [0, \shalf)$. This situation on complete graphs was considered in \autocite[Section
6]{Bansal2002}, showing a simple algorithm that with high probability makes
$\tilde{O}(n^\frac{3}{2})$ mistakes, and in \autocite[Theorem 2.6]{Ben-Dor99}, with an algorithm
recovering with high probability the planted partition of an unweighted graph in $O(n^2(\log n)^c)$,
where $c$ depends on the size of the smallest cluster and the noise probability.

\Textcite{Joachims2005} analyze a more refined weighted model, where weights are generated by a
probability distribution whose mean on true positive edges is larger than $\mu^+>0$ and whose mean
on true negative edges is smaller than $\mu^-<0$. They give a finite-sample bound on the number of
nodes misclustered w.r.t the planted partition as a function on the probability distribution
parameters. Indeed, as pointed out by \textcite{plantedAilon09}, independent and uniform noise is
not a good model of real situations, where the input of \pcc{}, \ie{} the similarity between nodes,
is often the result of a preprocessing, which may present strong correlations. Therefore, instead of
measuring the quality of a solution against the input (\ie{} the similarity information), they
argue it is more sensible to measure it against the (unknown) true optimal clustering that gave rise
to the input, and show that \ccpivot{} allows this, thanks to a new analysis. \Textcite{Mathieu2010}
also consider an adversarial model, where all edges are flipped with probability $p$ but the
adversary then decides whether to reveal the true sign or the flipped sign. On complete unweighted
graphs, they find a solution of \mind{} at most $1 + O(n^{-\nicefrac{1}{6}})$ times the optimal
whenever $p \leq \shalf - n^{-\nicefrac{1}{3}}$ by rounding the usual SDP solution. If, in addition,
there is no adversary, $p\leq\nicefrac{1}{3}$ and each planted cluster has at least
$\Theta(\sqrt{n})$ nodes, then the planted partition can be recovered exactly.
\Textcite{Makarychev2014} study the same adversarial model but on general weighted graphs, giving a PTAS
for \mind{} when $p\leq \nicefrac{1}{4}$. Under additional assumptions on the density of edges, they
present another algorithm that finds the ground truth clustering with an arbitrarily small
classification error.
% \footnote{They also mention a potentially interesting related work~\autocite{unobservedCC14}.}
% http://www.jmlr.org/papers/v15/chen14a.html
% \textcite{Noisy2CCLocality16} study the recovery of an arbitrary planted 2-partition under noisy
% measurement in special kind of graphs where each node is connected with a local neighborhood and
% show computationally efficient algorithms that only query an information theoretically optimal
% number of pair. I think it was extended in \url{https://doi.org/10.1109/TIT.2016.2600566}

\paragraph{\pcc{} under stability assumption}

Whereas clustering objective functions are \NPh{} to optimize, we expect meaningful instances
in which we are interested to have additional structure which allows for guaranteed polynomial time
algorithms. We refer the reader to a critical overview~\autocite{clusteringFeasibility15} of some
such notions of structure proposed recently. Informally, a general idea is that the clustering should not
change (or at least very little) if the data are slightly perturbed. For instance, a weighted graph
is \emph{$\alpha$-stable} (with $\alpha>1$) for some partition objective if its optimal partition
remains the same whenever every weight $w_i$ is multiplied by a factor $c_i$ between $1$ and
$\alpha$.
Another notion is the $(c, \epsilon)$-approximation stability. Formally, a dataset $X$ is $(c,
\epsilon)$-stable if, for an objective function $\Phi$, any clustering whose cost is within a factor
$c$ of the optimal cost $OPT_{\Phi}$ is $\epsilon$-close to the optimal clustering (as measured by
the fraction of points in which the two clustering disagree). When the objective is \mind{} on
complete graphs, \textcite{StableCC09} show that for $(1+\alpha, \epsilon)$-instances, the
approximation algorithms we describe in \autoref{ssub:cc_harness_approx} find a solution that is
$(\nicefrac{49}{\alpha} + 1)\epsilon$ close to the optimal clustering.

It is also possible to study the stability of a \pcc{} instance over a general graph w.r.t edge
weights via its canonical \mind{} LP. \Textcite{StableLP09} present a method to determine to which
extent the weights of a \pcc{} instance can be perturbed before the optimal solution changes.

\iffalse
\Textcite{StableCC17} provide a robust algorithm for $2-\nicefrac{2}{k}$-stable instance of
$k$-\textsc{Minimum Multiway Cut}, which therefore is not really useful in our case because I confused
multiway cut and multicut. To be fair, it seems I'm not the only one, \href{http://pages.cs.wisc.edu/~shuchi/papers/multicut-hardness-full.pdf}{on the hardness of
   approximating multicut}
   \enquote{multicut is known to be APX-hard
      (\href{https://pdfs.semanticscholar.org/1cf6/4c2bdd4f1c384a55910606a64c8d831a96ba.pdf}%
   {Dahlhaus et al. 1994}).} by that 1994 paper define the multiway cut problem, see
   \autoref{fig:cc_multiwhat}. \Textcite[Theorem 24]{Bansal2004} show that we can go from
   \textsc{Minimum Multiway Cut} to \pcc{} but I doubt about the other direction, mostly because
   \textsc{Minimum Multiway Cut} can actually be approximated with factor smaller than $1.3$.
   see \url{http://www.nowozin.net/sebastian/papers/nowozin2009lpstability.pdf} about LP multicut
   stability.
\begin{figure}[htpb]
   \centering
   \includegraphics[width=0.9\linewidth]{assets/raw/multicut_vs_multiway.png}
   \caption{Confusing statement} \label{fig:cc_multiwhat}
\end{figure}
\fi

% \enquote{We prove that if the Unique Games Conjecture of Khot (2002) is true, then for every constant L > 0
% it is NP-hard to approximate Multicut within factor L. If a quantitatively stronger version of the
% conjecture is true, then Multicut is NP-hard to approximate within a factor of $\Omega(\sqrt{\log
% \log n})$.}

% It's probably limited because LP with a lot of constraints do not scale that well (someone claimed
% $O(n^{4.5})$)

% SDP Gaps and UGC Hardness for Multiway Cut, 0-Extension and Metric Labelling,
% On the Unique Games Conjecture: Subhash Khot
% under the UGC, one cannot get better than $O(\log n)$ approximation using LP or SDP relaxation.