### 2017-12-10 conclusion/main.tex
write the conclusion

\label{chap:conclusion}

\section{Contributions}

The objective of this thesis was to provide \enquote{efficient and accurate methods to predict edge
type in complex networks}. By complex networks, we mean graphs where edges can have one among $k$
semantics (or types) and where nodes are connected by both partial homophily and heterophily.

\begin{itemize}

  \item In \autoref{chap:troll}~\autocite{trollSign17}, we first focused on directed signed graphs,
    as they model social interactions in several domains, such as e-commerce and collaborations.
    Here edges have two possible semantics, positive or negative. Our goal was to solve a supervised
    binary classification problem by predicting the sign of test edges. For that, we first
    introduced a simple sign generative model. It endows nodes with two parameters --trollness and
    unpleasantness--, governing their incoming and outgoing behavior. This might seem like a
    simplistic way of expressing the notion of partial homophily and heterophily. However, we
    presented two methods to approximate this Bayes predictor, suggesting a trade off between
    theoretical guarantees and input requirement. We showed experimentally that those methods are
    competitive with the state of the art, while being more scalable. Motivated by the regularity
    measure offered by trollness and unpleasantness, and the inability of existing methods to cope
    with new nodes joining the graph, we furthermore develop the first online algorithm for the
    \esp{} problem.

  \item We started \autoref{chap:cc} by showing that the methods of the previous chapter are not
    straightforward to extend to undirected signed graphs. Therefore, we changed our learning bias
    from a generative model to a clustering assumption inspired by social balance theories. Namely,
    we posit that the nodes are assigned to one of $K$ groups, and that the sign of the edge $(u,v)$
    is positive if the nodes $u$ and $v$ belong to the same group and negative otherwise. We then
    presented two approaches to leverage this bias. First, we drew a connection between the \esp{}
    problem and the \pcc{} problem. While \pcc{} is hard to approximate in general, we suggested,
    based on existing works, that the more an instance complies with our bias, the better \pcc{}
    methods could recover the $K$ underlying groups and thus provide accurate prediction. Second, we
    addressed the problem in an active setting, where $K=2$. For that, we implemented a promising
    spanning tree algorithm~\autocite{gtxFabio}, that queries some edges in order to connect the
    remaining edges by paths as short as possible. Experimentally, we showed that on both synthetic
    and real graphs, our method is competitive in terms of sign prediction with a strong natural
    baseline.

  \item Finally, in \autoref{chap:edge}, we considered the case where there are more than two types
    of edge. This means that instead of signed graphs, our input shifted to multilayer graphs, which
    are also ubiquitous to describe real world phenomena. Moreover, we modified the learning
    setting, so that no edge label are available, but instead nodes carry a profile vector. However,
    our goal remained to predict edge type, in what could then be described as a clustering problem
    with side information and interpretability requirements. In order to do so, we first defined a
    similarity between nodes, which accommodate both heterophily and homophily. We then formulate
    the problem as simultaneously finding $k$ special vectors called directions, and assigning one
    of them to every edge in order to maximize a scalar \emph{goodness} function. After introducing
    baselines oblivious of the graph topology, we added constraints resulting in one vectorial and
    one matricial optimization objectives. Finally, we showed on synthetic data how those methods
    can recover the planted ground truth.

\end{itemize}

\section{Future work}

In this section, we revisit the problems summarized above by showing how they could be applied in a
new context or enriched by existing node embedding methods.

\subsection{Reciprocal recommendation}

In this paragraph, we present the reciprocal recommendation problem~\autocite{Pizzato2013}. As we
shall see, it incorporates aspects from all the problems we discussed previously. Broadly speaking,
given two sets of users $L$ and $R$, we want to recommend users in $R$ to users in $L$, and vice
versa. These recommendations are based on mutual interest, but in most cases there is an asymmetry
between the two groups. To better illustrate this general definition, we now give some concrete
examples.

\begin{itemize}
  \item One recent application that quickly rose to prominence is online
    dating~\autocites{Krzywicki2015}{Xia2016a}{Alanazi}. In the most common case, users are divided
    between men and women, and we want to recommend users of the opposite gender in order to create
    couples. A well known example of such a system is Tinder~\autocite{TinderDesc16}.
  \item In another domain, we can also match job seekers and
    employers~\autocites{Siting2012}{Hong2013}{Kille2015}.
  \item As a specific case in point, \textcite{Zhang2016a} propose a method to match prospective PhD
    students and their future supervisors.
  \item In large organizations or associations, having a mentor is usually beneficial for
    both employees and employers~\autocite{mentorBenef04} and finding mentee/mentor pairs can be
    cast as a reciprocal recommendation problem~\autocite{OSSMentors12}.
    %\url{http://ieeexplore.ieee.org/abstract/document/7207130/}
  \item After moving to their new jobs, people also need a place to live. Again, we can match
    tenants and landlords, or people to share a flat.
  \item Finally, in a more critical setting, finding relevant matches between patients and organ
    donors can improve the success of transplants~\autocite{patientDonor17}.
\end{itemize}

Reciprocal recommendations present specific challenges~\autocite{Andrews2015}. Among them are
asymmetry and volume.  Asymmetry refers to the imbalance between the two groups, in terms of size
and bargaining power. For instance, depending of the economic situation and the field, there might
be a lot of job seekers for a few open positions, in which case employers can be very selective.
Likewise, it is generally observed that dating websites have a higher proportion of men, meaning
that women can afford to choose among numerous suitors~\autocite{TinderDesc16}. By volume, we mean
that users on both side can only handle a few recommendations at the same time. This is sharp
contrast with traditional user-item recommendation systems, where recommending simultaneously the
same item (say a movie) to thousands of users is not an issue. On the other hand, we have to keep
user engaged and prevent them from being idle. This is done by ensuring they are regularly the
subject or object of some recommendations.

As the name implies, reciprocal recommendation has often be seen as a special case of recommender
system~\autocite{Andrews2015}. Besides this paradigm, we could also cast it as an adaptive, active,
cost sensitive \esp{} problem on bipartite directed signed graphs with side information. Let us
decompose this last sentence, starting with the input graph. It is naturally bipartite, because of
the presence of two groups in which there is no internal recommendation. Furthermore, every time one
user expresses interest or disinterest in another user, this creates a directed signed edge. Such a
graph is thus a special case of the \dssn{} studied in \autoref{chap:troll}. Each nodes carries two
kind of information. The first is explicitly provided by the user, in order to arise interest in
users from the other group. The second is implicitly inferred from the user activity. In particular,
the interactions with other nodes allow us to build an implicit preference model for every user.
Combining these two sources of information yield a profile of each user, like in
\autoref{chap:vector}. Making recommendations in such a graph therefore boils down to finding non
existing positive edges. However, to cope with the two challenges presented in the previous
paragraph, we follow the setting of \autoref{chap:cc} and use an active approach. More precisely,
respecting the volume constraints would require an adaptive algorithm, which acquires the most
useful information without over- or under-loading any nodes. As for the group asymmetry, this could
be addressed by using a cost sensitive classification error, so that the error rate at a given node
is inversely proportional to its importance in the network.

Unfortunately, we did not study this problem further in this thesis. Indeed, getting data and
testing algorithms usually require an industrial partner.\footnote{See for
instance~\autocites{Andrews2015}{Kille2015}{Krzywicki2015}{Xia2015}{Alanazi}}

\subsection{Representation learning}
\label{par:representation_learning}

As we mentioned already in \autoref{sec:troll_related} and \autoref{sec:edge_related}, a recent and
promising trend regarding learning in graph is node embedding. In addition to the references cited
earlier, this also witnessed by recent works on embedding attributed nodes, either in signed
graphs~\autocites{SignedFS17}{attributedSNE17}{SHINE18} or in general
graphs~\autocites{Liao2017}{AANE17}{DynamicEmbedding17}. At first sight, those methods seem
radically different from traditional graph methods. However, just like word embedding methods have
been linked to matrix
factorization~\autocites{WordEmbeddingAsFactorization14}{Hellinger14}{AroraRandWalk16}, so have been
popular and successful node embedding methods~\autocite{embeddingAsFactorization18}. These methods
are the simple ones that do not consider node attributes and multiple types of edges, such as
DeepWalk~\autocite{DeepWalk14}, LINE~\autocite{LINE15} and node2vec~\autocite{node2vec16}. For
instance, \textcite{embeddingAsFactorization18} show that the node embedding of DeepWalk on the
weighted graph $G$ are approximately the left singular vectors of the following matrix:
\begin{equation}
  \label{eq:deepwalk}
  \log \left( vol(G) \left( \frac{1}{T} \sum_{r=1}^T (D^{-1}A)^r\right) D^{-1} \right) - \log b
\end{equation}
where $vol(G)$ is the sum of all weighted degrees, $T$ is the context window size, $D$ is the
diagonal matrix of degrees, $A$ is the adjacency matrix of the graph and $r$ is the number of
negative samples in skip-gram. There are several ways such findings could be used in our context:
\begin{itemize}

  \item If the edges of the input graph are fully labeled into $k$ types, then we have $k$ matrices
    $D_k$ and $A_k$, by considering the subgraphs induced by each edge type. If we simply decompose
    the $k$ matrices constructed according to equation \eqref{eq:deepwalk}, we would have $k$
    embedding for each node. Instead, it seems natural to stack those matrices together to form two
    third order tensors $\mathcal{D} \in \Rbb^{n \times n \times k}$ and $\mathcal{A} \in \Rbb^{n
    \times n \times k}$. One could then use one tensor decomposition
    methods~\autocites{tensorReview09}{tensorReview17} to extract node embedding taking into account
    the multiplicity of edge types.

  \item If the edges of the input graph are partially labeled, then predicting edge type can be seen
    as a tensor completion problem. Like in the matrix case, we seek a low rank tensor that
    coincides with the observed entries of the input data. Furthermore, when we have access to node
    profiles, we can use methods tailored to leverage such side information~\autocite[Section
    4]{tensorCompletion17}.

  \item If the input graph has no edge labels at all but we are given node profiles, we can use
    existing attributed embedding approaches~\autocites{Liao2017}{AANE17}{DynamicEmbedding17} to
    generate extra information about each node, before applying the methods presented in
    \autoref{chap:vector} on these extended profiles. We note however that
    doing so might reduce the interpretability of the final result.

\end{itemize}