### 2017-12-13 final.tex
start changing the layout

\documentclass[a4paper,svgnames,twoside,onecolumn,final,titlepage,11pt]{report}
\input{pmin}
\input{abbreviations}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\title{Characterizing edges in signed and vector-valued graphs}
\author{Géraud Le Falher}

\begin{document}
\newgeometry{vmargin=1.5cm,inner=3.8cm,bindingoffset=0.0cm,outer=3.3cm,marginparsep=0pt,marginparwidth=0pt}

\maketitle

% \section*{Abstract}
% \input{abstract}
%
% \tableofcontents
% \newpage
%
% \input{notations}
%
% \chapter{Introduction}
% \subimport*{intro/}{main}

\chapter{On the Troll-Trust Model for edge sign prediction in Social Networks}
\subimport*{troll/}{main}
\end{document}

\chapter{Edge sign prediction in general graphs and Correlation Clustering}
\subimport*{cc/}{main}

\chapter{Edge clustering in node attributed graphs}
\subimport*{edge/}{main}

\chapter{Conclusion}
\subimport*{conclusion/}{main}

\vspace{\baselineskip}
\begingroup
\newgeometry{hmargin=1.5cm,vmargin=1.3cm}
\setstretch{0.87}
\setlength\bibitemsep{1pt}
\todos
\printbibliography
\restoregeometry
\endgroup
\end{document}