### 2017-12-17 troll/related.tex

\section{Related work}
\label{sec:troll_related}

Interest in signed networks can be traced back to the psychological theory of structural
balance~\autocites{Cartwright56}{HeiderBook58} and its weak version~\autocite{davis1967clustering},
that we will describe with more details in \autoref{sub:bias_balance} \vpageref{sub:bias_balance}.
The advent of online signed social networks has enabled a
more thorough and quantitative understanding of that phenomenon. In this section, we provide an
overview of methods tackling the same \esp{} problem as us. Along the way, we give five of them a
name in small capitals, for they are recent and effective. Therefore, we will compare our approaches
with those methods in the experiments of the next section. At the end, we mention some closely
related variants of the original \esp{} problem.

\bigskip

Existing methods can be broadly divided into three strategies, which share some similarities:
\begin{enumerate}[1),nosep]
	\item embedding the nodes of the graph in a low dimensional space using spectral or neural
		techniques, before using node positions as features for a classifier;
	\item completing the adjacency matrix through a global optimization algorithm; and
	\item computing local features of the edges with several heuristics and train a classifier such as
		logistic regression or SVM.
\end{enumerate}

In the first direction, the spectral embedding is illustrated by works from
\textcites{Kunegis2009}{SignedEmbedding15}. We shall describe them more thoroughly in
\autoref{par:cc_spectral}, but here we note that, following the natural orientation of the study of
graph spectrum, they focus more on clustering than \esp{}. Furthermore, the
use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
makes those methods hardly scalable to large graphs. More recently, there has been great interest in
adapting word embedding techniques such as \textsf{word2vec}~\autocite{word2vec13} to a
\enquote{corpus} of random walks that are considered as documents, while nodes play the role of
word. This allows the unsupervised extraction of node features that can then used to train
downstream classifier, see
\autocites{nodeEmbeddingsurvey16}{representationLearning17}{graphEmbeddingSurvey17} for three recent
surveys. It can also be
tailored to exploit the specificity of signed graphs.
The goal is to find for every node $u$ a vector $x_u\in\Rbb^d$ in such a way that in this new space,
nodes are close to their positive neighbors and far from their negative neighbors. For instance in
SiNE~\autocite{neuralSigned17}, for a node triplet $u,v,w$ such that the edge $(u,v)$ is positive
and the edge $(u,w)$ is negative, the objective is to find vectors maximizing $f(x_u, x_w) + \delta
- f(x_u, x_v)$. The similarity function $f$ is chosen to be a Siamese multilayers neural network,
whose parameters are learned by back propagation. In SIGNet~\autocite{SIGNet17}, the similarity
function is a sigmoid whose argument $x_u^Tx_v$ is weighted by the sign of the corresponding edge,
with an optimization technique closer to \textsf{word2vec}. Finally, SNE~\autocite{SNE17} uses a
log-bilinear model, and the training objective is to predict, given a path $\mathcal{P}$, which node
should follow $\mathcal{P}$, weighting differently the vector nodes in $\mathcal{P}$ depending of
whether they are the source of a positive or negative edge in $\mathcal{P}$.

\medskip

Next we look at matrix completion approaches, which are global by nature.
For instance, \Textcite{LowRankCompletion14} tackle the \esp{} problem through this lens,
restricting themselves to undirected graphs. They consider the
observed adjacency matrix $A$, made of the edges in the training set \trainset{}, as a noisy
sampling of the adjacency matrix $A^\star$ of an underlying complete graph satisfying the weakly
balance condition (that is with no cycle containing only one negative edge, see
\autoref{def:weak_balance} \vpageref{def:weak_balance}).
This condition implies the existence of a small number $k$ of node clusters with positive links
within clusters and negative links across clusters, which in turn implies $\rank A^\star=k$.  By
recovering a complete matrix $\tilde{A}$ that matches the non-zeros entries of $A$, it is possible
to predict the sign of $(u,v) \notin \trainset$ as $\yhat_{u,v} = \sgn(\tilde{A}_{u,v})$. Although
the exact version of this problem is \NPh{}, the authors assume that $k$ is an hyperparameter known
beforehand and look for two matrices $W,H\in \mathbb{R}^{k\times|V|}$ that minimise a sigmoidal
reconstruction loss of $A$, subject to a nuclear norm regularization term. The minimization is
carried out by Stochastic Gradient Descent and we refer to this method as \emph{\complowrank{}}.
The approach of \textcite{OnlineCompletion17} is similar, but they consider directed graphs, use the
logistic loss and compute at each iteration a threshold optimizing the $F_1$-score on the observed
signs. Furthermore, they argue that to better handles class imbalance, it is preferable to minimize
the maxnorm of the recovered matrix, which is a tighter approximation to the rank function than the
nuclear norm. We thus refer to this method as \compmaxnorm{}.

\medskip

Finally, the last set of methods
are based on the computation of local features of the graph. These features are
evaluated on the subgraph induced by the training edges, and the resulting values are used to train
a supervised classification algorithm. The most basic set of local
features used to classify a given edge \euv{} are defined by
$\din^+(v),\din^-(v),\dout^+(u),\dout^-(u)$ computed over the training set \trainset{}, and by the
embeddedness coefficient $\big|\Nout(u) \cap \Nin(v)\big|$, which is the number of
common neighbors of $u$ and $v$. In turn, these degree features can be used to define more
complicated features, such as a notion of similarity between two nodes based on how they rate and
are rated by their neighbors~\autocite{Yuan2017}.
Another way of looking at neighborhoods and degrees is to mine ego networks\footnote{Recall that the
ego network of a node $u$ is the subgraph induced by the neighbors $\nei(u)$ of $u$.} with a
SVM~\autocite{Papaoikonomou2014}. \Textcite{Bachi2012} also use an approach based on ego networks,
in a data mining fashion. Namely, they extract frequent small subgraphs from the collection of all
ego networks of $G$. Then, they construct rules, which are made of two frequent subgraphs differing
by a single edge.

A sophisticated take on degree features is presented by \textcite{Bayesian15}, who note that a node
can belong to one of the 16 node-types based of whether the number of its positive (respectively
negative) outgoing (respectively incoming) edges is zero or not. The number of unobserved incoming
and outgoing edges of each node $u$ let us define a 16-dimensional vector $V_u$ containing the
probability of transitioning to any other type once the unobserved signs are revealed. Then each
edge \euv{} is associated with a feature vector consisting of the outer
product of $V_u$ with $V_v$ and also including additional features such as triads count and degree
information before training a Logistic Regression model. We refer to this method as
\emph{\compbayesian{}}.

Other types of features are derived from social status theory, which posits that a positive link
from $u$ to $v$ denotes that user $u$ considers user $v$ as having a higher status or skill than
herself~\autocite{Leskovec2010}. This has implications on the distribution of the so called
\emph{triads} in the network. A triad is a triangle formed by \euv{} together with $u \rightarrow w$
(or $w \rightarrow u$) and $v \rightarrow w$ (or $w \rightarrow v$) for any $w \in \NNout(u) \cap
\NNin(v)$. Taking signs and directions into account, there are 16 possible triads but according to
the status theory some must be more represented than others. The \emph{\comptriads{}}
method~\autocite{Leskovec2010} exploits this fact by counting for each edge in the training set how
frequently it is involved in each of the 16 triad types. It also adds 7 degree features before
training a Logistic Regression model.

A third group of features is based on node ranking scores. These scores (usually one or two per
node) are computed using a variety of methods, including
\begin{itemize}
	\item	PageTrust~\autocite{de2008pagetrust}, which adapts the random walk of
		PageRank~\autocite{pagerank99}, by making walkers keep in memory a list of nodes they do not
		like. If they ever reach such a node, their walk stop. The final PageTrust is computed
		iteratively until convergence. 
	\item Prestige~\autocite{zolfaghar2010mining}, which can be seen as a compounded degree feature,
		as it is defined by $P(u) = \frac{\dinp(u)-\dinm(u)}{\din(u)}$.
	\item	exponential ranking~\autocite{traag2010exponential}, which is the fixed point of the
		equation $\pi = A^T\frac{\exp{\frac{1}{\mu}\pi}}{||\exp{\frac{1}{\mu}\pi}||_1}$. This follows
		from the discrete choice theory and by assuming that we observe the reputation with some noise
		that is double exponentially distributed with parameter $\mu$.
	\item	Bias and Deserve~\autocite{mishra2011finding}, which are defined in terms of each other in
		weighted graphs. The bias is the tendency of a node $u$ to trust/mistrust others, that is the
		difference between its opinion of a neighbor $v$ and what $v$ truly deserves according to the
		network: \[bias(u) = \frac{1}{\dout(u)}\sum_{v\in\NNout(u)}(\wuv - deserve(v)).\] The deserve of a
		node $v$ is the aggregated opinion of its in-neighbors, discounted by their bias: \[deserve(v) =
		\frac{1}{\din(v)}\sum_{u\in\NNin(v)}\left(\wuv(1-\max\{0, \sgn(\wuv)\times bias(u)\})\right).\]
	\item	Reputation and Optimism~\autocite{shahriari2014ranking}, defined for a node $u$ by
		$\frac{\sum_{v \in \NNin(u)} y_{v,u}\sigma(v)}{\sum_{v \in \NNin(u)} \sigma(v)}$ and $
		\frac{\sum_{v \in \NNout(u)} \yuv\sigma(v)}{\sum_{v \in \NNout(u)} \sigma(v)}$, where
		$\sigma(v)$ is the ranking score assigned to node $v$. The former can be seen as a weighted
		version of Prestige, while the latter is its outgoing counterpart.
	\item	TrollTrust~\autocite{wu2016troll}, which builds upon \autocite{shahriari2014ranking} but
		defines the ranking $\sigma(u)$ as the trustworthiness of $u$. It follows from a recursive
		definition of trollness based on the opinion of one node's neighbors weighted by their own
		trollness, which allow to assign a $\sigma(u)$ to each node $u$ through a set of non linear
		equations solved by an iterative method. These $\sigma$ values are used to compute Reputation
		and Optimism scores, thus providing four features for each edge, which are in turn used to train
		a Logistic Regression model for the classification task. We refer to this method as
		\emph{\compranknodes{}}.
% Computing $\pi$ requires to choose two hyperparameters of the model,
% $\beta$ and $\lambda_1$ and the authors suggest to do it by holding out some training data to
% perform cross validation but in practice we found little difference in performance in our
% implementation while it consumes a lot of time.
\end{itemize}

\bigskip

Other works have also\done\Todo{maybe that part would be better in the conclusion, especially since side
information would help with the cold start problem showed in the experiments and trust version is
one of the future direction.}
considered versions of the problem where side information related to the network is available to
the learning system. For instance, \autocite{EdgeSignsRating15} uses the product purchased on
\epi{} in conjunction with a neural network, \autocite{TrollDetection15} identifies trolls by
analysing the textual content of their post, \autocite{attributedSNE17} improves
the embedding approach of \textcite{neuralSigned17} by considering the words
written in reviews as attributes of the users,  and~\autocite{SNTransfer13} uses SVM to perform
transfer learning from one network to another.
\autocite{Tang2013} uses a matrix completion approach, approximating $A$ by $PCP^T$, where the
row $P_u$ of the matrix $P\in\Rbb^{n\times d}$ is the low rank representation of the node $u$ and
$C\in\Rbb^{d\times d}$ is the correlation matrix between these representations. They also assume
they are given a symmetric matrix $Z$ of the homophily coefficients $\zeta_{u,v}$ between $u$ and
$v$ and add the following regularization term to be minimized: $\sum_{u<v} \zeta_{u,v}||P_u -
P_v||_2^2 = Tr(U^T L_Z U)$, where $L_Z$ is the Laplacian of $Z$.
While many of these approaches have interesting
performances, they often require extra information which is not always available (or reliable) and,
in addition, may face severe scaling issues.

\medskip

Whereas our focus is on \emph{binary} prediction, researchers have also considered a weighted
version of the problem, where edges measure the amount of trust or distrust between two users. Note
that typically, the embedding methods we discussed at the beginning of this section are able to
handle weighted networks.
%
One of the early and influential work on modeling how distrust propagate among users is
\autocite{guha2004propagation}. They propose to represent atomic conclusions (such as if $u$ trusts
$v$ and $v$ trusts $w$ then $u$ is likely to trusts $w$) as matrix operators and define four of them
that are assembled by a weighted linear combination. Starting from the observed adjacency matrix
$A$, they repeatedly apply this operator (with potentially a discount factor) to obtain a final
weigh matrix $F$ (that can also be rounded to provide sign prediction).
%
\Textcite{Qian2014sn} extend the binary case with categorical relationships (such as strongly
positive or weakly negative) and describe how every unbalanced triads experience some stress
depending on the strength of its contradictory relationships. Arguing that the network should
converge to a balanced state with as little change as possible, they express finding this minimal
transformation as a Multidimensional Scaling problem, and the resulting graph to characterize
unlabeled edges.
%
Finally, \textcite{edgeWeights16} explicitly consider the weight prediction problem in signed graphs.
They use a procedure similar to the ranking methods discussed above. Namely, they defined two scores
for each node that are computed iteratively from a uniform initialization. Assuming that weights lie
in $[-1,1]$, the fairness $f(u)$ of $u$ is a measure of how fair or reliable $u$ is when rating
others nodes, while the goodness $g(v)$ of $v$ measure how trustworthy is $v$ when evaluated with
complete fairness.  Formally, $f(u) =  1 - \frac{1}{2\NNout(u)} \sum_{v\in\NNout(u)}
\frac{|\wuv-g(v)|}{2}$ and $g(v) = \frac{1}{\NNin(v)}\sum_{u\in\NNin(v)}f(u)\wuv$. Note that these
definitions are very close to bias and deserve from \textcite{mishra2011finding}, but the absolute
value in the factor $2$ in the expression of fairness allows for better convergence property.
Finally, the weight of a test edge \euv{} is predicted as $f(u)\times g(v)$.

\medskip

While we presented methods that operate in the batch setting, and will present an online algorithm
as well in \autoref{s:algonline}, other works have addressed the \esp{} problem from an active
learning point of view~\autocites{Cesa-Bianchi2012a}{Cesa-Bianchi2012b}.  Recall that in the active
setting, we are given a budget of edge labels to observe, and are free to select them the way we
want within $E(Y)$. Again, the goal is to make as few mistakes as possible when predicting the sign
of the remaining edges. These two methods build a spanning subgraph of $G$ and query all its edge,
but they differ in its the construction. \Textcite{Cesa-Bianchi2012a} partition $G$ into stars and
connect them in a tree, whereas \textcite{Cesa-Bianchi2012b} cover the graph with cycles, each
containing one test edge and being queried for the other, while respecting the user specified query
budget.  This relies on a different bias than our generative model, and this will be the subject of
our next chapter.  One might also consider an online version of the problem where the topology is
not known in advance but discovered as prediction are made~\autocite{Gentile2013}. This naturally
increases and the difficulty of the problem, as reflected by the computational cost of the solution
proposed, which is quadratic in $|V|$ \emph{at each prediction}.

The survey~\autocite{Tang2015a} contains pointers to many papers on signed networks, in particular
for the \esp{} problem.