### 2017-12-18 cc/bias_main.tex

Let us recall first the sign model we used for \dssn{}. Each node $u$ is endowed with two parameters
$p_u$ and $q_u$ drawn from an arbitrary joint distribution $\mu$ over $[0,1]\times [0,1]$. $p_u$ can be
interpreted as the tendency of $u$ to send positive edges to other nodes (\ie{} the
\enquote{niceness} of $u$), while $q_u$ can be interpreted as the tendency of $u$ to receive positive
edges from other nodes (\ie{} the \enquote{popularity} of $u$). The sign of an edge \euv{} is then
chosen to be positive with probability $\frac{p_u+q_v}{2}$. This suggests that nodes in the
graph have a form of agency, and further imposes that edges are directed. We show experimentally that
failing to meet these two assumptions harms the performance of our previous methods. This is not
surprising, for our bias is no more justified in that case. Therefore, we suggest a different bias,
drawing heavily on social balance theories, although we shall see later this holds for many
nonsocial graphs as well.

\subsection{Sign generative model and behavior}
\label{sub:bias_sign_behavior}

\input{bias_model}

\subsection{Directed edges requirement}
\label{sub:need_for_a_directed_graph}

\input{bias_directed}

\subsection{Social balance as a learning bias}
\label{sub:bias_balance}

\input{bias_balance}