### 2017-12-21 assets/tikz/cc_4triads_tikz.tex

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
\begin{tikzpicture}[auto,vertex/.append style={minimum width=5mm},
  edge/.append style={line width=1.2pt,color=Black}]

  \node[vertex,medium,] (t11) at (0,0) {};
  \node[vertex,medium,] (t12) at (2,0) {};
  \node[vertex,medium,] (t13) at (1, 1.732) {};
  \draw[edge,Green] (t11) -- node[below] {$+$} (t12);
  \draw[edge,Green] (t11) -- node[left] {$+$} (t13);
  \draw[edge,Green] (t12) -- node[right] {$+$} (t13);
  \node[anchor=north,font=\small,scale=.8] (t1) at (1, -.5) {$T_3$ Mutual friends};

  \node[vertex,medium,] (t21) at (3.5,0) {};
  \node[vertex,medium,] (t22) at (5.5,0) {};
  \node[vertex,medium,] (t23) at (4.5, 1.732) {};
  \draw[edge,Green] (t21) -- node[below] {$+$} (t22);
  \draw[edge,Red] (t21) -- node[left] {$-$} (t23);
  \draw[edge,Red] (t22) -- node[right] {$-$} (t23);
  \node[anchor=north,font=\small,scale=.8] (t1) at (4.5, -.5) {$T_1$ Common enemy};

  \draw[edge,black,very thick] (-.20, 2.15) -- node[] {Balanced} (6.5, 2.15);

  \node[vertex,medium,] (t31) at (7,0) {};
  \node[vertex,medium,] (t32) at (9,0) {};
  \node[vertex,medium,] (t33) at (8, 1.732) {};
  \draw[edge,Red] (t31) -- node[below] {$-$} (t32);
  \draw[edge,Red] (t31) -- node[left] {$-$} (t33);
  \draw[edge,Red] (t32) -- node[right] {$-$} (t33);
  \node[anchor=north,font=\small,scale=.8] (t1) at (8, -.5) {$T_0$ Mutual enemies};

  \node[vertex,medium,] (t41) at (10.5,0) {};
  \node[vertex,medium,] (t42) at (12.5,0) {};
  \node[vertex,medium,] (t43) at (11.5, 1.732) {};
  \draw[edge,Red] (t41) -- node[below] {$-$} (t42);
  \draw[edge,Green] (t41) -- node[left] {$+$} (t43);
  \draw[edge,Green] (t42) -- node[right] {$+$} (t43);
  \node[anchor=north,font=\small,scale=.8] (t1) at (11.5, -.5) {$T_2$ Antagonist friends};
  \draw[edge,black,very thick,below] (-.20, -1.2) -- node[] {Weakly Balanced} (10, -1.2);
\end{tikzpicture}
\end{document}