### 2018-04-09 assets/tikz/defense_original_tikz.tex
material for the talk

\documentclass[svgnames,11pt]{standalone}
\input{preambule}
\begin{document}
       \begin{tikzpicture}[auto, every node/.style={transform shape},baseline=(b.base)]
         \tikzset{>=latex}
         \tikzstyle{peers}=[draw,circle,black,minimum width=9mm,inner sep=2pt]
         \tikzstyle{nedge}=[draw,rectangle,black,minimum width=8.5mm,minimum height=8.5mm, inner sep=0]
         \tikzstyle{edge}=[line width=2pt,color=Black,->]
         \tikzstyle{uedge}=[line width=2pt,color=Black]
         \tikzstyle{pnode}=[fill=White]
         \tikzstyle{elabel}=[color=Black]
         \tikzstyle{qnode}=[fill=White]
         \tikzstyle{posc}=[color=Green]
         \tikzstyle{negc}=[color=red]
         \foreach \place/\name in { {(0,0)/1}, {(3,0)/2}, {(0,-3)/3}, {(3,-3)/4}}
         \node[peers,pnode] (\name) at \place {\large $\name$};
         \draw[edge] (3) -- node [posc]                            {$+$}  (1);
         \draw[edge] (3) -- node [posc]                           {$+$}  (4);
         \draw[edge, dashed] (3) -- node [negc]                    {$-$}  (2);
         \draw[edge, dashed] (2) -- node [negc]                    {$-$}  (4);
         \draw[edge] (1) edge [bend left =10] node [posc]         {$+$} (2);
         \draw[edge, dashed] (2) edge [bend left =10] node [negc] {$-$} (1);
         \node[] (caption) at (0.5, 0.5) {$G$};
         \node[] (b) at (1.5, -1.5) {};
       \end{tikzpicture}
\end{document}